### 2013-09-09 Paper_Markdown.Rmd
Initial commit. RMarkdown file, R code and date from trial run.  MDS should remain unchanged, although changing the standard taxonomy might result in change to MDS distances.  `wa.res` is temporary, intended to facilitate some plotting.

Non-analogues in paleoecological reconstruction - the effects of analogue distance
========================================================

```{r}
source('R/data_setup.R')

```

```{r SiteLocations, fig.width=7, fig.height=6, fig.cap='Figure: North American site locations'}
plot(climate[,2:3], pch=19, cex=0.6, col=rgb(0.1, 0.1, 0.1, 0.3))

```

```{r SiteMDS, fig.width=7, fig.height=6, fig.cap='Figure: North American site MDS scores'}

mds <- data.frame(pol.mds$points)

ggplot(data = mds, aes(x = X1, y = X2)) +
  geom_point() + theme_bw()

```

```{r MinMaxDissim, fig.width=7, fig.height=6, fig.cap='Figure: North American site locations'}

dissim.scores <- data.frame(dissim = c(min.diss, max.diss),
                            class   = c(rep('Min', length(min.diss)),
                                        rep('Max', length(max.diss))))

ggplot(dissim.scores) + 
  geom_density(aes(x = dissim, fill=class), alpha=0.5) +
  facet_wrap(~class, scales='free') +
  theme_bw()

```

### 2013-09-09 R/bootstrap_mat.R

@@ -1,0 +1,97 @@
+library(snowfall)
+
+sfInit(parallel = TRUE, cpus = 3)
+
+quantiles <- c(0, quantile(ecdf(diag.dist), c(seq(0, 0.2, by=0.001), seq(0.21, 1, by=0.01))))
+
+vals <- c(seq(0, 0.2, by=0.001), seq(0.21, 1, by=0.01))
+
+wapls.res <- list(mean_prediction  = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                  sample_size      = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                  bias             = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                  variance         = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                  expectation      = matrix(ncol=length(vals), nrow=nrow(new.pol)))
+
+i <- 1; j <- 1
+
+subset.pol <- function(set){
+  
+  calib.samples <- sample((1:length(set))[set], replace=TRUE)
+  
+  zeros <- colSums(new.pol[calib.samples,], na.rm = TRUE) == 0
+  
+  list(calib.pol  = new.pol[calib.samples,!zeros],
+       calib.clim = climate[calib.samples,4],
+       zeros = zeros)
+
+}
+
+wapls.run <- function(j){
+  
+  set <- keep.pol[j,]
+  
+  x <- subset.pol(set)
+  
+  zeros <- colSums(x$calib.pol, na.rm = TRUE) == 0
+  
+  wapls <- try(WAPLS(y=x$calib.pol[,!zeros], 
+                     x = x$calib.clim))
+  
+  if (length(wapls) > 1){
+    pred.wapls <- predict(wapls, newdata = new.pol[j,!zeros])
+    output    <- pred.wapls$fit[which.min(performance(wapls)$object[,1])]
+    }
+  else{
+    output <- NA
+  }
+  return(output)
+  
+}
+
+#  Parallelize:
+sfExport(list = list('wapls.run'))
+sfExport(list = list('new.pol'))
+sfExport(list = list('subset.pol'))
+sfExport(list = list('climate'))
+sfLibrary(rioja)
+
+
+for(i in i:length(vals)){
+  #  This runs through each analogue distance
+  keep.pol <- aaply(diag.dist, 1, 
[…]

### 2013-09-09 R/bootstrap_wa.R

@@ -1,0 +1,100 @@
+library(snowfall)
+
+sfInit(parallel = TRUE, cpus = 6)
+
+quantiles <- c(0, quantile(ecdf(diag.dist), c(seq(0, 0.2, by=0.001), seq(0.21, 1, by=0.01))))
+
+vals <- c(seq(0, 0.2, by=0.001), seq(0.21, 1, by=0.01))
+
+
+wa.res <- list(mean_prediction  = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+               sample_size      = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+               bias             = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+               variance         = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+               expectation      = matrix(ncol=length(vals), nrow=nrow(new.pol)))
+
+if('wa.res.RData' %in% list.files('data/')) {
+  load('data/wa.res.RData')
+}
+
+i <- 1; j <- 1
+
+subset.pol <- function(set){
+  
+  calib.samples <- sample((1:length(set))[set], replace=TRUE)
+  
+  zeros <- colSums(new.pol[calib.samples,], na.rm = TRUE) == 0
+  
+  list(calib.pol  = new.pol[calib.samples,!zeros],
+       calib.clim = climate[calib.samples,4],
+       zeros = zeros)
+
+}
+
+wa.run <- function(j){
+  
+  set <- keep.pol[j,]
+  
+  x <- subset.pol(set)
+  
+  zeros <- colSums(x$calib.pol, na.rm = TRUE) == 0
+  
+  pred.wa <- try(predict(wa(x=x$calib.pol, 
+                            env = x$calib.clim, 
+                            deshrinking='monotonic'), 
+                         newdata = new.pol[j,!x$zeros]))
+  
+  if(length(pred.wa) > 1){
+    
+    output <- pred.wa$pred$pred
+    
+  }
+  else{
+    output <- NA
+  }
+  return(output)
+  
+}
+
+#  Parallelize:
+sfExport(list = list('wa.run'))
+sfExport(list = list('new.pol'))
+sfExport(list = list('subset.pol'))
+sfExport(list = list('climate'))
+sfLibrary(analogue)
+
+
+for(i in i:length(vals)){
+  #  This runs through each analogue distance
[…]

### 2013-09-09 R/bootstrap_wapls.R

@@ -1,0 +1,90 @@
+library(snowfall)
+
+sfInit(parallel = TRUE, cpus = 3)
+
+quantiles <- c(0, quantile(ecdf(diag.dist), c(seq(0, 0.2, by=0.001), seq(0.21, 1, by=0.01))))
+
+vals <- c(seq(0, 0.2, by=0.001), seq(0.21, 1, by=0.01))
+
+wapls.res <- list(mean_prediction  = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                  sample_size      = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                  bias             = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                  variance         = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                  expectation      = matrix(ncol=length(vals), nrow=nrow(new.pol)))
+
+i <- 1; j <- 1
+
+subset.pol <- function(set){
+  
+  calib.samples <- sample((1:length(set))[set], replace=TRUE)
+  
+  zeros <- colSums(new.pol[calib.samples,], na.rm = TRUE) == 0
+  
+  list(calib.pol  = new.pol[calib.samples,!zeros],
+       calib.clim = climate[calib.samples,4],
+       zeros = zeros)
+
+}
+
+wapls.run <- function(j){
+  
+  set <- keep.pol[j,]
+  
+  x <- subset.pol(set)
+  
+  zeros <- colSums(x$calib.pol, na.rm = TRUE) == 0
+  
+  wapls <- try(WAPLS(y=x$calib.pol[,!zeros], 
+                     x = x$calib.clim))
+  
+  if (length(wapls) > 1){
+    pred.wapls <- predict(wapls, newdata = new.pol[j,!zeros])
+    output    <- pred.wapls$fit[which.min(performance(wapls)$object[,1])]
+    }
+  else{
+    output <- NA
+  }
+  return(output)
+  
+}
+
+#  Parallelize:
+sfExport(list = list('wapls.run'))
+sfExport(list = list('new.pol'))
+sfExport(list = list('subset.pol'))
+sfExport(list = list('climate'))
+sfLibrary(rioja)
+
+
+for(i in i:length(vals)){
+  #  This runs through each analogue distance
+  keep.pol <- aaply(diag.dist, 1, 
+                    function(x) {x > quantiles[i]}, 
[…]

### 2013-09-09 R/bootstrapping.noaanalogs.R

@@ -1,0 +1,156 @@
+library(analogue)
+library(MASS)
+library(plyr)
+library(rioja)
+library(randomForest)
+library(vegan)
+
+climate <- read.csv('data/Clim_Pollen_Zirbel.csv', row.names=1)
+mod_pol <- read.csv('data/ModP35c_East.csv', row.names=1, sep='\t')
+
+new.pol <- mod_pol[,7:ncol(mod_pol)] / rowSums(mod_pol[,7:ncol(mod_pol)])
+
+climate <- climate[rownames(mod_pol),]
+
+dists <- dist(new.pol^(1/2))
+
+dists[dists==0] <- min(dists[dists > 0], na.rm=TRUE)
+
+if(!'pol.mds.RData' %in% list.files('data/')){
+  pol.mds <- isoMDS(dists, k=2)
+  save(pol.mds, file='data/pol.mds.RData')
+}
+else{
+  load('data/pol.mds.RData')
+}
+
+#  Find the density of minimum distances:
+diag.dist <- as.matrix(dists); diag(diag.dist) <- NA
+
+min.diss <- apply(diag.dist, 2, min, na.rm=TRUE)
+max.diss <- apply(diag.dist, 2, max, na.rm=TRUE)
+
+min.dist.dens <- density(min.diss, from=0, to=1.42, n=100)
+max.dist.dens <- density(max.diss, from=0, to=1.42, n=100)
+
+quantiles <- c(0, quantile(ecdf(diag.dist), c(seq(0, 0.2, by=0.001), seq(0.21, 1, by=0.01))))
+
+vals <- c(seq(0, 0.2, by=0.001), seq(0.21, 1, by=0.01))
+
+mat.res <- list(prediction     = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                r.squared      = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                closest.n      = matrix(ncol=length(vals), nrow=nrow(new.pol)))
+
+wa.res <- list(prediction      = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+               r.squared       = matrix(ncol=length(vals), nrow=nrow(new.pol)))
+
+wapls.res <- list(prediction   = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                  r.squared    = matrix(ncol=length(vals), nrow=nrow(new.pol)),
[…]

### 2013-09-09 R/data_setup.R

@@ -1,0 +1,36 @@
+library(ggplot2)
+library(analogue)
+library(MASS)
+library(plyr)
+library(rioja)
+library(randomForest)
+library(vegan)
+
+climate <- read.csv('data/Clim_Pollen_Zirbel.csv', row.names=1)
+mod_pol <- read.csv('data/ModP35c_East.csv', row.names=1, sep='\t')
+
+new.pol <- mod_pol[,7:ncol(mod_pol)] / rowSums(mod_pol[,7:ncol(mod_pol)])
+
+climate <- climate[rownames(mod_pol),]
+
+dists <- dist(new.pol^(1/2))
+
+dists[dists==0] <- min(dists[dists > 0], na.rm=TRUE)
+
+if(!'pol.mds.RData' %in% list.files('data/')){
+  pol.mds <- isoMDS(dists, k=2)
+  save(pol.mds, file='data/pol.mds.RData')
+}
+
+if('pol.mds.RData' %in% list.files('data/')){
+  load('data/pol.mds.RData')
+}
+
+#  Find the density of minimum distances:
+diag.dist <- as.matrix(dists); diag(diag.dist) <- NA
+
+min.diss <- apply(diag.dist, 2, min, na.rm=TRUE)
+max.diss <- apply(diag.dist, 2, max, na.rm=TRUE)
+
+min.dist.dens <- density(min.diss, from=0, to=1.5, n=100)
+max.dist.dens <- density(max.diss, from=0, to=1.5, n=100)

### 2013-09-17 Paper_Markdown.Rmd
Took the bootstrap_wa file and rewrote it for MAT.  Some potential issues remain, partcularly at larger analogue distances, but should be enough to get started with.  Made some minor plotting changes to the paper file.

@@
 ggplot(data = mds, aes(x = X1, y = X2)) +
-  geom_point() + theme_bw()
+  geom_point(alpha=0.5, size = 2) + theme_bw() +
+  xlab('MDS Axis One') + ylab('MDS Axis Two') +
+    theme(text = element_text(size=24, family='serif'))
 
@@
 
-dissim.scores <- data.frame(dissim = c(min.diss, max.diss),
+dissim.scores <- data.frame(dissim = c(min.diss, max.diss, as.vector(dists)),
                             class   = c(rep('Min', length(min.diss)),
-                                        rep('Max', length(max.diss))))
+                                        rep('Max', length(max.diss)), rep('Dists', length(as.vector(dists)))))
 
@@
   geom_density(aes(x = dissim, fill=class), alpha=0.5) +
-  facet_wrap(~class, scales='free') +
-  theme_bw()
+  #facet_wrap(~class, scales='free') +
+  theme_bw() +
+  theme(text = element_text(size=24, family='serif'))
 

### 2013-09-17 R/bootstrap_mat.R

@@ -1,97 +1,110 @@
 library(snowfall)
 
-sfInit(parallel = TRUE, cpus = 3)
+sfInit(parallel = TRUE, cpus = 4)
 
+# Find the quantiles & set up the analogue distance values:
 quantiles <- c(0, quantile(ecdf(diag.dist), c(seq(0, 0.2, by=0.001), seq(0.21, 1, by=0.01))))
-
 vals <- c(seq(0, 0.2, by=0.001), seq(0.21, 1, by=0.01))
 
-wapls.res <- list(mean_prediction  = matrix(ncol=length(vals), nrow=nrow(new.pol)),
-                  sample_size      = matrix(ncol=length(vals), nrow=nrow(new.pol)),
-                  bias             = matrix(ncol=length(vals), nrow=nrow(new.pol)),
-                  variance         = matrix(ncol=length(vals), nrow=nrow(new.pol)),
-                  expectation      = matrix(ncol=length(vals), nrow=nrow(new.pol)))
+mat.res <- list(mean_prediction  = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                sample_size      = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                bias             = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                variance         = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                expectation      = matrix(ncol=length(vals), nrow=nrow(new.pol)))
+
+#  Turn the distance table into a matrix.
+diag.dist <- as.matrix(dists); diag(diag.dist) <- NA
 
 i <- 1; j <- 1
 
+#  For MAT to work we don't need to re-calculate the distace matrix, we just need to know the
+#  sample of interest and the subset:
+
+#  A function 
 subset.pol <- function(set){
   
   calib.samples <- sample((1:length(set))[set], replace=TRUE)
   
-  zeros <- colSums(new.pol[calib.samples,], na.rm = TRUE) == 0
-  
-  list(calib.pol  = new.pol[calib.samples,!zeros],
-       calib.clim = climate[calib.samples,4],
-       zeros = zeros)
+  list(calib.pol  = new.pol[calib.samples, ],
+       calib.clim = climate[calib.samples,4])
 
 }
 
-wapls.run <- function(j){
+fmat <- function(x){
[…]

### 2013-09-18 Paper_Markdown.Rmd
Major changes.  Plotting changes in the paper markdown, added libraries in `data_setup.R`, `bootstrap_mat.R` now works and has output `mat.res.RData`, which should be the final output.  `bootstrap_mat.R` is pretty fast as a parallel process.  Cleaned up `bootstrap_wa.R`.
Should have pushed this as a small set of changes, sorry.

@@
 ```{r}
+
 source('R/data_setup.R')
@@
 ```{r SiteLocations, fig.width=7, fig.height=6, fig.cap='Figure: North American site locations'}
+map <- map_data('world')
+
+ggplot(data = map, aes(long, lat, group=group)) + geom_path(color='black') +
+   coord_map(xlim=c(-170, -40), ylim=c(20, 83))
+
 plot(climate[,2:3], pch=19, cex=0.6, col=rgb(0.1, 0.1, 0.1, 0.3))
-
-```
-
-```{r SiteMDS, fig.width=7, fig.height=6, fig.cap='Figure: North American site MDS scores'}
-
-mds <- data.frame(pol.mds$points)
-
-ggplot(data = mds, aes(x = X1, y = X2)) +
-  geom_point(alpha=0.5, size = 2) + theme_bw() +
-  xlab('MDS Axis One') + ylab('MDS Axis Two') +
-    theme(text = element_text(size=24, family='serif'))
 
@@
 ```
+Figure X.
+Distribution of squared-chord dissimilarity values.
+The 95% confidence interval for the minumum spanning distance in the North American Modern Pollen Database is 0.10 - 0.48.
+We choose 0.48 as our cut-off for 'non-analogue' distances in the database.
+
+```{r SiteMDS, fig.width=7, fig.height=6, fig.cap='Figure: North American site MDS scores'}
+
+mds <- data.frame(pol.mds$points,
+                  non.ana = factor(min.diss > 0.48))
+
+ggplot(data = mds, aes(x = X1, y = X2)) +
+  geom_point(alpha=0.5, size = 4, aes(color = non.ana)) + theme_bw() +
+  xlab('MDS Axis One') + ylab('MDS Axis Two') +
+    theme(text = element_text(size=24, family='serif'))
+
+```

### 2013-09-18 R/bootstrap_mat.R

@@ -20,16 +20,6 @@ i <- 1; j <- 1
 #  For MAT to work we don't need to re-calculate the distace matrix, we just need to know the
 #  sample of interest and the subset:
 
-#  A function 
-subset.pol <- function(set){
-  
-  calib.samples <- sample((1:length(set))[set], replace=TRUE)
-  
-  list(calib.pol  = new.pol[calib.samples, ],
-       calib.clim = climate[calib.samples,4])
-
-}
-
 fmat <- function(x){
   #  returns the means of the 1:10 closest to calculate the RMSE for analogue selection:
   
@@ -69,7 +59,6 @@ mat.fun <- function(z, j, min){
 sfExport(list = list('diag.dist'))
 sfExport(list = list('mat.fun'))
 sfExport(list = list('new.pol'))
-sfExport(list = list('subset.pol'))
 sfExport(list = list('climate'))
 
 for(i in i:length(vals)){
@@ -108,3 +97,10 @@ for(i in i:length(vals)){
   save(mat.res, file = 'data/mat.res.RData')
   cat('\n', i)
 }
+
+par(mfrow=c(2, 1))
+plot(quantiles[-282], mat.res$mean_prediction[1,], ylim = range(mat.res$mean_prediction, na.rm=TRUE), type='l', col=rgb(0.1, 0.1, 0.1,0.1))
+for(i in 1:2289)lines(quantiles[-282], mat.res$mean_prediction[i,], col=rgb(0.1, 0.1, 0.1,0.01))
+rug(x = clim)
+plot(quantiles[-282], wa.res$mean_prediction[1,], ylim = range(wa.res$mean_prediction, na.rm=TRUE), type='l', col=rgb(0.1, 0.1, 0.1,0.1))
+for(i in 1:2289)lines(quantiles[-282], wa.res$mean_prediction[i,], col=rgb(0.1, 0.1, 0.1,0.01))

### 2013-09-18 R/bootstrap_wa.R

@@ -6,7 +6,6 @@ quantiles <- c(0, quantile(ecdf(diag.dist), c(seq(0, 0.2, by=0.001), seq(0.21, 1
 
 vals <- c(seq(0, 0.2, by=0.001), seq(0.21, 1, by=0.01))
 
-
 wa.res <- list(mean_prediction  = matrix(ncol=length(vals), nrow=nrow(new.pol)),
                sample_size      = matrix(ncol=length(vals), nrow=nrow(new.pol)),
                bias             = matrix(ncol=length(vals), nrow=nrow(new.pol)),

### 2013-09-18 R/data_setup.R

@@ -5,6 +5,9 @@ library(plyr)
 library(rioja)
 library(randomForest)
 library(vegan)
+library(maps)
+library(maptools)
+library(ggmap)
 
 climate <- read.csv('data/Clim_Pollen_Zirbel.csv', row.names=1)
 mod_pol <- read.csv('data/ModP35c_East.csv', row.names=1, sep='\t')

### 2013-09-18 Paper_Markdown.Rmd
Fixed regional plotting.

@@
 
-ggplot(data = map, aes(long, lat, group=group)) + geom_path(color='black') +
-   coord_map(xlim=c(-170, -40), ylim=c(20, 83))
+map <- subset(map, map$long > -170 & map$long < -40)
+map <- subset(map, map$lat > 20 & map$lat < 83)
 
-plot(climate[,2:3], pch=19, cex=0.6, col=rgb(0.1, 0.1, 0.1, 0.3))
+ggplot(data = data.frame(map), aes(long, lat)) + geom_path(aes(group=group), color='black') +
+  geom_point(data = data.frame(climate), aes(x = LONDD, y = LATDD)) +
+  coord_map(xlim=c(-170, -40), ylim=c(21, 83))
 

### 2013-09-26 Paper_Markdown.Rmd
Modified plotting for figures in the markdown text.

@@
   theme_bw() +
-  theme(text = element_text(size=24, family='serif'))
+  theme(text = element_text(size=24, family='serif')) +
+  scale_x_sqrt()
 
@@
 
-ggplot(data = mds, aes(x = X1, y = X2)) +
-  geom_point(alpha=0.5, size = 4, aes(color = non.ana)) + theme_bw() +
+non.ana <- ggplot(data = mds, aes(x = X1, y = X2)) +
+  geom_point(alpha=0.5, size = 4, aes(color = non.ana)) +   
+  theme_bw() +
   xlab('MDS Axis One') + ylab('MDS Axis Two') +
@@
 
+non.ana.map <- ggplot(data = data.frame(map), aes(long, lat)) + 
+  geom_path(aes(group=group), color='black') +
+  geom_point(data = data.frame(climate), 
+             aes(x = LONDD, y = LATDD, color = mds$non.ana)) +
+  coord_map(xlim=c(-170, -40), ylim=c(21, 83))
+
 ```

### 2013-09-30 Paper_Markdown.Rmd
Updated text, running wapls bootstraps, some updates (fairly rapid, must be close to the end), re-wrote the bootstrap code to be faster on an unused maching (more cores).

--- 2013-09-18 Paper_Markdown.Rmd
@@
 
+**Abstract**
+Some jibberish.
+
 ```{r}
+library(gridExtra)
 
@@
 
+Introduction
+=========================
+Pollen assemblages are widely used for paleoclimatic reconstruction during the Holocene, but this method is also applied to pollen assemblages from the last Glacial period (Bartlein et al., 2011), as well as to earlier pollen assemblages, such as those during the Eemian (Brewer et al, 2008).
+A particular challenge for climate reconstruction using these much earlier samples is the presence of non-analogue vegetation.
+Non-analogue pollen assemblages pose a problem for many reconstruction techniques, these vegetation assemblages may be the result of sets of climate variables that no longer co-exist, ecological conditions resulting from differing rates of species migration or release from herbivores (Gill et al., 2009?), or as a result of changes in human land use (St.
+Jacques?, maybe not the best.
+.
+.).
+
+Predicting climate from pollen in non-analogue space is likely to increase the uncertainty of models, possibly introduce systematic bias in predictions, and is likely to increase the variability of predictions across a time-series of predictions from a single site.
+However, high-quality terrestrial proxies for climate are important for understanding past climate change, and can act as a constraint on models of climate change in the past derived from GCMs in regions where little proxy data is available.
+Thus, understanding the behaviour of pollen-based climate reconstructions in non-analogue space is an important first stem in understanding these reconstructions.
+
+Although we cannot know the values for temperature or
+
+Methods:
+=========================
+
+Sites from across North America were used to build a modern analogue dataset.
+
 ```{r SiteLocations, fig.width=7, fig.height=6, fig.cap='Figure: North American site locations'}
@@
 
-ggplot(data = map, aes(long, lat, group=group)) + geom_path(color='black') +
-   coord_map(xlim=c(-170, -40), ylim=c(20, 83))
+map <- subset(map, map$long > -170 & map$long < -30)
+map <- subset(map, map$lat > 20 & map$lat < 83)
 
-plot(climate[,2:3], pch=19, cex=0.6, col=rgb(0.1, 0.1, 0.1, 0.3))
+ggplot(data = data.frame(map), aes(long, lat)) + geom_path(aes(group=group), color='gray') +
+  geom_point(data = data.frame(climate), aes(x = LONDD, y = LATDD), alpha = 0.5) +
+  theme_bw() +
+  xlab('') + ylab('') +
+  theme(axis.text=element_text(size=12, family='serif')) +
+  coord_map(projection='albers', lat0=30, lat1=50, xlim=c(-145, -60), ylim=c(26, 75))
 
 ```
+**Figure 1**. Location of sample plots used for analysis.  Samples are derived from a modern pollen dataset.
 
+When we 
 ```{r MinMaxDissim, fig.width=7, fig.height=6, fig.cap='Figure: North American site locations'}
@@
 dissim.scores <- data.frame(dissim = c(min.diss, max.diss, as.vector(dists)),
-                            class   = c(rep('Min', length(min.diss)),
-                                        rep('Max', length(max.diss)), rep('Dists', length(as.vector(dists)))))
+                            class   = c(rep('Minimum', length(min.diss)),
+                                        rep('Maximim', length(max.diss)), rep('All', length(as.vector(dists)))))
 
@@
   geom_density(aes(x = dissim, fill=class), alpha=0.5) +
-  #facet_wrap(~class, scales='free') +
   theme_bw() +
-  theme(text = element_text(size=24, family='serif'))
+  xlab('Squared-chord Dissimilarity') +
+  ylab('Density') +
+  theme(text = element_text(size=24, family='serif'),
+        strip.text = element_text(size=14, family='serif'),
+        axis.text = element_text(size = 12, family='serif')) +
+  scale_x_log10() + scale_y_sqrt()
 
 ```
-Figure X.
-Distribution of squared-chord dissimilarity values.
+**Figure X**.
+*Distribution of squared-chord dissimilarity values.
 The 95% confidence interval for the minumum spanning distance in the North American Modern Pollen Database is 0.10 - 0.48.
-We choose 0.48 as our cut-off for 'non-analogue' distances in the database.
+We choose 0.48 as our cut-off for 'non-analogue' distances in the database.*
 
@@
 
-ggplot(data = mds, aes(x = X1, y = X2)) +
-  geom_point(alpha=0.5, size = 4, aes(color = non.ana)) + theme_bw() +
+non.ana <- ggplot(data = mds, aes(x = X1, y = X2)) +
+  geom_point(data = mds[mds$non.ana == FALSE,], alpha=0.5, size = 2, color='black') +   
+  geom_point(data = mds[mds$non.ana == TRUE,], alpha=0.7, size = 4, color='red') +
+  theme_bw() +
   xlab('MDS Axis One') + ylab('MDS Axis Two') +
@@
 
+non.ana.map <- ggplot(data = data.frame(map), aes(long, lat)) + 
+  geom_path(aes(group=group), color='black') +
+  geom_point(data = data.frame(climate)[mds$non.ana == FALSE,], 
+             aes(x = LONDD, y = LATDD), alpha=0.5) +
+  geom_point(data = data.frame(climate)[mds$non.ana == TRUE,], 
+             aes(x = LONDD, y = LATDD), color = 'red') +
+   theme_bw() +
+  coord_map(projection='albers', lat0=30, lat1=50, xlim=c(-140, -60), ylim=c(26, 75))
+
+grid.arrange(non.ana, non.ana.map, ncol=2)
+
 ```
+**Figure 2**.
+*Location of non-analogue points in the modern pollen data, presented within a non-metric multidimensional scaling ordination (Panel 1) and geographically.
+Interestingly, although a number of geographically isolated points exist in the dataset, the majority of non-analogue points exist within the most dense region of sampling, indicating that non-analogue sites are not ecologically stratified.*
+
+References
+=========================
+Bartlein, P.
+J., Harrison, S.
+P., Brewer, S., Connor, S., Davis, B.
+A.
+S., Gajewski, K., ...
+& Wu, H.
+(2011).
+Pollen-based continental climate reconstructions at 6 and 21 ka: a global synthesis.
+Climate Dynamics, 37(3-4), 775-802.
+Brewer, S., Guiot, J., Sánchez-Goñi, M. F., & Klotz, S. (2008). The climate in Europe during the Eemian: a multi-method approach using pollen data. Quaternary Science Reviews, 27(25), 2303-2315.

### 2013-09-30 R/bootstrap_wapls.R

@@ -1,16 +1,21 @@
 library(snowfall)
 
-sfInit(parallel = TRUE, cpus = 4)
+sfInit(parallel = TRUE, cpus = 6)
 
 quantiles <- c(0, quantile(ecdf(diag.dist), c(seq(0, 0.2, by=0.001), seq(0.21, 1, by=0.01))))
 
 vals <- c(seq(0, 0.2, by=0.001), seq(0.21, 1, by=0.01))
 
-wapls.res <- list(mean_prediction  = matrix(ncol=length(vals), nrow=nrow(new.pol)),
-                  sample_size      = matrix(ncol=length(vals), nrow=nrow(new.pol)),
-                  bias             = matrix(ncol=length(vals), nrow=nrow(new.pol)),
-                  variance         = matrix(ncol=length(vals), nrow=nrow(new.pol)),
-                  expectation      = matrix(ncol=length(vals), nrow=nrow(new.pol)))
+if('wapls.res.RData' %in% list.files('data')){
+  load('data/wapls.res.RData')
+}
+if(!'wapls.res.RData' %in% list.files('data')){
+  wapls.res <- list(mean_prediction  = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                    sample_size      = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                    bias             = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                    variance         = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                    expectation      = matrix(ncol=length(vals), nrow=nrow(new.pol)))
+}
 
 i <- 1; j <- 1
 
@@ -78,7 +83,7 @@ for(i in 1:length(vals)){
       wapls.res$var[j, i]  <- mean((mean(prediction) - prediction)^2)
     }
     
-    cat('\n', j)  
+    #cat('\n', j)  
     
   }
     

### 2013-10-01 Paper_Markdown.Rmd
Began dividing data into Results section - some problems with plotting exist (really hard plots to make!). Changed bootstrap_wa to reconstruct July temperatures.

@@
 
-ggplot(data = data.frame(map), aes(long, lat)) + geom_path(aes(group=group), color='gray') +
+ggplot(data = data.frame(map), aes(long, lat)) + 
+  geom_polygon(aes(group=group), color='blue', alpha=0.2) +
   geom_point(data = data.frame(climate), aes(x = LONDD, y = LATDD), alpha = 0.5) +
@@
 
-When we 
+Results
+=========================
+
+When we examine the dissimilarity scores we can see that the distribution between minimum, maximum and overall dissimilarities can be understood numerically and spatially.
+
 ```{r MinMaxDissim, fig.width=7, fig.height=6, fig.cap='Figure: North American site locations'}
 
-dissim.scores <- data.frame(dissim = c(min.diss, max.diss, as.vector(dists)),
+dissim.scores <- data.frame(dissim = c(min.diss, 
+                                       max.diss, 
+                                       as.vector(dists)[sample(length(dists), 10000)]),
                             class   = c(rep('Minimum', length(min.diss)),
-                                        rep('Maximim', length(max.diss)), rep('All', length(as.vector(dists)))))
+                                        rep('Maximim', length(max.diss)), 
+                                        rep('All', 10000)))
 
@@
         axis.text = element_text(size = 12, family='serif')) +
+  annotate("segment", x=0.48, xend = -.32, y = 0, yend = 1, size=3, color='red') +
   scale_x_log10() + scale_y_sqrt()
@@
 
+Model outputs
+=========================
+```{r LoadOutputs, echo=FALSE, message=FALSE, warning=FALSE}
+load('data/wa.res.RData')
+load('data/mat.res.RData')
+load('data/wapls.res.RData')
+
+vals <- c(seq(0, 0.2, by=0.001), seq(0.21, 1, by=0.01))
+
+na.samples <- data.frame(size = as.vector(wa.res$sample_size),
+                         distance = rep(vals, each = nrow(wa.res$sample_size)),
+                         alpha.val = rep(diff(c(-.5,vals)), each = nrow(wa.res$sample_size))+0.005)
+
+na.samples <- na.samples[!is.na(na.samples$size),]
+
+ggplot(na.samples, aes(x=distance, y=size)) + geom_point(aes(alpha=alpha.val)) +
+  xlab('Analogue Distance') + ylab('Available Samples') +
+  theme(text = element_text(size=24, family='serif')) +
+  theme_bw() +
+  scale_alpha_identity(guide='none')
+
+```
+**Figure X**.
+*As the analogue distance exclusion is increased the available data set size is reduced, although some points continue to maintain a large number of potential analogues, this is directly related to the minimum distance to the closest analogue.
+Points that maintain a large pool of analogues are far from neighbours.*
+
+```{r PredictedValues, message=FALSE, echo=FALSE, warning=FALSE}
+
+load('data/mat.res.RData')
+load('data/wa.res.RData')
+load('data/wapls.res.RData')
+
+model.out <- data.frame(site   = rep(rep(climate[,1], ncol(mat.res$bias)),3),
+                        preds  = c(as.vector(mat.res$mean_prediction),
+                                   as.vector(wa.res$mean_prediction),
+                                   as.vector(wapls.res$mean_prediction)),
+                        dissim = rep(rep(vals, each=nrow(mat.res$bias)),3),
+                        model  = rep(c('MAT', 'WA', 'WAPLS'), each = length(mat.res$bias)))
+
+model.out <- model.out[!colSums(is.na(model.out)) > 0,]
+
+ggplot(model.out, aes(x = dissim, y = preds, group = site)) +
+  geom_path(alpha = 0.01) +
+  theme_bw() +
+  facet_wrap(~mode)
+  scale_x_continuous(expand=c(0,0)) +
+  annotate('segment', x=0, xend=1, 
+           y=mean(climate[,10], na.rm=TRUE),
+           yend=mean(climate[,10], na.rm=TRUE),
+           color='red', size=1.5, linetype=2)
+
+```
+**Figure X**.
+
 References

### 2013-10-01 Paper_Markdown.Rmd
Added site information, changed plotting outputs.

@@
 
-na.samples <- data.frame(size = as.vector(wa.res$sample_size),
+na.samples <- data.frame(site = rep(climate[,1], ncol(wa.res$bias)),
+                         size = as.vector(wa.res$sample_size),
                          distance = rep(vals, each = nrow(wa.res$sample_size)),
@@
 
-ggplot(na.samples, aes(x=distance, y=size)) + geom_point(aes(alpha=alpha.val)) +
+ggplot(na.samples, aes(x=distance, y=size, group )) + 
+  geom_path(aes(alpha=alpha.val)) +
   xlab('Analogue Distance') + ylab('Available Samples') +

### 2013-10-02 Paper_Markdown.Rmd
Expanded introductory text to introduce method outline.  Revised plotting of model-exclusion ghostplots, improved figure caption for the ghostplots.  Also fixed the MAT bootstrap to run for July temperatures and to run a lower number of quantiles given preliminary outputs.

@@
 
-Although we cannot know the values for temperature or
+Although we cannot know the values for temperature or precipitation during this time, statistical non-analogues do exist in a modern context, and we can construct false non-analogues using a modified h-block sampling procedure
 
@@
 
-ggplot(na.samples, aes(x=distance, y=size, group )) + 
-  geom_path(aes(alpha=alpha.val)) +
+ggplot(na.samples, aes(x=distance, y=size)) + 
+  geom_path(aes(group=site), alpha=0.001) +
   xlab('Analogue Distance') + ylab('Available Samples') +
   theme(text = element_text(size=24, family='serif')) +
-  theme_bw() +
-  scale_alpha_identity(guide='none')
+  theme_bw()
 
@@
   theme_bw() +
-  facet_wrap(~mode)
-  scale_x_continuous(expand=c(0,0)) +
-  annotate('segment', x=0, xend=1, 
-           y=mean(climate[,10], na.rm=TRUE),
-           yend=mean(climate[,10], na.rm=TRUE),
-           color='red', size=1.5, linetype=2)
+  facet_wrap(~model, scales='free_y') +
+  scale_x_continuous(expand=c(0,0))
 
@@
 **Figure X**.
+*Change in site predictions over time.
+Of great interest is the difference between model outputs as the exclusion zone increases.
+MAT can never predict beyond the extent of the data samples in the pool, and so as the exclusion zone increases the predictions move toward the dataset mean.
+WAPLS uses an initial transformation and then regresses against it to predict new values, so as the exclusion zone increases we see an increase in the spread of the predicted values, with much greater variability and a loss of accuracy.*
+
+Discussion
+=========================
+What does this tell us about predicting climate in non-analogue space?
 

### 2013-10-02 R/bootstrap_mat.R

@@ -1,10 +1,10 @@
 library(snowfall)
 
-sfInit(parallel = TRUE, cpus = 4)
+sfInit(parallel = TRUE, cpus = 3)
 
 # Find the quantiles & set up the analogue distance values:
 quantiles <- c(0, quantile(ecdf(diag.dist), c(seq(0, 0.2, by=0.001), seq(0.21, 1, by=0.01))))
-vals <- c(seq(0, 0.2, by=0.001), seq(0.21, 1, by=0.01))
+vals <- c(seq(0, 1, by=0.01))
 
 mat.res <- list(mean_prediction  = matrix(ncol=length(vals), nrow=nrow(new.pol)),
                 sample_size      = matrix(ncol=length(vals), nrow=nrow(new.pol)),
@@ -61,7 +61,7 @@ sfExport(list = list('mat.fun'))
 sfExport(list = list('new.pol'))
 sfExport(list = list('climate'))
 
-for(i in i:length(vals)){
+for(i in 1:length(vals)){
   #  At each quantile, figure out which samples should be acceptable for a calibration set targeting
   #  each sample return 'keep.pol'
   keep.pol <- aaply(diag.dist, 1, 
@@ -70,7 +70,7 @@ for(i in i:length(vals)){
 
   diag(keep.pol) <- FALSE
   
-  if(any(is.na(mat.res$prediction[,i]))){
+  if(any(is.na(mat.res$mean_prediction[,i]))){
     fast.mat <- laply(1:nrow(diag.dist), fmat, .progress = "text")
     rmse <- aaply(fast.mat, 2, function(x) sqrt(mean((x - climate[,10])^2, na.rm = T)))
   }
@@ -86,8 +86,8 @@ for(i in i:length(vals)){
       mat.res$mean_prediction[j,i] <- mean(prediction, na.rm=TRUE)
       mat.res$sample_size[j,i] <- sum(keep.pol[j,], na.rm=TRUE)
       mat.res$bias[j, i] <- (climate[j,10] - mean(prediction, na.rm=TRUE))^2
-      mat.res$exp[j, i]  <- mean((climate[j,10] - prediction)^2, na.rm=TRUE)
-      mat.res$var[j, i]  <- mean((mean(prediction, na.rm=TRUE) - prediction)^2, na.rm=TRUE)
[…]

### 2013-10-02 Paper_Markdown.Rmd
Minor change in bootstrap_mat, removed `cat` calls for `j` and removed text update to improve speed.  Added a bibtex file for potential references, more text editing.

@@
 
-Although we cannot know the values for temperature or precipitation during this time, statistical non-analogues do exist in a modern context, and we can construct false non-analogues using a modified h-block sampling procedure
+Although we cannot know the values for temperature or precipitation during this time, statistical non-analogues do exist in a modern context, and we can construct false non-analogues using a modified h-block sampling procedure (e.g., Telford & Birks) that examines analogue distance rather than geographical distance.
 

### 2013-10-02 R/bootstrap_mat.R

@@ -61,12 +61,11 @@ sfExport(list = list('mat.fun'))
 sfExport(list = list('new.pol'))
 sfExport(list = list('climate'))
 
-for(i in 1:length(vals)){
+for(i in i:length(vals)){
   #  At each quantile, figure out which samples should be acceptable for a calibration set targeting
   #  each sample return 'keep.pol'
   keep.pol <- aaply(diag.dist, 1, 
-                    function(x) {x > quantiles[i]}, 
-                    .progress = "text")
+                    function(x) {x > quantiles[i]})
 
   diag(keep.pol) <- FALSE
   

### 2013-10-02 Paper_Markdown.Rmd
Commented out the individual site `cat` in the `j` loop for `bootstrap_mat`, added the pCCA ratio clump into `Paper_Markdown` methods to provide better choices of climate variables.  Began writing some results text for this section.  Finished July temp MAT bootstraps.

@@
 
+Climate Variables
+=========================
+
+We used the `r nrow(climate)` records from the North American Modern Pollen Database and associated climate data to provide reconstructions of key climatic variables.
+The process of calculating the lambda/lambda ratio from RDA (ref) produces interesting relationships, perhaps impacted by the breadth of the available climate and pollen data across North America.
+
+```{r echo=FALSE, message=FALSE, warning=FALSE}
+pol <- mod_pol[,7:ncol(mod_pol)]
+clim <- climate[,4:ncol(climate)]
+
+clim.ratio <- rep(NA, ncol(clim))
+
+for(i in 1:ncol(clim)){
+  test.struct <- rda(pol ~ clim[,i])
+  clim.ratio[i] <- test.struct$CCA$tot.chi / test.struct$CA$tot.chi
+}
+
+names(clim.ratio) <- colnames(clim)
+
+```
+
+Key variables based on the $latex \frac {\lambda_1} {\lambda_2} $ using pCCA appear to be climate variables relating to mean monthly percent summer sunshine.
+These variables are likely to be variables that are combinations of temperature and preciptiation variables, however this value is not linear, nor is it continuous, meaning that .
+While Telford & Birks (2009) have criticised sunshine reconstructions (Freshette et al., 2008) in the past (and perhaps with good reason), the strength of the sunshine signal in the climate data is worth noting, but the limitations of CCA may impact our ability to detect and accurately assess its importance.
+
 Model outputs
 =========================
+
+Once the climate variables of interest were established we began by bootstrapping the models for these climate variables.
+
 ```{r LoadOutputs, echo=FALSE, message=FALSE, warning=FALSE}

### 2013-10-02 R/bootstrap_mat.R

@@ -89,7 +89,7 @@ for(i in i:length(vals)){
       mat.res$variance[j, i]  <- mean((mean(prediction, na.rm=TRUE) - prediction)^2, na.rm=TRUE)
     }
     
-    cat('\n', j)  
+#    cat('\n', j)  
     
   }
     

### 2013-10-02 Paper_Markdown.Rmd
Comments & modifications from the CCA analysis of the climate variables structring pollen data.

@@
 ```{r echo=FALSE, message=FALSE, warning=FALSE}
+
+#  Sakari, can you take a look at this and let me know what you think?
 pol <- mod_pol[,7:ncol(mod_pol)]
@@
 names(clim.ratio) <- colnames(clim)
+best.clim <- clim.ratio[order(clim.ratio, decreasing = TRUE)]
 

### 2013-10-03 Paper_Markdown.Rmd
Some major changes to the coding, mostly adding echo, message and warning blocks to the Markdown R code sections.  Some revisions on the figures.  New table & text to accompany the CCA variance analysis.
Most importantly, I re-ran the MAT runs with fewer values.  It looks like the small intervals we use aren't really all that useful from 0 - 0.2, so I ditched them (hence the change to `bootstrap_mat.R`).  They're not useful for WAPLS or WA but I haven't implemented that yet.  That's why the Figure right before the Discussion has a weird dark/light pattern for WA and WAPLS.  New files come fro

@@
 
-```{r}
+```{r LoadData, echo=FALSE, warning=FALSE, message=FALSE}
 library(gridExtra)
@@
 
-```{r SiteLocations, fig.width=7, fig.height=6, fig.cap='Figure: North American site locations'}
+```{r SiteLocations, fig.width=7, fig.height=6, fig.cap='Figure: North American site locations', echo=FALSE, message=FALSE, warning=FALSE}
 map <- map_data('world')
@@
 
-```{r MinMaxDissim, fig.width=7, fig.height=6, fig.cap='Figure: North American site locations'}
+```{r MinMaxDissim, fig.width=7, fig.height=6, fig.cap='Figure: North American site locations', echo=FALSE, message=FALSE, warning=FALSE}
 
@@
 ```
-**Figure X**.
+**Figure 2**.
 *Distribution of squared-chord dissimilarity values.
@@
 
-```{r SiteMDS, fig.width=7, fig.height=6, fig.cap='Figure: North American site MDS scores'}
+```{r SiteMDS, fig.width=7, fig.height=6, fig.cap='Figure: North American site MDS scores', echo=FALSE, message=FALSE, warning=FALSE}
 
@@
 ```
-**Figure 2**.
+**Figure 3**.
 *Location of non-analogue points in the modern pollen data, presented within a non-metric multidimensional scaling ordination (Panel 1) and geographically.
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE}
+```{r message=FALSE, warning=FALSE}
 
-#  Sakari, can you take a look at this and let me know what you think?
+#  Sakari, can you take a look at this and let me know what you think?  Am I doing this right, and what do you make of these results?  Are we simply using too large a dataset?
+
 pol <- mod_pol[,7:ncol(mod_pol)]
@@
 names(clim.ratio) <- colnames(clim)
-best.clim <- clim.ratio[order(clim.ratio, decreasing = TRUE)]
+best.clim <- round(clim.ratio[order(clim.ratio, decreasing = TRUE)]*100, 0)
 
@@
 
-Key variables based on the $latex \frac {\lambda_1} {\lambda_2} $ using pCCA appear to be climate variables relating to mean monthly percent summer sunshine.
+Variable | Percent Variability
+-------- | -------------------
+`r names(best.clim)[1]` | `r best.clim[1]`
+`r names(best.clim)[2]` | `r best.clim[2]`
+`r names(best.clim)[3]` | `r best.clim[3]`
+`r names(best.clim)[4]` | `r best.clim[4]`
+`r names(best.clim)[5]` | `r best.clim[5]`
+`r names(best.clim)[6]` | `r best.clim[6]`
+`r names(best.clim)[7]` | `r best.clim[7]`
+
+Key variables based on the $latex \frac {\lambda_1} {\lambda_2} $ using pCCA appear to be climate variables relating to mean monthly percent summer sunshine (variables preceded by an 's').
 These variables are likely to be variables that are combinations of temperature and preciptiation variables, however this value is not linear, nor is it continuous, meaning that .
@@
 Once the climate variables of interest were established we began by bootstrapping the models for these climate variables.
+From the total pool of samples we generate a calibration dataset by sampling with replacement.
 
@@
 
-na.samples <- data.frame(site = rep(climate[,1], ncol(wa.res$bias)),
-                         size = as.vector(wa.res$sample_size),
-                         distance = rep(vals, each = nrow(wa.res$sample_size)),
-                         alpha.val = rep(diff(c(-.5,vals)), each = nrow(wa.res$sample_size))+0.005)
+na.samples <- data.frame(site = rep(climate[,1], ncol(mat.res$bias)),
+                         size = as.vector(mat.res$sample_size),
+                         distance = rep(seq(0, 1, by=0.01), each = nrow(mat.res$sample_size)))
 
@@
 
-ggplot(na.samples, aes(x=distance, y=size)) + 
-  geom_path(aes(group=site), alpha=0.001) +
+loss <- ggplot(na.samples, aes(x=distance, y=size)) + 
+  geom_path(aes(group=site), alpha=0.01) +
   xlab('Analogue Distance') + ylab('Available Samples') +
-  theme(text = element_text(size=24, family='serif')) +
-  theme_bw()
+  theme_bw() +
+  theme(text = element_text(size=24, family='serif'),
+        axis.text = element_text(size=14, family='serif')) +
+  scale_x_continuous(expand = c(0,0)) +
+  scale_y_continuous(expand=c(0,0)) +
+  geom_segment(aes(x=0.48, xend = 0.48, y = min(mat.res$sample_size, na.rm=TRUE), yend = nrow(climate)), size=1, color='red', linetype=2)
+
+loc.loss <- data.frame(lat =  climate$LATDD,
+                       long = climate$LONDD,
+                       loss = 1 - mat.res$sample_size[,49] / nrow(climate))
+
+loss.map <-  ggplot(data = data.frame(map), aes(long, lat)) + 
+  geom_path(aes(group=group), color='black') +
+  geom_point(data = loc.loss, 
+             aes(x = long, y = lat, color = loss), alpha=0.5) +
+   theme_bw() +
+  theme(text = element_text(size=24, family='serif'),
+        axis.text = element_text(size=14, family='serif')) +
+  coord_map(projection='albers', lat0=30, lat1=50, xlim=c(-140, -60), ylim=c(26, 75)) +
+  scale_color_continuous(low = '#F08080', high='#4169E1') +
+  xlab('') + ylab('')
+
+grid.arrange(loss, loss.map, ncol=2)
 
@@
 
-load('data/mat.res.RData')
-load('data/wa.res.RData')
-load('data/wapls.res.RData')
+mat.df <- data.frame(site = rep(climate[,1], ncol(mat.res$bias)),
+                     preds = as.vector(mat.res$mean_prediction),
+                     dissim = rep(seq(0, 1, by=0.01),each = nrow(clim)),
+                     model = 'MAT', 
+                     init = rep(climate[,10], ncol(mat.res$bias)))
 
-model.out <- data.frame(site   = rep(rep(climate[,1], ncol(mat.res$bias)),3),
-                        preds  = c(as.vector(mat.res$mean_prediction),
-                                   as.vector(wa.res$mean_prediction),
+model.out <- data.frame(site   = rep(rep(climate[,1], ncol(wa.res$bias)),2),
+                        preds  = c(as.vector(wa.res$mean_prediction),
                                    as.vector(wapls.res$mean_prediction)),
-                        dissim = rep(rep(vals, each=nrow(mat.res$bias)),3),
-                        model  = rep(c('MAT', 'WA', 'WAPLS'), each = length(mat.res$bias)))
+                        dissim = rep(rep(vals, each=nrow(wa.res$bias)),2),
+                        model  = rep(c('WA', 'WAPLS'), each = length(wa.res$bias)),
+                        init = rep(climate[,10], ncol(wa.res$bias)))
 
+#  Temporary fix until we re-run wa.re and wapls.res
+model.out <- rbind(mat.df, model.out)
 model.out <- model.out[!colSums(is.na(model.out)) > 0,]
 
-ggplot(model.out, aes(x = dissim, y = preds, group = site)) +
-  geom_path(alpha = 0.01) +
+#  This plot takes a really long time to plot.
+
+ggplot(model.out, aes(x = dissim, y = preds, group = site, color = init)) +
+  geom_path(alpha = 0.05) +
   theme_bw() +
   facet_wrap(~model, scales='free_y') +
-  scale_x_continuous(expand=c(0,0))
+  scale_x_continuous(expand=c(0,0)) +
+  scale_color_continuous(low='#F08080', high='#4169E1')
 

### 2013-10-03 R/bootstrap_mat.R

@@ -4,7 +4,7 @@ sfInit(parallel = TRUE, cpus = 3)
 
 # Find the quantiles & set up the analogue distance values:
 quantiles <- c(0, quantile(ecdf(diag.dist), c(seq(0, 0.2, by=0.001), seq(0.21, 1, by=0.01))))
-vals <- c(seq(0, 1, by=0.01))
+vals <- seq(0, 1, by=0.01)
 
 mat.res <- list(mean_prediction  = matrix(ncol=length(vals), nrow=nrow(new.pol)),
                 sample_size      = matrix(ncol=length(vals), nrow=nrow(new.pol)),

### 2013-11-15 R/BRT_time_test.r
Create BRT_time_test.r
A test version of BRT cross-validation code, to examine time requirements.

@@ -1,0 +1,51 @@
+# Test version of the BRT cross-validation code.
+#
+# This version does not save the results, but only tests the time requirement of the BRT
+# cross validation. 23 samples (i.e. 1% of total) are selected at random, and 50 bootstrap
+# iterations are performed for each sample. Predicted value and used time are printed
+# in each iteration.
+
+library (gbm)
+
+# Read the data from Excel. Note that the data must be organised as in NA_pollen_for_BRT.csv
+# (same column names and column order)
+
+cal_data <- sqrt(read.table("clipboard", header=TRUE))
+attach (cal_data)
+
+# Construct the formula string for gbm
+
+long_formula <- as.formula(paste("tjul ~ ", paste(paste("v", 1:56, sep = ""), collapse = " + "), sep = ""))
+
+# Select 23 random samples to test with
+
+test_samples <- sample(x=2289, size=23, replace=FALSE)
+
+# Run 50 bootstrap cycles with 23 random samples. The predicted values are not stored: 
+# they are over-written in each iteration in variable "pred".
+
+start_time <- Sys.time()
+pred <- numeric()
+
+for (i in 1:23) {
+
+   for (j in 1:50) {
+   
+      iter_start_time <- Sys.time()
+      calibration_set <- cal_data[sample(x=2289, size=2289, replace=TRUE),]
+      attach (calibration_set, warn.conflicts = FALSE)
+      xval_model <- gbm(long_formula , distribution="gaussian", n.trees=500, shrinkage=0.005, interaction.depth=6, cv.folds = 4, verbose=FALSE)  # Constructs the BRT model
+      xval_best_iter <- gbm.perf(xval_model,method="cv",plot.it=FALSE)  # Returns the number of trees that gives the best performance
+      test_set <- cal_data[test_samples[i],4:59]
+      attach (test_set, warn.conflicts = FALSE)
+      pred <- predict(xval_model,test_set,xval_best_iter, type="response") # Predicts for the test-set sample
+	  iter_end_time <- Sys.time()
+	  print(paste("Sample ",test_samples[i]," (",i,"/23), bootstrap cycle ",j,"/50, predicted value ",pred,".", sep=""))
+	  print (iter_end_time - iter_start_time)
+	  
+   }
+   
+}
+
+end_time <- Sys.time()
+print (end_time-start_time)

### 2013-11-15 R/BRT_time_test.r
Test version of the BRT cross-validation code, just to see the time requirements

@@ -39,9 +39,9 @@ for (i in 1:23) {
       test_set <- cal_data[test_samples[i],4:59]
       attach (test_set, warn.conflicts = FALSE)
       pred <- predict(xval_model,test_set,xval_best_iter, type="response") # Predicts for the test-set sample
-	  iter_end_time <- Sys.time()
-	  print(paste("Sample ",test_samples[i]," (",i,"/23), bootstrap cycle ",j,"/50, predicted value ",pred,".", sep=""))
-	  print (iter_end_time - iter_start_time)
+      iter_end_time <- Sys.time()
+      print(paste("Sample ",test_samples[i]," (",i,"/23), bootstrap cycle ",j,"/50, predicted value ",pred,".", sep=""))
+      print (iter_end_time - iter_start_time)
 	  
    }
    

### 2013-11-16 Paper_Markdown.Rmd
Added text to the document, cleaned up some extra lines in the bootstrap_mat.R, no actual code changes.

@@
 **Abstract**
-Some jibberish.
+Fossil pollen provides a widespread proxy for past vegetation.
+It is increasingly used in paleoclimatic reconstruction, but the limits of its utility are not well known.
+Methodological advances, such as the development of tests for spatial autocorrelation, and newer methods for climate reconstruction using boosted regression trees may improve the abilities of these climate reconstruction techniques, but little is known about the accuracy of these models under conditions of non-analogue vegetation known to occur at the end of the late glacial.
+In this paper we generate non-analogue pollen assemblages by excluding close neighbours from calibration datasets at increasingly larger squared chord distances.
+Using three standard and two newer techniques we then test the ability of pollen-based climate models to accurately recostruct climate under conditions of no-analogue vegetation.
 
@@
 A particular challenge for climate reconstruction using these much earlier samples is the presence of non-analogue vegetation.
-Non-analogue pollen assemblages pose a problem for many reconstruction techniques, these vegetation assemblages may be the result of sets of climate variables that no longer co-exist, ecological conditions resulting from differing rates of species migration or release from herbivores (Gill et al., 2009?), or as a result of changes in human land use (St.
+Non-analogue pollen assemblages pose a problem for many reconstruction techniques, these vegetation assemblages may be the result of sets of climate variables that no longer co-exist, ecological conditions resulting from differing rates of species migration or release from herbivores (Gill et al., 2009), or as a result of changes in human land use (St.
 Jacques?, maybe not the best.
@@
 dissim.scores <- data.frame(dissim = c(min.diss,
-                                       max.diss, 
                                        as.vector(dists)[sample(length(dists), 10000)]),
                             class   = c(rep('Minimum', length(min.diss)),
-                                        rep('Maximim', length(max.diss)), 
                                         rep('All', 10000)))
@@
         axis.text = element_text(size = 12, family='serif')) +
-  annotate("segment", x=0.48, xend = -.32, y = 0, yend = 1, size=3, color='red') +
+  annotate("segment", x=0.48, 
+           xend = -.32, y = 0, yend = 1, size=3, color='red') +
   scale_x_log10() + scale_y_sqrt()
@@
    theme_bw() +
-  coord_map(projection='albers', lat0=30, lat1=50, xlim=c(-140, -60), ylim=c(26, 75))
+  coord_map(projection='albers', lat0=30, lat1=50, xlim=c(-140, -60), ylim=c(26, 75)) +
+  xlab('Longitude') + ylab('Latitude') +
+  theme(text = element_text(size=24, family='serif'))
 
@@
 load('data/wapls.res.RData')
+load('data/rfor.res.RData')
+load('data/brt.res.RData')
 
-vals <- c(seq(0, 0.2, by=0.001), seq(0.21, 1, by=0.01))
+vals <- c(seq(0, 1, by=0.01))
 

### 2013-11-16 R/bootstrap_mat.R

@@ -77,7 +77,6 @@ for(i in i:length(vals)){
   }
   
   sfExport(list = list('keep.pol'))
-  
 
   for(j in 1:nrow(new.pol)){
 
@@ -97,4 +96,3 @@ for(i in i:length(vals)){
   save(mat.res, file = 'data/mat.res.RData')
   cat(i)
 }
-

### 2013-11-16 Paper_Markdown.Rmd
Added figure names to the Rmd chunks and started working on the plots.  Moved loading the RData files into data_setup.R.

@@
 
-```{r SiteLocations, fig.width=7, fig.height=6, fig.cap='Figure: North American site locations', echo=FALSE, message=FALSE, warning=FALSE}
+```{r Fig1SiteLocations, fig.width=7, fig.height=6, fig.cap='Figure: North American site locations', echo=FALSE, message=FALSE, warning=FALSE}
 map <- map_data('world')
@@
 
-```{r MinMaxDissim, fig.width=7, fig.height=6, fig.cap='Figure: North American site locations', echo=FALSE, message=FALSE, warning=FALSE}
+```{r Fig2MinDissim, fig.width=7, fig.height=6, fig.cap='Figure: North American site locations', echo=FALSE, message=FALSE, warning=FALSE}
 
@@
 
-```{r SiteMDS, fig.width=7, fig.height=6, fig.cap='Figure: North American site MDS scores', echo=FALSE, message=FALSE, warning=FALSE}
+```{r Fig3SiteMDS, fig.width=7, fig.height=6, fig.cap='Figure: North American site MDS scores', echo=FALSE, message=FALSE, warning=FALSE}
 
@@
 
-```{r LoadOutputs, echo=FALSE, message=FALSE, warning=FALSE}
-load('data/wa.res.RData')
-load('data/mat.res.RData')
-load('data/wapls.res.RData')
-load('data/rfor.res.RData')
-load('data/brt.res.RData')
+```{r Fig4SampleLoss, echo=FALSE, message=FALSE, warning=FALSE}
 

### 2013-11-17 R/bootstrap_wapls.R
The variance term for wapls was being entered as `var` rather than `variance`.  Fixed bootstrap_wapls to reflect the change in the variable name & re-saved the wapls.res.RData file.

@@ -81,7 +81,7 @@ for(i in 1:length(vals)){
       wapls.res$sample_size[j,i] <- sum(keep.pol[j,], na.rm=TRUE)
       wapls.res$bias[j, i] <- (climate[j,10] - mean(prediction, na.rm=TRUE))^2
       wapls.res$exp[j, i]  <- mean((climate[j,10] - prediction)^2)
-      wapls.res$var[j, i]  <- mean((mean(prediction) - prediction)^2)
+      wapls.res$variance[j, i]  <- mean((mean(prediction) - prediction)^2)
     }
     
   }

### 2013-11-17 Paper_Markdown.Rmd
Added the figures for Bias and Variance.  Need to look into them a bit more.

@@
 ```
-**Figure X**.
+**Figure 4**.
 *As the analogue distance exclusion is increased the available data set size is reduced, although some points continue to maintain a large number of potential analogues, this is directly related to the minimum distance to the closest analogue.
@@
 
-```{r PredictedValues, message=FALSE, echo=FALSE, warning=FALSE}
+```{r Fig5PredictedValues, message=FALSE, echo=FALSE, warning=FALSE}
 
-mat.df <- data.frame(site = rep(climate[,1], ncol(mat.res$bias)),
-                     preds = as.vector(mat.res$mean_prediction),
-                     dissim = rep(seq(0, 1, by=0.01),each = nrow(clim)),
-                     model = 'MAT', 
-                     init = rep(climate[,10], ncol(mat.res$bias)))
+model.out <- data.frame(site   = rep(rep(climate[,1], ncol(wa.res$bias)),5),
+                        preds  = c(as.vector(mat.res$mean_prediction),
+                                   as.vector(wa.res$mean_prediction),
+                                   as.vector(wapls.res$mean_prediction),
+                                   as.vector(rfor.res$mean_prediction),
+                                   as.vector(brt.res$mean_prediction)),
+                        bias   = c(as.vector(mat.res$bias),
+                                   as.vector(wa.res$bias),
+                                   as.vector(wapls.res$bias),
+                                   as.vector(rfor.res$bias),
+                                   as.vector(brt.res$bias)),
+                        variance = c(as.vector(mat.res$variance),
+                                   as.vector(wa.res$variance),
+                                   as.vector(wapls.res$variance),
+                                   as.vector(rfor.res$variance),
+                                   as.vector(brt.res$variance)),
+                        dissim = rep(rep(vals, each=nrow(wa.res$bias)),5),
+                        model  = rep(c('MAT', 'WA', 'WAPLS', 'rFor', 'BRT'), each = length(wa.res$bias)),
+                        init = rep(climate[,10], ncol(wa.res$bias), 5))
 
-model.out <- data.frame(site   = rep(rep(climate[,1], ncol(wa.res$bias)),2),
-                        preds  = c(as.vector(wa.res$mean_prediction),
-                                   as.vector(wapls.res$mean_prediction)),
-                        dissim = rep(rep(vals, each=nrow(wa.res$bias)),2),
-                        model  = rep(c('WA', 'WAPLS'), each = length(wa.res$bias)),
-                        init = rep(climate[,10], ncol(wa.res$bias)))
+#  This kills ~ 300,000 rows right now because of gaps in rfor.res and brt.res.
+model.out <- model.out[!rowSums(is.na(model.out)) > 0,]
+#  This is a temporary fix, 
+samp.set <- sapply(c('MAT', 'WA', 'WAPLS', 'rFor', 'BRT'), function(x){
+  set.len <- sum(model.out$model == x)
+  if( set.len > 5000) set.len <- 5000
+  sample(which(model.out$model == x), set.len)
+})
 
-#  Temporary fix until we re-run wa.re and wapls.res
-model.out <- rbind(mat.df, model.out)
-model.out <- model.out[!colSums(is.na(model.out)) > 0,]
+model.samp <- model.out[unlist(samp.set),]
 
@@
 
-ggplot(model.out, aes(x = dissim, y = preds, group = site, color = init)) +
+ggplot(model.samp, aes(x = dissim, y = preds, group = site, color = init)) +
+  geom_path(alpha = 0.1) +
+  theme_bw() +
+  facet_wrap(~model, scales='free_y', nrow = 1) +
+  scale_x_continuous(expand=c(0,0)) +
+  scale_color_continuous(low='#F08080', high='#4169E1') +
+  theme(legend.position = 'top',
+        axis.title = element_text(family = 'serif', face = 'bold', size = 14),
+        strip.text = element_text(family = 'serif', face = 'bold', size = 14),
+        axis.text = element_text(family = 'serif', size = 10),
+        strip.text = element_text(family = 'serif', face = 'bold', size = 12)) +
+  xlab('Calibration Dissimilarity') +
+  ylab('Predicted Value')
+
+```
+**Figure 5**.
+*Change in site predictions over time.
+Of great interest is the difference between model outputs as the exclusion zone increases.
+MAT can never predict beyond the extent of the data samples in the pool, and so as the exclusion zone increases the predictions move toward the dataset mean.
+WAPLS uses an initial transformation and then regresses against it to predict new values, so as the exclusion zone increases we see an increase in the spread of the predicted values, with much greater variability and a loss of accuracy.*
+
+```{r Fig6BiasValues, message=FALSE, echo=FALSE, warning=FALSE}
+
+ggplot(model.samp, aes(x = dissim, y = bias, group = site, color = init)) +
+  geom_path(alpha = 0.1) +
+  theme_bw() +
+  facet_wrap(~model, scales='free_y', nrow = 1) +
+  scale_x_continuous(expand=c(0,0)) +
+  scale_color_continuous(low='#F08080', high='#4169E1') +
+  theme(legend.position = 'top',
+        axis.title = element_text(family = 'serif', face = 'bold', size = 14),
+        strip.text = element_text(family = 'serif', face = 'bold', size = 14),
+        axis.text = element_text(family = 'serif', size = 10),
+        strip.text = element_text(family = 'serif', face = 'bold', size = 12)) +
+  xlab('Calibration Dissimilarity') +
+  ylab('Bootstrap Variance')
+
+```
+**Figure 6**. *Change in bias over time.*
+
+```{r Fig7VarianceValues, message=FALSE, echo=FALSE, warning=FALSE}
+
+ggplot(model.out, aes(x = dissim, y = variance, group = site, color = init)) +
   geom_path(alpha = 0.05) +
@@
   scale_x_continuous(expand=c(0,0)) +
-  scale_color_continuous(low='#F08080', high='#4169E1')
+  scale_color_continuous(low='#F08080', high='#4169E1') +
+  theme(legend.position = 'top',
+        axis.title = element_text(family = 'serif', face = 'bold', size = 14),
+        strip.text = element_text(family = 'serif', face = 'bold', size = 14),
+        axis.text = element_text(family = 'serif', size = 10),
+        strip.text = element_text(family = 'serif', face = 'bold', size = 12)) +
+  xlab('Calibration Dissimilarity') +
+  ylab('Bootstrap Variance')
 
 ```
-**Figure X**.
-*Change in site predictions over time.
-Of great interest is the difference between model outputs as the exclusion zone increases.
-MAT can never predict beyond the extent of the data samples in the pool, and so as the exclusion zone increases the predictions move toward the dataset mean.
-WAPLS uses an initial transformation and then regresses against it to predict new values, so as the exclusion zone increases we see an increase in the spread of the predicted values, with much greater variability and a loss of accuracy.*
+**Figure 7**. *Change in site variance over time.*
 

### 2013-11-18 R/bootstrap_wa.R
Error in the calculations of bias and expectation, the climate variable should be July (column 10) not January (column 4).  Made the errors overly large.  Other variables were unaffected.

@@ -82,8 +82,8 @@ for(i in i:length(vals)){
     
       wa.res$mean_prediction[j,i] <- mean(prediction, na.rm=TRUE)
       wa.res$sample_size[j,i] <- sum(keep.pol[j,], na.rm=TRUE)
-      wa.res$bias[j, i] <- (climate[j,4] - mean(prediction, na.rm=TRUE))^2
-      wa.res$expectation[j, i]  <- mean((climate[j,4] - prediction)^2, na.rm=TRUE)
+      wa.res$bias[j, i] <- (climate[j,10] - mean(prediction, na.rm=TRUE))^2
+      wa.res$expectation[j, i]  <- mean((climate[j,10] - prediction)^2, na.rm=TRUE)
       wa.res$variance[j, i]  <- mean((mean(prediction, na.rm=TRUE) - prediction)^2, na.rm=TRUE)
     }
         
@@ -97,3 +97,5 @@ for(i in i:length(vals)){
   cat(i)  
   
 }
+
+sfStop()

### 2013-11-18 R/bootstrap_wa.R
Changed clim variables in bootstrap WA.

@@ -82,8 +82,8 @@ for(i in i:length(vals)){
     
       wa.res$mean_prediction[j,i] <- mean(prediction, na.rm=TRUE)
       wa.res$sample_size[j,i] <- sum(keep.pol[j,], na.rm=TRUE)
-      wa.res$bias[j, i] <- (climate[j,4] - mean(prediction, na.rm=TRUE))^2
-      wa.res$expectation[j, i]  <- mean((climate[j,4] - prediction)^2, na.rm=TRUE)
+      wa.res$bias[j, i] <- (climate[j,10] - mean(prediction, na.rm=TRUE))^2
+      wa.res$expectation[j, i]  <- mean((climate[j,10] - prediction)^2, na.rm=TRUE)
       wa.res$variance[j, i]  <- mean((mean(prediction, na.rm=TRUE) - prediction)^2, na.rm=TRUE)
     }
         

### 2013-11-18 R/bootstrap_wa.R
Updated wa.res to fix the bug in the code around the climate variables.

@@ -86,7 +86,7 @@ for(i in i:length(vals)){
       wa.res$expectation[j, i]  <- mean((climate[j,10] - prediction)^2, na.rm=TRUE)
       wa.res$variance[j, i]  <- mean((mean(prediction, na.rm=TRUE) - prediction)^2, na.rm=TRUE)
     }
-        
+     
   }
   if(ncol(wa.res$bias) > i){
     if(!sum(is.na(wa.res$bias[,i+1])) == 0){

### 2013-11-18 Paper_Markdown.Rmd
Changed the bias plot to fix y axis and sqrt the yaxis values.

@@
 
-ggplot(model.samp, aes(x = dissim, y = bias, group = site, color = init)) +
+ggplot(model.samp, aes(x = dissim, y = sqrt(bias), group = site, color = init)) +
   geom_path(alpha = 0.1) +
@@
   xlab('Calibration Dissimilarity') +
-  ylab('Bootstrap Variance')
+  ylab('Bias (RMSEP)')
 

### 2013-11-18 Paper_Markdown.Rmd
Changed variable names, I think this is in conflict with the GitHub version.

@@
 
-ggplot(model.samp, aes(x = dissim, y = sqrt(bias), group = site, color = init)) +
+ggplot(model.samp, aes(x = dissim, y = bias, group = site, color = init)) +
   geom_path(alpha = 0.1) +
@@
   scale_x_continuous(expand=c(0,0)) +
+  scale_y_continuout(expand = c(0,0)) +
   scale_color_continuous(low='#F08080', high='#4169E1') +
@@
   xlab('Calibration Dissimilarity') +
-  ylab('Bias (RMSEP)')
+  ylab('Estimate Bias')
 
@@
 
-ggplot(model.out, aes(x = dissim, y = variance, group = site, color = init)) +
+ggplot(model.samp, aes(x = dissim, y = sqrt(variance), group = site, color = init)) +
   geom_path(alpha = 0.05) +
   theme_bw() +
-  facet_wrap(~model, scales='free_y') +
+  facet_wrap(~model, nrow = 1) +
   scale_x_continuous(expand=c(0,0)) +

### 2013-11-18 Paper_Markdown.Rmd
Re-ran the markdown.  Updated the axes on the plots.

@@
 **Abstract**
-Fossil pollen provides a widespread proxy for past vegetation.
+*Fossil pollen provides a widespread proxy for past vegetation.
 It is increasingly used in paleoclimatic reconstruction, but the limits of its utility are not well known.
@@
 In this paper we generate non-analogue pollen assemblages by excluding close neighbours from calibration datasets at increasingly larger squared chord distances.
-Using three standard and two newer techniques we then test the ability of pollen-based climate models to accurately recostruct climate under conditions of no-analogue vegetation.
+Using three standard and two newer techniques we then test the ability of pollen-based climate models to accurately recostruct climate under conditions of no-analogue vegetation.*
 
@@
   scale_x_continuous(expand=c(0,0)) +
-  scale_y_continuout(expand = c(0,0)) +
+  scale_y_continuous(expand = c(0,0)) +
   scale_color_continuous(low='#F08080', high='#4169E1') +
@@
   xlab('Calibration Dissimilarity') +
-  ylab('Estimate Bias')
+  ylab('Bias (RMSEP)')
 

### 2013-12-16 Paper_Markdown.Rmd
Re-ran the markdown file & updating wapls.res.RData
Re-compiled the markdown file but noticed an error with the wapls.res
files.  I've fixed that up (hence edits to bootstrap_wapls.R)  and have
a partially complete wapls.res file as well.  Approx 60% done.  It seems
a lot slower than usual though.
Removed free scales for the plotting (except this screws up wapls plots!), e-ran wa.res to try to fix bug.  Partial run results (up to 16 runs).

@@
 ```{r LoadData, echo=FALSE, warning=FALSE, message=FALSE}
-library(gridExtra)
 
@@
   theme_bw() +
-  facet_wrap(~model, scales='free_y', nrow = 1) +
+  facet_wrap(~model, nrow = 1) +
   scale_x_continuous(expand=c(0,0)) +

### 2013-12-17 Paper_Markdown.Rmd
Updated the WA results, I've standardized the plotting somewhat.  Not sure why the colors aren't working out as well as I'd like.

@@
 
-```{r message=FALSE, warning=FALSE}
+```{r echo = FALSE, message=FALSE, warning=FALSE}
 
@@
   xlab('Calibration Dissimilarity') +
-  ylab('Predicted Value')
+  ylab('Predicted Value') +
+  ylim(0, 40)
 
@@
   theme_bw() +
-  facet_wrap(~model, scales='free_y', nrow = 1) +
+  facet_wrap(~model, nrow = 1) +
   scale_x_continuous(expand=c(0,0)) +
-  scale_y_continuous(expand = c(0,0)) +
+  scale_y_sqrt(expand = c(0,0)) +
   scale_color_continuous(low='#F08080', high='#4169E1') +
@@
   scale_x_continuous(expand=c(0,0)) +
+  scale_y_sqrt(expand=c(0,0)) +
   scale_color_continuous(low='#F08080', high='#4169E1') +

### 2013-12-18 Paper_Markdown.Rmd
Local changes to plotting.

--- 2013-11-18 Paper_Markdown.Rmd
@@
 **Abstract**
-*Fossil pollen provides a widespread proxy for past vegetation.
+Fossil pollen provides a widespread proxy for past vegetation.
 It is increasingly used in paleoclimatic reconstruction, but the limits of its utility are not well known.
@@
 In this paper we generate non-analogue pollen assemblages by excluding close neighbours from calibration datasets at increasingly larger squared chord distances.
-Using three standard and two newer techniques we then test the ability of pollen-based climate models to accurately recostruct climate under conditions of no-analogue vegetation.*
+Using three standard and two newer techniques we then test the ability of pollen-based climate models to accurately recostruct climate under conditions of no-analogue vegetation.
 
@@
   theme_bw() +
-  facet_wrap(~model, scales='free_y', nrow = 1) +
-  scale_x_continuous(expand=c(0,0)) +
+  facet_wrap(~model, nrow = 1) +
+  scale_x_continuous(expand=c(0,0), limits=c()) +
   scale_color_continuous(low='#F08080', high='#4169E1') +
@@
   theme_bw() +
-  facet_wrap(~model, nrow = 1) +
+  facet_wrap(~model, nrow = 1, scales='free_y') +
   scale_x_continuous(expand=c(0,0)) +

### 2013-12-30 Paper_Markdown.Rmd

--- 2013-12-17 Paper_Markdown.Rmd
@@
 ```{r Fig1SiteLocations, fig.width=7, fig.height=6, fig.cap='Figure: North American site locations', echo=FALSE, message=FALSE, warning=FALSE}
+
 map <- map_data('world')
@@
 
-```{r Fig4SampleLoss, echo=FALSE, message=FALSE, warning=FALSE}
+```{r Fig4SampleLoss, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3, fig.width=6}
 
@@
 
+init.color <- colorRampPalette(c('#F08080', '#68d64f', '#4169E1'))
+clim.colors <- seq(0, 31, length.out = 300)
+init.set <- init.color(300)[findInterval(x = climate[,10], vec=clim.colors)]
+
 model.out <- data.frame(site   = rep(rep(climate[,1], ncol(wa.res$bias)),5),
@@
                         model  = rep(c('MAT', 'WA', 'WAPLS', 'rFor', 'BRT'), each = length(wa.res$bias)),
-                        init = rep(climate[,10], ncol(wa.res$bias), 5))
+                        init = rep(rep(climate[,10], 101), 5))
 
@@
 model.out <- model.out[!rowSums(is.na(model.out)) > 0,]
+
 #  This is a temporary fix, 
-samp.set <- sapply(c('MAT', 'WA', 'WAPLS', 'rFor', 'BRT'), function(x){
+sample.model <- function(x){
+  #  How many samples exist for the model?
   set.len <- sum(model.out$model == x)
   if( set.len > 5000) set.len <- 5000
-  sample(which(model.out$model == x), set.len)
-})
+
+  set <- sample(which(model.out$model == x), set.len)
+  start.end <- which(model.out$model == x & 
+                       (model.out$dissim %in% c(0, 0.5, 1)))
+
+  unique(c(set, start.end))
+
+}
+
+samp.set <- sapply(c('MAT', 'WA', 'WAPLS', 'rFor', 'BRT'), sample.model)
 
@@
   facet_wrap(~model, nrow = 1) +
-  scale_x_continuous(expand=c(0,0)) +
-  scale_color_continuous(low='#F08080', high='#4169E1') +
-  theme(legend.position = 'top',
+  scale_x_continuous(expand=c(0,0), breaks = c(0.5, 1)) +
+  scale_y_continuous(expand=c(0,0), limits=c(0, 40)) +
+  scale_color_continuous(low = 'blue', high='pink', trans='sqrt') +
+  theme(legend.position = 'none',
         axis.title = element_text(family = 'serif', face = 'bold', size = 14),
@@
         strip.text = element_text(family = 'serif', face = 'bold', size = 12)) +
+  annotate("segment", x=0.48, 
+           xend = 0.48, y = 0, yend = 100, size=1, linetype='dashed') +
   xlab('Calibration Dissimilarity') +
-  ylab('Predicted Value') +
-  ylim(0, 40)
+  ylab('Predicted Value')
 
@@
   facet_wrap(~model, nrow = 1) +
-  scale_x_continuous(expand=c(0,0)) +
-  scale_y_sqrt(expand = c(0,0)) +
-  scale_color_continuous(low='#F08080', high='#4169E1') +
-  theme(legend.position = 'top',
+  scale_x_continuous(expand=c(0,0), breaks = c(0.5, 1)) +
+  scale_y_sqrt(expand=c(0,0)) +
+  scale_color_continuous(low = 'blue', high='pink', trans='sqrt') +
+  theme(legend.position = 'none',
         axis.title = element_text(family = 'serif', face = 'bold', size = 14),
@@
 ggplot(model.samp, aes(x = dissim, y = sqrt(variance), group = site, color = init)) +
-  geom_path(alpha = 0.05) +
+  geom_path(alpha = 0.1) +
   theme_bw() +
   facet_wrap(~model, nrow = 1) +
-  scale_x_continuous(expand=c(0,0)) +
+  scale_x_continuous(expand=c(0,0), breaks = c(0.5, 1)) +
   scale_y_sqrt(expand=c(0,0)) +
-  scale_color_continuous(low='#F08080', high='#4169E1') +
-  theme(legend.position = 'top',
+  scale_color_continuous(low = 'blue', high='pink', trans='sqrt') +
+  theme(legend.position = 'none',
         axis.title = element_text(family = 'serif', face = 'bold', size = 14),

### 2014-03-25 Paper_Markdown.Rmd
Started editing again.  main change was to create both parallelized and non-parallelized instances of the program.

@@
 ```{r Fig3SiteMDS, fig.width=7, fig.height=6, fig.cap='Figure: North American site MDS scores', echo=FALSE, message=FALSE, warning=FALSE}
-
+X
 mds <- data.frame(pol.mds$points,

### 2014-03-25 R/SourceRun.R

@@ -1,11 +1,13 @@
 #  Run all bootstraps:
 source('R/data_setup.R')
 
-source('R/bootstrap_mat.R')
+source('R/nonPar/bootstrap_mat.R')
 cat('Done MAT/n')
-source('R/bootstrap_wa.R')
+source('R/nonPar/bootstrap_wa.R')
 cat('Done WA/n')
-source('R/bootstrap_wapls.R')
+source('R/nonPar/bootstrap_wapls.R')
 cat('Done WAPLS/n')
-source('R/bootstrap_rfor.R')
+source('R/nonPar/bootstrap_rfor.R')
+cat('Done rFor/n')
+source('R/nonPar/bootstrap_brt.R')
 cat('Done rFor/n')

### 2014-03-25 R/bootstrapping.noaanalogs.R

@@ -52,6 +52,10 @@ rfor.res <- list(prediction     = matrix(ncol=length(vals), nrow=nrow(new.pol)),
                  r.squared      = matrix(ncol=length(vals), nrow=nrow(new.pol)),
                  closest.n      = matrix(ncol=length(vals), nrow=nrow(new.pol)))
 
+brt.res <- list(prediction     = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                r.squared      = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                closest.n      = matrix(ncol=length(vals), nrow=nrow(new.pol)))
+
 
 fmat <- function(x){
   #  returns the means of the 1:10 closest:
@@ -131,6 +135,17 @@ for(i in 1:length(vals)){
       
     }
             
+    if(is.na(brt.res$prediction[j,i])){
+      rfor <- try(randomForest(x=new.pol[keep.pol[j,],!zeros], y = climate[keep.pol[j,],4]))
+      
+      if(length(rfor) > 1){
+        pred.brt <- predict(rfor, newdata = new.pol[j,!zeros])
+        brt.res$prediction[j,i] <- pred.rfor
+        brt.res$r.squared[j,i]  <- mean(rfor$rsq)
+      }
+      
+    }
+    
     if(j %% 120 == 0) cat('.')
   }
   cat('\n', i)

### 2014-03-25 R/data_setup.R

@@ -1,6 +1,5 @@
 library(gridExtra)
 library(ggplot2)
-library(analogue)
 library(MASS)
 library(plyr)
 library(rioja)

### 2014-03-25 R/nonPar/bootstrap_brt.R

@@ -1,0 +1,132 @@
+# This is a modification of bootstrap_rfor.R, with the following changes:
+# - instead of "randomForest", loads (and passes to snowfall)
+#   the "gbm" library
+# - all references to "rfor.res" replaced with "brt.res"
+# - modified subset.pol() to return the climate values as a single-column 
+#   data frame (and not as vector)
+# - wrote function brt.run(), which replaced rfor.run()
+
+library(gbm)
+
+vals <- seq(0, 1, by=0.01)
+
+if('brt.res.RData' %in% list.files('data')){
+  load('data/brt.res.RData')
+}
+if(!'brt.res.RData' %in% list.files('data')){
+  brt.res <- list(mean_prediction = matrix(ncol=length(vals), nrow=nrow(new.pol)),   
+                   sample_size = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                   bias = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                   variance = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                   expectation = matrix(ncol=length(vals), nrow=nrow(new.pol)))
+}
+
+i <- 1; j <- 1
+
+subset.pol <- function(set){
+  
+  calib.samples <- sample((1:length(set))[set], replace=TRUE)
+  
+  zeros <- colSums(new.pol[calib.samples,], na.rm = TRUE) == 0
+
+  # Modified the following to return the climate values as a 
+  # single-column data frame.
+  
+  list(calib.pol = new.pol[calib.samples,!zeros],
+       calib.clim = climate[calib.samples,10,drop=FALSE], 
+       zeros = zeros)   
+
+}
+
+brt.run <- function(j){
+  
+  set <- keep.pol[j,]
+  
+  x <- subset.pol(set)
+  
+  zeros <- colSums(x$calib.pol, na.rm = TRUE) == 0
+
+  # Construct the brt formula based on x$calib.clim and x$calib.pol column 
+  # names, omitting the zero-sum taxa which won't be used.
+  
+  brt.formula <- as.formula(paste (colnames(x$calib.clim)[1], "~", paste(colnames(x$calib.pol)[!zeros], collapse=" + ")))     
+
+  # Construct brt model. x$calib.clim and x$calib.pol combined into a single
+  # data frame and passed as the "data" parameter.
+  
[…]

### 2014-03-25 R/nonPar/bootstrap_mat.R

@@ -1,0 +1,85 @@
+#  Set up the analogue distance exclusion values.
+#  This was originally a set of quantiles, but it played havoc on plotting and
+#  didn't seem to add much to the understanding of what was going on.
+vals <- seq(0, 1, by=0.01)
+
+mat.res <- list(mean_prediction  = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                sample_size      = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                bias             = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                variance         = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                expectation      = matrix(ncol=length(vals), nrow=nrow(new.pol)))
+
+#  Turn the distance table into a matrix.
+diag.dist <- as.matrix(dists); diag(diag.dist) <- NA
+
+i <- 1; j <- 1
+
+#  For MAT to work we don't need to re-calculate the distace matrix, we just need to know the
+#  sample of interest and the subset:
+
+fmat <- function(x){
+  #  returns the means of the 1:10 closest to calculate the RMSE for analogue selection:
+  
+  index.used <- order(diag.dist[x,])[keep.pol[x,]]
+  ten.or.less <- min(c(10, sum(keep.pol[x,])))
+  
+  vals <- cumsum(climate[index.used,4][1:ten.or.less]) / 1:ten.or.less
+  vars <- c(0, lapply(1:ten.or.less, function(x)var(climate[index.used,4][1:x])))
+  
+  if(length(vals) < 10) vals <- c(vals, rep(NA, 10-length(vals)))
+  
+  vals
+  
+}
+
+mat.fun <- function(z, j, min){
+  #  For sample j, pass in a vector of samples to be used for calibration.
+  
+  #  These are the samples that can be used:
+  calib.samples <- (1:nrow(new.pol))[ keep.pol[j,] ]
+  
+  #  now we resample:
+  samples <- sample(calib.samples, replace=TRUE)
+  
+  distances <- order(diag.dist[j,samples])
+  
+  #  This gives the closest 10 climate variables to the sample of interest:
+  clim.var <- climate[samples, 10][distances[1:10]]
+  
+  #  And the means of the 1 - 10 closest -
[…]

### 2014-03-25 R/nonPar/bootstrap_rfor.R

@@ -1,0 +1,125 @@
+library(snowfall)
+
+library(randomForest)
+
+sfStop()
+sfInit(parallel = TRUE, cpus = 6)
+
+vals <- seq(0, 1, by=0.01)
+
+if('rfor.res.RData' %in% list.files('data')){
+  load('data/rfor.res.RData')
+}
+if(!'rfor.res.RData' %in% list.files('data')){
+  rfor.res <- list(mean_prediction  = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                   sample_size      = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                   bias             = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                   variance         = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                   expectation      = matrix(ncol=length(vals), nrow=nrow(new.pol)))
+}
+
+i <- 1; j <- 1
+
+subset.pol <- function(set){
+  
+  calib.samples <- sample((1:length(set))[set], replace=TRUE)
+  
+  zeros <- colSums(new.pol[calib.samples,], na.rm = TRUE) == 0
+  
+  list(calib.pol  = new.pol[calib.samples,!zeros],
+       calib.clim = climate[calib.samples,10],
+       zeros = zeros)
+
+}
+
+rfor.run <- function(j){
+  
+  set <- keep.pol[j,]
+  
+  x <- subset.pol(set)
+  
+  zeros <- colSums(x$calib.pol, na.rm = TRUE) == 0
+  
+  rfor <- try(randomForest(x=x$calib.pol[,!zeros], 
+                     y = x$calib.clim))
+  
+  if (length(rfor) > 1){
+    output <- predict(rfor, newdata = new.pol[j,!zeros])
+    }
+  else{
+    output <- NA
+  }
+  return(output)
+  
+}
+
+#  Parallelize:
+sfExport(list = list('rfor.run'))
+sfExport(list = list('new.pol'))
+sfExport(list = list('subset.pol'))
+sfExport(list = list('climate'))
+sfLibrary(randomForest)
+
+#  Create a set of all possible xy pairs in the rfor tables, and then look to see if
+#  they've been sampled yet.
+longlist <- expand.grid(row = 1:nrow(new.pol), col=1:length(vals))
+longlist <- data.frame(longlist, calc = is.na(as.vector(rfor.res$bias)))
+
+#  Now to sample in an efficient way:  This expanded grid gives the probability
[…]

### 2014-03-25 R/nonPar/bootstrap_wa.R

@@ -1,0 +1,101 @@
+library(snowfall)
+
+sfStop()
+sfInit(parallel = TRUE, cpus = 6)
+
+# Set up the analogue distance exclusion values.
+#  This was originally a set of quantiles, but it played havoc on plotting and
+#  didn't seem to add much to the understanding of what was going on.
+vals <- seq(0, 1, by=0.01)
+
+wa.res <- list(mean_prediction  = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+               sample_size      = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+               bias             = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+               variance         = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+               expectation      = matrix(ncol=length(vals), nrow=nrow(new.pol)))
+
+if('wa.res.RData' %in% list.files('data/')) {
+  load('data/wa.res.RData')
+}
+
+i <- 1; j <- 1
+
+subset.pol <- function(set){
+  
+  calib.samples <- sample((1:length(set))[set], replace=TRUE)
+  
+  zeros <- colSums(new.pol[calib.samples,], na.rm = TRUE) == 0
+  
+  list(calib.pol  = new.pol[calib.samples,!zeros],
+       calib.clim = climate[calib.samples,10],
+       zeros = zeros)
+
+}
+
+wa.run <- function(j){
+  
+  set <- keep.pol[j,]
+  
+  x <- subset.pol(set)
+  
+  zeros <- colSums(x$calib.pol, na.rm = TRUE) == 0
+  
+  pred.wa <- try(predict(wa(x=x$calib.pol, 
+                            env = x$calib.clim, 
+                            deshrinking='monotonic'), 
+                         newdata = new.pol[j,!x$zeros]))
+  
+  if(length(pred.wa) > 1){
+    
+    output <- pred.wa$pred$pred
+    
+  }
+  else{
+    output <- NA
+  }
+  return(output)
+  
+}
+
+#  Parallelize:
+sfExport(list = list('wa.run'))
+sfExport(list = list('new.pol'))
+sfExport(list = list('subset.pol'))
+sfExport(list = list('climate'))
+sfLibrary(analogue)
+
+
+for(i in i:length(vals)){
[…]

### 2014-03-25 R/nonPar/bootstrap_wapls.R

@@ -1,0 +1,93 @@
+library(snowfall)
+
+sfStop()
+sfInit(parallel = TRUE, cpus = 6)
+
+# Set up the analogue distance exclusion values.
+#  This was originally a set of quantiles, but it played havoc on plotting and
+#  didn't seem to add much to the understanding of what was going on.
+vals <- c(seq(0, 1, by=0.01))
+
+if('wapls.res.RData' %in% list.files('data')){
+  load('data/wapls.res.RData')
+}
+if(!'wapls.res.RData' %in% list.files('data')){
+  wapls.res <- list(mean_prediction  = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                    sample_size      = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                    bias             = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                    variance         = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                    expectation      = matrix(ncol=length(vals), nrow=nrow(new.pol)))
+}
+
+i <- 1; j <- 1
+
+subset.pol <- function(set){
+  
+  calib.samples <- sample((1:length(set))[set], replace=TRUE)
+  
+  zeros <- colSums(new.pol[calib.samples,], na.rm = TRUE) == 0
+  
+  list(calib.pol  = new.pol[calib.samples,!zeros],
+       calib.clim = climate[calib.samples,10],
+       zeros = zeros)
+
+}
+
+wapls.run <- function(j){
+  
+  set <- keep.pol[j,]
+  
+  x <- subset.pol(set)
+  
+  zeros <- colSums(x$calib.pol, na.rm = TRUE) == 0
+  
+  wapls <- try(WAPLS(y=x$calib.pol[,!zeros], 
+                     x = x$calib.clim))
+  
+  if (length(wapls) > 1){
+    pred.wapls <- predict(wapls, newdata = new.pol[j,!zeros])
+    output    <- pred.wapls$fit[which.min(performance(wapls)$object[,1])]
+    }
+  else{
+    output <- NA
+  }
+  return(output)
+  
+}
+
+#  Parallelize:
+sfExport(list = list('wapls.run'))
+sfExport(list = list('new.pol'))
+sfExport(list = list('subset.pol'))
+sfExport(list = list('climate'))
+sfLibrary(rioja)
+
+
+for(i in 1:length(vals)){
[…]

### 2014-03-25 R/SourceRun.R
Finished editing all source files for non-parallel versions.

@@ -1,13 +1,22 @@
 #  Run all bootstraps:
 source('R/data_setup.R')
 
-source('R/nonPar/bootstrap_mat.R')
+
+par <- FALSE
+
+file.lead <- ifelse(par == TRUE, 'R/', 'R/nonPar/')
+
+source(paste(file.lead, 'bootstrap_mat.R', sep ='')
 cat('Done MAT/n')
-source('R/nonPar/bootstrap_wa.R')
+
+source(paste(file.lead, 'bootstrap_wa.R', sep ='')
 cat('Done WA/n')
-source('R/nonPar/bootstrap_wapls.R')
+
+source(paste(file.lead, 'bootstrap_wapls.R', sep ='')
 cat('Done WAPLS/n')
-source('R/nonPar/bootstrap_rfor.R')
+
+source(paste(file.lead, 'bootstrap_rfor.R', sep ='')
 cat('Done rFor/n')
-source('R/nonPar/bootstrap_brt.R')
+
+source(paste(file.lead, 'bootstrap_brt.R', sep ='')
 cat('Done rFor/n')

### 2014-03-25 R/nonPar/bootstrap_rfor.R

@@ -1,9 +1,4 @@
-library(snowfall)
-
 library(randomForest)
-
-sfStop()
-sfInit(parallel = TRUE, cpus = 6)
 
 vals <- seq(0, 1, by=0.01)
 
@@ -53,13 +48,6 @@ rfor.run <- function(j){
   
 }
 
-#  Parallelize:
-sfExport(list = list('rfor.run'))
-sfExport(list = list('new.pol'))
-sfExport(list = list('subset.pol'))
-sfExport(list = list('climate'))
-sfLibrary(randomForest)
-
 #  Create a set of all possible xy pairs in the rfor tables, and then look to see if
 #  they've been sampled yet.
 longlist <- expand.grid(row = 1:nrow(new.pol), col=1:length(vals))
@@ -103,7 +91,7 @@ for(k in samp){
     
   if(is.na(rfor.res$mean_prediction[j,i])){
     #  Run randomForest with raw defaults
-    prediction <- unlist(sfLapply(rep(j, 50), fun = rfor.run))
+    prediction <- unlist(lapply(rep(j, 50), fun = rfor.run))
   
     rfor.res$mean_prediction[j,i] <- mean(prediction, na.rm=TRUE)
     rfor.res$sample_size[j,i] <- sum(keep.pol[j,], na.rm=TRUE)

### 2014-03-25 R/nonPar/bootstrap_wa.R

@@ -1,8 +1,3 @@
-library(snowfall)
-
-sfStop()
-sfInit(parallel = TRUE, cpus = 6)
-
 # Set up the analogue distance exclusion values.
 #  This was originally a set of quantiles, but it played havoc on plotting and
 #  didn't seem to add much to the understanding of what was going on.
@@ -57,28 +52,19 @@ wa.run <- function(j){
   
 }
 
-#  Parallelize:
-sfExport(list = list('wa.run'))
-sfExport(list = list('new.pol'))
-sfExport(list = list('subset.pol'))
-sfExport(list = list('climate'))
-sfLibrary(analogue)
-
 
 for(i in i:length(vals)){
   #  This runs through each analogue distance
   keep.pol <- aaply(diag.dist, 1, 
                     function(x) {x > vals[i]})
 
-  sfExport(list = list('keep.pol'))
-  
   diag(keep.pol) <- FALSE
   
   for(j in 1:nrow(new.pol)){
     
     if(is.na(wa.res$mean_prediction[j,i])){
       #  Run WA with monotonic deshrinking.
-      prediction <- unlist(sfLapply(rep(j, 100), fun = wa.run))
+      prediction <- unlist(lapply(rep(j, 100), fun = wa.run))
     
       wa.res$mean_prediction[j,i] <- mean(prediction, na.rm=TRUE)
       wa.res$sample_size[j,i] <- sum(keep.pol[j,], na.rm=TRUE)
@@ -97,5 +83,3 @@ for(i in i:length(vals)){
   cat(i)  
   
 }
-
-sfStop()

### 2014-03-25 R/nonPar/bootstrap_wapls.R

@@ -1,8 +1,3 @@
-library(snowfall)
-
-sfStop()
-sfInit(parallel = TRUE, cpus = 6)
-
 # Set up the analogue distance exclusion values.
 #  This was originally a set of quantiles, but it played havoc on plotting and
 #  didn't seem to add much to the understanding of what was going on.
@@ -55,13 +50,7 @@ wapls.run <- function(j){
   
 }
 
-#  Parallelize:
-sfExport(list = list('wapls.run'))
-sfExport(list = list('new.pol'))
-sfExport(list = list('subset.pol'))
-sfExport(list = list('climate'))
 sfLibrary(rioja)
-
 
 for(i in 1:length(vals)){
   #  This runs through each analogue distance
@@ -69,13 +58,11 @@ for(i in 1:length(vals)){
                     function(x) {x > vals[i]})
   diag(keep.pol) <- FALSE
   
-  sfExport(list = list('keep.pol'))
-  
   for(j in 1:nrow(new.pol)){
     
     if(is.na(wapls.res$mean_prediction[j,i])){
       #  Run WA with monotonic deshrinking.
-      prediction <- unlist(sfLapply(rep(j, 100), fun = wapls.run))
+      prediction <- unlist(lapply(rep(j, 100), fun = wapls.run))
     
       wapls.res$mean_prediction[j,i] <- mean(prediction, na.rm=TRUE)
       wapls.res$sample_size[j,i] <- sum(keep.pol[j,], na.rm=TRUE)
@@ -89,5 +76,3 @@ for(i in 1:length(vals)){
   save(wapls.res, file = 'data/wapls.res.RData')
   cat(i)
 }
-
-  

### 2014-03-25 R/data_setup.R
Added string at the top that will install packages if they aren't there for use on the cloud.

@@ -1,3 +1,5 @@
+install.packages(c('gridExtra', 'ggplot2', 'MASS', 'plyr', 'rioja', 'randomForest', 'vegan', 'maps', 'maptools', 'ggmap'))
+
 library(gridExtra)
 library(ggplot2)
 library(MASS)

### 2014-03-25 R/SourceRun.R
Forgot to close brackes in SourceRun.  Whoops.

@@ -6,17 +6,17 @@ par <- FALSE
 
 file.lead <- ifelse(par == TRUE, 'R/', 'R/nonPar/')
 
-source(paste(file.lead, 'bootstrap_mat.R', sep ='')
+source(paste(file.lead, 'bootstrap_mat.R', sep =''))
 cat('Done MAT/n')
 
-source(paste(file.lead, 'bootstrap_wa.R', sep ='')
+source(paste(file.lead, 'bootstrap_wa.R', sep =''))
 cat('Done WA/n')
 
-source(paste(file.lead, 'bootstrap_wapls.R', sep ='')
+source(paste(file.lead, 'bootstrap_wapls.R', sep =''))
 cat('Done WAPLS/n')
 
-source(paste(file.lead, 'bootstrap_rfor.R', sep ='')
+source(paste(file.lead, 'bootstrap_rfor.R', sep =''))
 cat('Done rFor/n')
 
-source(paste(file.lead, 'bootstrap_brt.R', sep ='')
+source(paste(file.lead, 'bootstrap_brt.R', sep =''))
 cat('Done rFor/n')

### 2014-03-25 R/data_setup.R
Changed the install so that it only installs when the flag `install.it` is set to TRUE.

@@ -1,4 +1,6 @@
-install.packages(c('gridExtra', 'ggplot2', 'MASS', 'plyr', 'rioja', 'randomForest', 'vegan', 'maps', 'maptools', 'ggmap'))
+if(install.it == TRUE){
+  install.packages(c('gridExtra', 'ggplot2', 'MASS', 'plyr', 'rioja', 'randomForest', 'vegan', 'maps', 'maptools', 'ggmap'))
+}
 
 library(gridExtra)
 library(ggplot2)

### 2014-03-25 R/data_setup.R
Removed the install thingy, added a loader for mat.res, because it was missing.

@@ -1,6 +1,6 @@
-if(install.it == TRUE){
-  install.packages(c('gridExtra', 'ggplot2', 'MASS', 'plyr', 'rioja', 'randomForest', 'vegan', 'maps', 'maptools', 'ggmap'))
-}
+#if(install.it == TRUE){
+#  install.packages(c('gridExtra', 'ggplot2', 'MASS', 'plyr', 'rioja', 'randomForest', 'vegan', 'maps', 'maptools', 'ggmap'))
+#}
 
 library(gridExtra)
 library(ggplot2)

### 2014-03-25 R/nonPar/bootstrap_mat.R

@@ -9,6 +9,11 @@ mat.res <- list(mean_prediction  = matrix(ncol=length(vals), nrow=nrow(new.pol))
                 variance         = matrix(ncol=length(vals), nrow=nrow(new.pol)),
                 expectation      = matrix(ncol=length(vals), nrow=nrow(new.pol)))
 
+
+if('mat.res.RData' %in% list.files('data/')) {
+  load('data/mat.res.RData')
+}
+
 #  Turn the distance table into a matrix.
 diag.dist <- as.matrix(dists); diag(diag.dist) <- NA
 

### 2014-03-27 R/SourceRun.R
Running BRTs, accidentally overwrote & re-ran MAT, but results should be similar.  Some minor coding.

@@ -19,4 +19,4 @@ source(paste(file.lead, 'bootstrap_rfor.R', sep =''))
 cat('Done rFor/n')
 
 source(paste(file.lead, 'bootstrap_brt.R', sep =''))
-cat('Done rFor/n')
+cat('Done BRT/n')

### 2014-03-27 R/nonPar/bootstrap_brt.R

@@ -110,7 +110,7 @@ for(k in samp){
     
   if(is.na(brt.res$mean_prediction[j,i])){
 
-    prediction <- unlist(lapply(rep(j, 50), fun = brt.run))
+    prediction <- unlist(lapply(rep(j, 50), FUN = brt.run))
   
     brt.res$mean_prediction[j,i] <- mean(prediction, na.rm=TRUE)
     brt.res$sample_size[j,i] <- sum(keep.pol[j,], na.rm=TRUE)

### 2014-03-27 R/nonPar/bootstrap_wapls.R

@@ -50,8 +50,6 @@ wapls.run <- function(j){
   
 }
 
-sfLibrary(rioja)
-
 for(i in 1:length(vals)){
   #  This runs through each analogue distance
   keep.pol <- aaply(diag.dist, 1, 

### 2014-07-08 Paper_Markdown.Rmd
Added a bibtex file for references and started more work on the text.
Found a bug in the pollen data that may cause problems down the line, there is only AbiesEast in the dataset, no AbiesWest or Abies undiff.  This might mean that I need to re-run everything again.

--- 2013-12-30 Paper_Markdown.Rmd
@@
 =========================
-Pollen assemblages are widely used for paleoclimatic reconstruction during the Holocene, but this method is also applied to pollen assemblages from the last Glacial period (Bartlein et al., 2011), as well as to earlier pollen assemblages, such as those during the Eemian (Brewer et al, 2008).
+Pollen assemblages are widely used for paleoclimatic reconstruction during the Holocene, but this method is also applied to pollen assemblages from the last Glacial period (bartlein2011pollen, kaufman2012multi), and earlier (tarasov2011progress, brewer2008climate, cheddadi1998climate).
 A particular challenge for climate reconstruction using these much earlier samples is the presence of non-analogue vegetation.
-Non-analogue pollen assemblages pose a problem for many reconstruction techniques, these vegetation assemblages may be the result of sets of climate variables that no longer co-exist, ecological conditions resulting from differing rates of species migration or release from herbivores (Gill et al., 2009), or as a result of changes in human land use (St.
-Jacques?, maybe not the best.
-.
-.).
+Non-analogue pollen assemblages pose a problem for many reconstruction techniques, these vegetation assemblages may be the result of sets of climate variables that no longer co-exist, ecological conditions resulting from differing rates of species migration () or release from herbivores (gill2009pleistocene), or as a result of changes in human land use (Goring et al.
+in review?).
 
@@
 Sites from across North America were used to build a modern analogue dataset.
+Samples were obtained from the North American Modern Pollen Database east of 100oW.
+Pollen taxa were standardized as in Gill et al.
+(in prep), using a set of 54 taxa, similar to those reported in Whitmore et al.
+(2005).
+We chose to exclude Western North America since pollen taxa such as Tsuga canadensis and Tsuga heterophylla may be morphologically indistinct, but taxa have clearly differentiated distributions, thus it may be possible to find false close-analogues using the entire North American dataset.
+While this is not, strictly speaking, a major problem for this analysis, it will raise baseline error rates for models which then cause us to underestimate the change in error as analogue distance is increased.
 
@@
 
-map <- subset(map, map$long > -170 & map$long < -30)
+map <- subset(map, map$long > -160 & map$long < -30)
 map <- subset(map, map$lat > 20 & map$lat < 83)
@@
 ggplot(data = data.frame(map), aes(long, lat)) + 
-  geom_polygon(aes(group=group), color='blue', alpha=0.2) +
+  geom_polygon(aes(group=group), color='blue', alpha=0.7) +
   geom_point(data = data.frame(climate), aes(x = LONDD, y = LATDD), alpha = 0.5) +
@@
   xlab('') + ylab('') +
-  theme(axis.text=element_text(size=12, family='serif')) +
-  coord_map(projection='albers', lat0=30, lat1=50, xlim=c(-145, -60), ylim=c(26, 75))
+  theme(axis.text=element_text(size=12, family='serif'),
+        panel.background = element_rect(fill = 'lightblue')) +
+  coord_map(projection='albers', lat0=35, lat1=55, xlim=c(-100, -60), ylim=c(35, 80))
 
 ```
-**Figure 1**. Location of sample plots used for analysis.  Samples are derived from a modern pollen dataset.
+**Figure 1**. *Location of sample plots used for analysis.  Samples are derived from a modern pollen dataset.*
+
+To analyse the effects of non-analogue pollen on model skill we chose five pollen-based reconstruction methods from the literature and apply them to the modern pollen data from eastern North America.
+The models are MAT, WA, WAPLS, BRT and randomForest.
+Each of these models has been used previously in the pollen literature, and has strengths and weaknesses.
+MAT (Overpeck, 1985) is commonly applied to reconstruct climate, for pollen, diatoms, chironomids and other taxa.
+WA has a long history in the literature, and the recent development of monotone deshrinking has apparently improved performance further (Simpson, 2014?), but WA continues to suffer fro a basic assumption of unimodal taxon distributions along climate gradients.
 
@@
 
+Mean minimum dissimilarity for pollen samples in the modern pollen dataset ranges from `r min(min.diss)` to `r min(min.diss)`, with a 95% confidence interval from `r quantile(min.diss, 0.025)` to `r quantile(min.diss, 0.975)`.
+The highest minimum analogue distance is for an unpublished site along the New York, Pennsylvannia border.
+The site is notable for very high proportions of *Fraxinus pennsylvanica/americana* pollen, more than 5x higher than the next closest site.
+The lowest minimum distance occurs across four sites, two from Thinn Lake, identified as potentially anomalous in Williams et al (2009),
+
 ```{r Fig2MinDissim, fig.width=7, fig.height=6, fig.cap='Figure: North American site locations', echo=FALSE, message=FALSE, warning=FALSE}
@@
 dissim.scores <- data.frame(dissim = c(min.diss,
-                                       as.vector(dists)[sample(length(dists), 10000)]),
+                                       as.vector(dists)[sample(length(dists), 100000)]),
                             class   = c(rep('Minimum', length(min.diss)),
-                                        rep('All', 10000)))
+                                        rep('All', 100000)))
 
 ggplot(dissim.scores) + 
-  geom_density(aes(x = dissim, fill=class), alpha=0.5) +
+  geom_density(aes(x = dissim, linetype=class), alpha=0.5, size = 1.3, from = 0) +
   theme_bw() +
   xlab('Squared-chord Dissimilarity') +
-  ylab('Density') +
+  ylab('Kernel Density') +
   theme(text = element_text(size=24, family='serif'),
         strip.text = element_text(size=14, family='serif'),
-        axis.text = element_text(size = 12, family='serif')) +
+        axis.text = element_text(size = 12, family='serif'),
+        legend.position = 'none') +
   annotate("segment", x=0.48, 
-           xend = -.32, y = 0, yend = 1, size=3, color='red') +
-  scale_x_log10() + scale_y_sqrt()
+           xend = 0.48, y = 0, yend = 0.1, size=2, color='black', alpha = 0.7) +
+  scale_y_sqrt(limits = c(0, 5.5), expand=c(0,0)) +
+  scale_x_continuous(expand=c(0,0))
 
@@
 The 95% confidence interval for the minumum spanning distance in the North American Modern Pollen Database is 0.10 - 0.48.
-We choose 0.48 as our cut-off for 'non-analogue' distances in the database.*
+We choose 0.48 (marked on the x axis) as our cut-off for 'non-analogue' distances in the database.*
 
@@
 non.ana.map <- ggplot(data = data.frame(map), aes(long, lat)) + 
-  geom_path(aes(group=group), color='black') +
+  geom_polygon(aes(group=group), color='blue', alpha=0.2) +
   geom_point(data = data.frame(climate)[mds$non.ana == FALSE,], 
@@
   geom_point(data = data.frame(climate)[mds$non.ana == TRUE,], 
-             aes(x = LONDD, y = LATDD), color = 'red') +
+             aes(x = LONDD, y = LATDD), color = 'red', size = 3) +
    theme_bw() +
-  coord_map(projection='albers', lat0=30, lat1=50, xlim=c(-140, -60), ylim=c(26, 75)) +
   xlab('Longitude') + ylab('Latitude') +
-  theme(text = element_text(size=24, family='serif'))
+  theme(axis.text=element_text(size=12, family='serif'),
+        panel.background = element_rect(fill = 'lightblue')) +
+  coord_map(projection='albers', lat0=35, lat1=55, xlim=c(-100, -60), ylim=c(35, 80))
 
@@
   xlab('Calibration Dissimilarity') +
-  ylab('Bias (RMSEP)')
+  ylab('Sample Bias (RMSEP)')
 
 ```
-**Figure 6**. *Change in bias over time.*
+**Figure 6**. *Change in model bias with increasing analogue distance.*
 
@@
 ```
-**Figure 7**. *Change in site variance over time.*
+**Figure 7**. *Change in bootstrap variance with increasing analogue distance.*
 
@@
 Climate Dynamics, 37(3-4), 775-802.
+
 Brewer, S., Guiot, J., Sánchez-Goñi, M. F., & Klotz, S. (2008). The climate in Europe during the Eemian: a multi-method approach using pollen data. Quaternary Science Reviews, 27(25), 2303-2315.

### 2014-07-08 R/bootstrap_wa.R

@@ -1,4 +1,5 @@
 library(snowfall)
+library(analogue)
 
 sfStop()
 sfInit(parallel = TRUE, cpus = 6)
@@ -88,11 +89,8 @@ for(i in i:length(vals)){
     }
      
   }
-  if(ncol(wa.res$bias) > i){
-    if(!sum(is.na(wa.res$bias[,i+1])) == 0){
-      save(wa.res, file = 'data/wa.res.RData')
-    }
-  }
+  
+  save(wa.res, file = 'data/wa.res.RData')
   
   cat(i)  
   

### 2014-07-08 R/data_setup.R

@@ -13,11 +13,20 @@ library(maps)
 library(maptools)
 library(ggmap)
 library(reshape2)
+library(analogue)
 
 climate <- read.csv('data/Clim_Pollen_Zirbel.csv', row.names=1)
 mod_pol <- read.csv('data/ModP35c_East.csv', row.names=1, sep='\t')
 
-new.pol <- mod_pol[,7:ncol(mod_pol)] / rowSums(mod_pol[,7:ncol(mod_pol)])
+mod_pol <- mod_pol[mod_pol$LONDD > -100, ]
+climate <- climate[climate$LONDD > -100, ]
+
+bad.pol <- c("LATDD", "LONDD", "XXX", "XXX.1", "ELEVATION", "PolSum", 
+             'SPHAGNUM', 'CYPERACE', 'PTERIDIUM')
+
+new.pol <- mod_pol[, !colnames(mod_pol)%in% bad.pol]
+  
+new.pol <- new.pol / rowSums(new.pol, na.rm=TRUE)
 
 climate <- climate[rownames(mod_pol),]
 

### 2014-07-15 Paper_Markdown.Rmd
Found a bug in the codeing of pollen taxa that excludes western Abies from pollen assemblages, effectively changing the assemblages.  Re-ran the analysis using only eastern sample sites.

@@
 (2005).
-We chose to exclude Western North America since pollen taxa such as Tsuga canadensis and Tsuga heterophylla may be morphologically indistinct, but taxa have clearly differentiated distributions, thus it may be possible to find false close-analogues using the entire North American dataset.
+We chose to exclude Western North America since pollen taxa such as *Tsuga canadensis* and *Tsuga heterophylla* may be morphologically indistinct, but their parent plant taxa have clearly differentiated distributions, thus it may be possible to find false close-analogues using the entire North American dataset.
 While this is not, strictly speaking, a major problem for this analysis, it will raise baseline error rates for models which then cause us to underestimate the change in error as analogue distance is increased.

### 2014-07-15 R/bootstrap_mat.R

@@ -1,7 +1,7 @@
 library(snowfall)
 
 sfStop()
-sfInit(parallel = TRUE, cpus = 3)
+sfInit(parallel = TRUE, cpus = 6)
 
 # Set up the analogue distance exclusion values.
 #  This was originally a set of quantiles, but it played havoc on plotting and
@@ -94,5 +94,5 @@ for(i in i:length(vals)){
   }
     
   save(mat.res, file = 'data/mat.res.RData')
-  cat(i)
+  cat(i, '\n')
 }

### 2014-07-15 R/bootstrap_wa.R

@@ -2,7 +2,7 @@ library(snowfall)
 library(analogue)
 
 sfStop()
-sfInit(parallel = TRUE, cpus = 6)
+sfInit(parallel = TRUE, cpus = 4)
 
 # Set up the analogue distance exclusion values.
 #  This was originally a set of quantiles, but it played havoc on plotting and

### 2014-07-15 R/bootstrap_wapls.R

@@ -74,7 +74,7 @@ for(i in 1:length(vals)){
   for(j in 1:nrow(new.pol)){
     
     if(is.na(wapls.res$mean_prediction[j,i])){
-      #  Run WA with monotonic deshrinking.
+      #  Run WAPLS
       prediction <- unlist(sfLapply(rep(j, 100), fun = wapls.run))
     
       wapls.res$mean_prediction[j,i] <- mean(prediction, na.rm=TRUE)

### 2014-07-17 R/bootstrap_rfor.R
small runs for randomForest

@@ -3,7 +3,7 @@ library(snowfall)
 library(randomForest)
 
 sfStop()
-sfInit(parallel = TRUE, cpus = 6)
+sfInit(parallel = TRUE, cpus = 8)
 
 vals <- seq(0, 1, by=0.01)
 
@@ -39,9 +39,10 @@ rfor.run <- function(j){
   x <- subset.pol(set)
   
   zeros <- colSums(x$calib.pol, na.rm = TRUE) == 0
+  bad.clim <- is.na(x$calib.clim)
   
-  rfor <- try(randomForest(x=x$calib.pol[,!zeros], 
-                     y = x$calib.clim))
+  rfor <- try(randomForest(x=x$calib.pol[!bad.clim, !zeros], 
+                     y = x$calib.clim[!bad.clim]))
   
   if (length(rfor) > 1){
     output <- predict(rfor, newdata = new.pol[j,!zeros])
@@ -54,11 +55,9 @@ rfor.run <- function(j){
 }
 
 #  Parallelize:
-sfExport(list = list('rfor.run'))
-sfExport(list = list('new.pol'))
-sfExport(list = list('subset.pol'))
-sfExport(list = list('climate'))
+sfExport(list = list('rfor.run', 'new.pol', 'subset.pol', 'climate'))
 sfLibrary(randomForest)
+sfClusterSetupRNG()
 
 #  Create a set of all possible xy pairs in the rfor tables, and then look to see if
 #  they've been sampled yet.
@@ -103,7 +102,7 @@ for(k in samp){
     
   if(is.na(rfor.res$mean_prediction[j,i])){
     #  Run randomForest with raw defaults
-    prediction <- unlist(sfLapply(rep(j, 50), fun = rfor.run))
+    prediction <- unlist(sfLapply(rep(j, 30), fun = rfor.run))
   
     rfor.res$mean_prediction[j,i] <- mean(prediction, na.rm=TRUE)
     rfor.res$sample_size[j,i] <- sum(keep.pol[j,], na.rm=TRUE)

### 2014-07-17 R/bootstrap_brt.R
Changed bootstraps to 30.

@@ -124,7 +124,7 @@ for(k in samp){
     
   if(is.na(brt.res$mean_prediction[j,i])){
 
-    prediction <- unlist(sfLapply(rep(j, 50), fun = brt.run))
+    prediction <- unlist(sfLapply(rep(j, 30), fun = brt.run))
   
     brt.res$mean_prediction[j,i] <- mean(prediction, na.rm=TRUE)
     brt.res$sample_size[j,i] <- sum(keep.pol[j,], na.rm=TRUE)

### 2014-07-17 R/bootstrap_brt.R
Added the bad.clim to each of the bootstrap functions.  Now I've got all of them running with only thirty samples so everything is effectively equivalent.

@@ -50,16 +50,23 @@ brt.run <- function(j){
   x <- subset.pol(set)
   
   zeros <- colSums(x$calib.pol, na.rm = TRUE) == 0
+  bad.clim <- is.na(x$calib.clim)
 
   # Construct the brt formula based on x$calib.clim and x$calib.pol column 
   # names, omitting the zero-sum taxa which won't be used.
   
-  brt.formula <- as.formula(paste (colnames(x$calib.clim)[1], "~", paste(colnames(x$calib.pol)[!zeros], collapse=" + ")))     
+  brt.formula <- as.formula(paste (colnames(x$calib.clim)[1], "~", 
+                                   paste(colnames(x$calib.pol)[!zeros], collapse=" + ")))
 
   # Construct brt model. x$calib.clim and x$calib.pol combined into a single
   # data frame and passed as the "data" parameter.
   
-  brt.model <- gbm(brt.formula, distribution="gaussian", n.trees=500, shrinkage=0.005, interaction.depth=6, cv.folds=4, verbose=FALSE, data=cbind(x$calib.clim, x$calib.pol[,!zeros]))
+  brt.model <- gbm(brt.formula, distribution="gaussian", 
+                   n.trees=500, shrinkage=0.005, 
+                   interaction.depth=6, cv.folds=4, 
+                   verbose=FALSE, 
+                   data=cbind(x$calib.clim, 
+                              x$calib.pol[,!zeros])[!bad.clim,])
 
   # This probably always returns 500
   brt.ntrees <- gbm.perf(brt.model, method="cv", plot.it=FALSE)   
@@ -118,6 +125,7 @@ for(k in samp){
   
   keep.pol <- aaply(diag.dist, 1,
                     function(x) {x > vals[i]})
+  
   diag(keep.pol) <- FALSE
   
   sfExport(list = list('keep.pol'))

### 2014-07-17 R/bootstrap_mat.R

@@ -80,7 +80,7 @@ for(i in i:length(vals)){
 
   for(j in 1:nrow(new.pol)){
 
-    prediction <- unlist(sfLapply(1:100, mat.fun, j = j, min = which.min(rmse)))
+    prediction <- unlist(sfLapply(1:30, mat.fun, j = j, min = which.min(rmse)))
     
     if(is.na(mat.res$mean_prediction[j,i])){
       mat.res$mean_prediction[j,i] <- mean(prediction, na.rm=TRUE)

### 2014-07-17 R/bootstrap_wa.R

@@ -40,9 +40,10 @@ wa.run <- function(j){
   x <- subset.pol(set)
   
   zeros <- colSums(x$calib.pol, na.rm = TRUE) == 0
+  bad.clim <- is.na(x$calib.clim)
   
-  pred.wa <- try(predict(wa(x=x$calib.pol, 
-                            env = x$calib.clim, 
+  pred.wa <- try(predict(wa(x=x$calib.pol[!bad.clim, ], 
+                            env = x$calib.clim[!bad.clim], 
                             deshrinking='monotonic'), 
                          newdata = new.pol[j,!x$zeros]))
   
@@ -79,7 +80,7 @@ for(i in i:length(vals)){
     
     if(is.na(wa.res$mean_prediction[j,i])){
       #  Run WA with monotonic deshrinking.
-      prediction <- unlist(sfLapply(rep(j, 100), fun = wa.run))
+      prediction <- unlist(sfLapply(rep(j, 30), fun = wa.run))
     
       wa.res$mean_prediction[j,i] <- mean(prediction, na.rm=TRUE)
       wa.res$sample_size[j,i] <- sum(keep.pol[j,], na.rm=TRUE)

### 2014-07-17 R/bootstrap_wapls.R

@@ -40,9 +40,10 @@ wapls.run <- function(j){
   x <- subset.pol(set)
   
   zeros <- colSums(x$calib.pol, na.rm = TRUE) == 0
+  bad.clim <- is.na(x$calib.clim)
   
-  wapls <- try(WAPLS(y=x$calib.pol[,!zeros], 
-                     x = x$calib.clim))
+  wapls <- try(WAPLS(y=x$calib.pol[!bad.clim, !zeros], 
+                     x = x$calib.clim[!bad.clim]))
   
   if (length(wapls) > 1){
     pred.wapls <- predict(wapls, newdata = new.pol[j,!zeros])

### 2014-07-17 R/bootstrap_rfor.R
Running as longlist

@@ -3,7 +3,7 @@ library(snowfall)
 library(randomForest)
 
 sfStop()
-sfInit(parallel = TRUE, cpus = 8)
+sfInit(parallel = TRUE, cpus = 30)
 
 vals <- seq(0, 1, by=0.01)
 
@@ -34,7 +34,7 @@ subset.pol <- function(set){
 
 rfor.run <- function(j){
   
-  set <- keep.pol[j,]
+  set <- keep.pol
   
   x <- subset.pol(set)
   
@@ -89,36 +89,42 @@ longlist <- longlist[longlist[,3],]
 
 for(k in samp){
   #  This runs through each analogue distance
+  
   run.time <- proc.time()
   
   i <- longlist[k,2]
   j <- longlist[k,1]
   
-  keep.pol <- aaply(diag.dist, 1, 
-                    function(x) {x > vals[i]})
-  diag(keep.pol) <- FALSE
-  
-  sfExport(list = list('keep.pol'))
-    
+  if(!is.na(rfor.res$mean_prediction[j,i])){
+    cat('.')
+  }
   if(is.na(rfor.res$mean_prediction[j,i])){
+    
+    keep.pol <- aaply(diag.dist, 1, 
+                      function(x) {x[j] > vals[i]})
+    
+    keep.pol[j] <- FALSE
+    
+    sfExport(list = list('keep.pol'))
+      
     #  Run randomForest with raw defaults
     prediction <- unlist(sfLapply(rep(j, 30), fun = rfor.run))
   
     rfor.res$mean_prediction[j,i] <- mean(prediction, na.rm=TRUE)
-    rfor.res$sample_size[j,i] <- sum(keep.pol[j,], na.rm=TRUE)
+    rfor.res$sample_size[j,i] <- sum(keep.pol[j], na.rm=TRUE)
     rfor.res$bias[j, i] <- (climate[j,10] - mean(prediction, na.rm=TRUE))^2
     rfor.res$expectation[j, i]  <- mean((climate[j,10] - prediction)^2)
     rfor.res$variance[j, i]  <- mean((mean(prediction) - prediction)^2)
+  
+    save(rfor.res, file = 'data/rfor.res.RData')
+    
+    end.time <- proc.time()
+    
+    st <- Sys.time()
+    
+    cat(round(sum(!is.na(rfor.res$bias))/length(rfor.res$bias) * 100, 4), '% done on', 
[…]

### 2014-07-17 R/shared_run.R
Running as longlist part 2

@@ -9,10 +9,11 @@ vals <- seq(0, 1, by=0.01)
 
 #  Sakari, I doubt it will, but you need to make sure that 'goring' isn't anywhere in your
 #  working directory path!
-analyst <- ifelse(regexpr('goring', tolower(getwd()))>0, 'Simon', 'Sakari')
+#analyst <- ifelse(regexpr('goring', tolower(getwd()))>0, 'Simon', 'Sakari')
+analyst <- 'Simon'
 
 if(paste('longlist.', analyst, '.RData', sep='') %in% list.files('data')){
-  load(paste('data/longlist.', analyst, '.RData', sep=''))
+  load(paste0('data/longlist.', analyst, '.RData'))
 }
 
 if(!exists('longlist')){
@@ -29,11 +30,10 @@ if(exists('longlist')){
 save(longlist, file=paste('data/longlist.', analyst, '.RData', sep=''))
 
 library(snowfall)
-
 library(gbm)
 
 sfStop()
-sfInit(parallel = TRUE, cpus = 2)
+sfInit(parallel = TRUE, cpus = 30)
 
 subset.pol <- function(set){
   
@@ -52,11 +52,12 @@ subset.pol <- function(set){
 
 brt.run <- function(px, pa){
   
-  set <- keep.pol[px,]
+  set <- keep.pol
   
   x <- subset.pol(set)
   
   zeros <- colSums(x$calib.pol, na.rm = TRUE) == 0
+  bad.clim <- is.na(x$calib.clim)
   
   # Construct the brt formula based on x$calib.clim and x$calib.pol column 
   # names, omitting the zero-sum taxa which won't be used.
@@ -66,7 +67,10 @@ brt.run <- function(px, pa){
   # Construct brt model. x$calib.clim and x$calib.pol combined into a single
   # data frame and passed as the "data" parameter.
   
-  brt.model <- gbm(brt.formula, distribution="gaussian", n.trees=500, shrinkage=0.005, interaction.depth=6, cv.folds=4, verbose=FALSE, data=cbind(x$calib.clim, x$calib.pol[,!zeros]))
+  brt.model <- gbm(brt.formula, distribution="gaussian", 
+                   n.trees=500, shrinkage=0.005, interaction.depth=6, 
+                   cv.folds=4, verbose=FALSE, 
+                   data=cbind(x$calib.clim, x$calib.pol[,!zeros])[!bad.clim,])
   
[…]

### 2014-07-18 R/shared_run.R
0.58% of runs.

@@ -67,15 +67,14 @@ brt.run <- function(px, pa){
   # Construct brt model. x$calib.clim and x$calib.pol combined into a single
   # data frame and passed as the "data" parameter.
   
-  brt.model <- gbm(brt.formula, distribution="gaussian", 
+  brt.model <- try(gbm(brt.formula, distribution="gaussian", 
                    n.trees=500, shrinkage=0.005, interaction.depth=6, 
                    cv.folds=4, verbose=FALSE, 
-                   data=cbind(x$calib.clim, x$calib.pol[,!zeros])[!bad.clim,])
+                   data=cbind(x$calib.clim, x$calib.pol[,!zeros])[!bad.clim,]))
   
+  if(length(brt.model) > 1){
   # This probably always returns 500
-  brt.ntrees <- gbm.perf(brt.model, method="cv", plot.it=FALSE)   
-  
-  if (length(brt.model) > 1){
+    brt.ntrees <- gbm.perf(brt.model, method="cv", plot.it=FALSE)   
     output <- predict(brt.model, new.pol[px,!zeros], brt.ntrees, type="response")
   }
   else{

### 2015-01-13 Paper_Markdown.Rmd
Mostly text and mapping changes to the Paper_Markdown.Rmd` file. * Moved "shared" run files to the `shared run` folder to provide a clean folder with just the bootstrap code. * Moved a few files to the `deprecated` folder. * Made sure to explicitly call the `rioja` version of `performance` in the WAPLS code.

@@
+---
+output: word_document
+---
 Non-analogues in paleoecological reconstruction - the effects of analogue distance
@@
 for(i in 1:ncol(clim)){
-  test.struct <- rda(pol ~ clim[,i])
+  test.struct <- rda(pol ~ clim[,i], na.action = na.omit)
   clim.ratio[i] <- test.struct$CCA$tot.chi / test.struct$CA$tot.chi
@@
 
+point.loss <- 1 - mat.res$sample_size[,49] / nrow(climate)
+
 na.samples <- data.frame(site = rep(climate[,1], ncol(mat.res$bias)),
+                         lat = rep(climate$LATDD, ncol(mat.res$bias)),
+                         long = rep(climate$LONDD, ncol(mat.res$bias)),
+                         loss = point.loss,
+                         bin = factor(round(ecdf(point.loss)(point.loss), 1)),
                          size = as.vector(mat.res$sample_size),
@@
 
-loss <- ggplot(na.samples, aes(x=distance, y=size)) + 
-  geom_path(aes(group=site), alpha=0.01) +
+loss <- ggplot(data = na.samples, aes(x=distance, y=size)) + 
+  geom_path(aes(group=site, color = bin)) +
+  scale_color_brewer(palette = 'PiYG', type = 'div') +
   xlab('Analogue Distance') + ylab('Available Samples') +
-  theme_bw() +
   theme(text = element_text(size=24, family='serif'),
-        axis.text = element_text(size=14, family='serif')) +
+        axis.text = element_text(size=14, family='serif'),
+        legend.position = 'none') +
   scale_x_continuous(expand = c(0,0)) +
   scale_y_continuous(expand=c(0,0)) +
-  geom_segment(aes(x=0.48, xend = 0.48, y = min(mat.res$sample_size, na.rm=TRUE), yend = nrow(climate)), size=1, color='red', linetype=2)
+  geom_segment(aes(x=0.48, 
+                   xend = 0.48, 
+                   y = min(mat.res$sample_size, na.rm=TRUE), 
+                   yend = nrow(climate)), 
+               size=1, color='red', linetype=2)
 
@@
                        long = climate$LONDD,
-                       loss = 1 - mat.res$sample_size[,49] / nrow(climate))
+                       loss = point.loss,
+                       bin  = factor(round(ecdf(point.loss)(point.loss), 1)))
 
 loss.map <-  ggplot(data = data.frame(map), aes(long, lat)) + 
-  geom_path(aes(group=group), color='black') +
+  geom_polygon(aes(group=group), color='blue', alpha=0.2) +
   geom_point(data = loc.loss, 
-             aes(x = long, y = lat, color = loss), alpha=0.5) +
+             aes(x = long, y = lat, color = bin), alpha=0.7, size = 3) +
   theme_bw() +
-  theme(text = element_text(size=24, family='serif'),
-        axis.text = element_text(size=14, family='serif')) +
-  coord_map(projection='albers', lat0=30, lat1=50, xlim=c(-140, -60), ylim=c(26, 75)) +
-  scale_color_continuous(low = '#F08080', high='#4169E1') +
-  xlab('') + ylab('')
+  xlab('Longitude') + ylab('Latitude') +
+  theme(axis.text=element_text(size=12, family='serif'),
+        panel.background = element_rect(fill = 'lightblue'),
+        legend.position = 'none') +
+  coord_map(projection='albers', lat0=35, lat1=55, xlim=c(-100, -60), ylim=c(35, 80)) +
+  scale_color_brewer(palette = 'PiYG', type = 'div')
+  #scale_color_continuous(low = '#F08080', high='#4169E1')
 
@@
 Points that maintain a large pool of analogues are far from neighbours.*
+
+We have an interesting distribution of loss here.
+First, the color coding is based on loss to 'non-analogue' conditions, and we see that it is possible for some samples to have relatively low loss to the no-analogue point, but then much higher losses to the extreme end of our analogue space (see the magenta lines at the very bottom end of Figure 4a.).
+Secondly, the points that lose the most samples tend to be in the boreal forest.
+To some degree this is surprising, however, the dominance of Pinus and Picea in these forests likely means that they have smaller analogue distances from one another than samples in the southern forests (with more diversity and therfore, likely fewer close analogues), but also, surpisingly, arctic samples also have higer inter-sample distances, and are thus less sensitive to losses of close analogues.
 

### 2015-01-13 R/deprecated/BRT_time_test.r

@@ -1,0 +1,51 @@
+# Test version of the BRT cross-validation code.
+#
+# This version does not save the results, but only tests the time requirement of the BRT
+# cross validation. 23 samples (i.e. 1% of total) are selected at random, and 50 bootstrap
+# iterations are performed for each sample. Predicted value and used time are printed
+# in each iteration.
+
+library (gbm)
+
+# Read the data from Excel. Note that the data must be organised as in NA_pollen_for_BRT.csv
+# (same column names and column order)
+
+cal_data <- sqrt(read.table("clipboard", header=TRUE))
+attach (cal_data)
+
+# Construct the formula string for gbm
+
+long_formula <- as.formula(paste("tjul ~ ", paste(paste("v", 1:56, sep = ""), collapse = " + "), sep = ""))
+
+# Select 23 random samples to test with
+
+test_samples <- sample(x=2289, size=23, replace=FALSE)
+
+# Run 50 bootstrap cycles with 23 random samples. The predicted values are not stored: 
+# they are over-written in each iteration in variable "pred".
+
+start_time <- Sys.time()
+pred <- numeric()
+
+for (i in 1:23) {
+
+   for (j in 1:50) {
+   
+      iter_start_time <- Sys.time()
+      calibration_set <- cal_data[sample(x=2289, size=2289, replace=TRUE),]
+      attach (calibration_set, warn.conflicts = FALSE)
+      xval_model <- gbm(long_formula , distribution="gaussian", n.trees=500, shrinkage=0.005, interaction.depth=6, cv.folds = 4, verbose=FALSE)  # Constructs the BRT model
+      xval_best_iter <- gbm.perf(xval_model,method="cv",plot.it=FALSE)  # Returns the number of trees that gives the best performance
+      test_set <- cal_data[test_samples[i],4:59]
+      attach (test_set, warn.conflicts = FALSE)
+      pred <- predict(xval_model,test_set,xval_best_iter, type="response") # Predicts for the test-set sample
+      iter_end_time <- Sys.time()
+      print(paste("Sample ",test_samples[i]," (",i,"/23), bootstrap cycle ",j,"/50, predicted value ",pred,".", sep=""))
+      print (iter_end_time - iter_start_time)
+	  
+   }
+   
+}
+
+end_time <- Sys.time()
+print (end_time-start_time)

### 2015-01-13 R/deprecated/bootstrapping.noaanalogs.R

@@ -1,0 +1,171 @@
+library(analogue)
+library(MASS)
+library(plyr)
+library(rioja)
+library(randomForest)
+library(vegan)
+
+climate <- read.csv('data/Clim_Pollen_Zirbel.csv', row.names=1)
+mod_pol <- read.csv('data/ModP35c_East.csv', row.names=1, sep='\t')
+
+new.pol <- mod_pol[,7:ncol(mod_pol)] / rowSums(mod_pol[,7:ncol(mod_pol)])
+
+climate <- climate[rownames(mod_pol),]
+
+dists <- dist(new.pol^(1/2))
+
+dists[dists==0] <- min(dists[dists > 0], na.rm=TRUE)
+
+if(!'pol.mds.RData' %in% list.files('data/')){
+  pol.mds <- isoMDS(dists, k=2)
+  save(pol.mds, file='data/pol.mds.RData')
+}
+else{
+  load('data/pol.mds.RData')
+}
+
+#  Find the density of minimum distances:
+diag.dist <- as.matrix(dists); diag(diag.dist) <- NA
+
+min.diss <- apply(diag.dist, 2, min, na.rm=TRUE)
+max.diss <- apply(diag.dist, 2, max, na.rm=TRUE)
+
+min.dist.dens <- density(min.diss, from=0, to=1.42, n=100)
+max.dist.dens <- density(max.diss, from=0, to=1.42, n=100)
+
+quantiles <- c(0, quantile(ecdf(diag.dist), c(seq(0, 0.2, by=0.001), seq(0.21, 1, by=0.01))))
+
+vals <- c(seq(0, 0.2, by=0.001), seq(0.21, 1, by=0.01))
+
+mat.res <- list(prediction     = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                r.squared      = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                closest.n      = matrix(ncol=length(vals), nrow=nrow(new.pol)))
+
+wa.res <- list(prediction      = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+               r.squared       = matrix(ncol=length(vals), nrow=nrow(new.pol)))
+
+wapls.res <- list(prediction   = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                  r.squared    = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                  components.n = matrix(ncol=length(vals), nrow=nrow(new.pol)))
+
[…]

### 2015-01-13 R/recombine_analysts.R

@@ -1,23 +1,0 @@
-#  Taking the shared run files and stitching them back into the respective boot files:
-library(reshape2)
-
-source('R/data_setup.R')
-
-load('data/total.output.RData')
-load('data/longlist.Simon.RData')
-
-dump.it <- function(x, y, set){
-  
-  dump.set <- x[x$method == set,]
-  
-  y$mean_prediction <- acast(data=dump.set, Var1 ~ Var2, fun.aggregate=sum, value.var='mean_prediction')
-  y$bias <- acast(data=dump.set, Var1 ~ Var2, fun.aggregate=sum, value.var='bias')
-  y$expectation <- acast(data=dump.set, Var1 ~ Var2, fun.aggregate=sum, value.var='expectation')
-  y$variance <- acast(data=dump.set, Var1 ~ Var2, fun.aggregate=sum, value.var='variance')
-  y$sample_size <- acast(data=dump.set, Var1 ~ Var2, fun.aggregate=sum, value.var='sample_size')
-  
-  y
-}
-
-rfor.res <- dump.it(longlist, rfor.res, 'rfor')
-brt.res <- dump.it(longlist, brt.res, 'brt')

### 2015-01-13 R/shared run/recombine_analysts.R

@@ -1,0 +1,28 @@
+#  Taking the shared run files and stitching them back into the respective boot files:
+library(reshape2)
+
+source('R/data_setup.R')
+
+load('data/total.output.RData')
+load('data/longlist.Simon.RData')
+
+
+dump.it <- function(x, y, set){
+  
+  dump.set <- x[x$method == set,]
+  
+  y$mean_prediction <- acast(data=dump.set, Var1 ~ Var2, fun.aggregate=sum, value.var='mean_prediction')
+  y$bias <- acast(data=dump.set, Var1 ~ Var2, fun.aggregate=sum, value.var='bias')
+  y$expectation <- acast(data=dump.set, Var1 ~ Var2, fun.aggregate=sum, value.var='expectation')
+  y$variance <- acast(data=dump.set, Var1 ~ Var2, fun.aggregate=sum, value.var='variance')
+  y$sample_size <- acast(data=dump.set, Var1 ~ Var2, fun.aggregate=sum, value.var='sample_size')
+  
+  
+}
+
+rfor.res <- dump.it(longlist, rfor.res, 'rfor')
+brt.res <- dump.it(longlist, brt.res, 'brt')
+
+load('data/longlist.Work.RData')
+rfor.res <- dump.it(longlist, rfor.res, 'rfor')
+brt.res <- dump.it(longlist, brt.res, 'brt')

### 2015-01-13 R/shared_run.R

@@ -1,205 +1,0 @@
-#  Sakari, make sure that the default directory is set as the main 'NoAnalogueError' directory,
-#  and not the R directory.
-
-source('R/data_setup.R')
-
-load('data/total.output.RData')
-
-vals <- seq(0, 1, by=0.01)
-
-#  Sakari, I doubt it will, but you need to make sure that 'goring' isn't anywhere in your
-#  working directory path!
-#analyst <- ifelse(regexpr('goring', tolower(getwd()))>0, 'Simon', 'Sakari')
-analyst <- 'Work'
-
-if(paste('longlist.', analyst, '.RData', sep='') %in% list.files('data')){
-  load(paste0('data/longlist.', analyst, '.RData'))
-}
-
-if(!exists('longlist')){
-  longlist <- data.frame(total.output, 
-                         calc = !is.na(total.output[,3]))
-}
-if(exists('longlist')){
-  if(any(is.na(longlist$bias) & !is.na(total.output$bias))){
-    longlist[is.na(longlist$bias) & !is.na(total.output$bias), colnames(total.output)] <-
-      total.output[is.na(longlist$bias) & !is.na(total.output$bias), ]
-  }
-}
-
-save(longlist, file=paste('data/longlist.', analyst, '.RData', sep=''))
-
-library(snowfall)
-library(gbm)
-
-sfStop()
-sfInit(parallel = TRUE, cpus = 3)
-
-
-subset.pol <- function(set){
-  
-  calib.samples <- sample((1:length(set))[set], replace=TRUE)
-  
-  zeros <- colSums(new.pol[calib.samples,], na.rm = TRUE) == 0
-  
-  # Modified the following to return the climate values as a 
-  # single-column data frame.
-  
-  list(calib.pol = new.pol[calib.samples,!zeros],
-       calib.clim = climate[calib.samples,10,drop=FALSE], 
-       zeros = zeros)   
-  
-}
-
-brt.run <- function(px, pa){
-  
-  set <- keep.pol
-  
-  x <- subset.pol(set)
-  
-  zeros <- colSums(x$calib.pol, na.rm = TRUE) == 0
-  bad.clim <- is.na(x$calib.clim)
-  
-  # Construct the brt formula based on x$calib.clim and x$calib.pol column 
-  # names, omitting the zero-sum taxa which won't be used.
-  
[…]

### 2015-05-06 R/SourceRun.R
Turned on the parallel flag for the source run.

@@ -1,7 +1,7 @@
 #  Run all bootstraps:
 source('R/data_setup.R')
 
-par <- FALSE
+par <- TRUE
 
 file.lead <- ifelse(par == TRUE, 'R/', 'R/nonPar/')
 

### 2015-05-06 R/bootstrap_brt.R
Reformatted the brt runs so that the lines look nicer, reduced the bootstrap size to 30 and increased the number of available cores to 30.
Reformatted the randomForest runs, increased the cores to 30, dropped the bootstraps to 30, moved some text around, but didn't make too many major changes.

MAT up to 30 cores as well, 300 bootstraps.

WA up to 30 cores, 30 bootstraps.

WAPLS up to 30 cores, 30 bootstraps.

@@ -11,7 +11,7 @@ library(snowfall)
 library(gbm)
 
 sfStop()
-sfInit(parallel = TRUE, cpus = 6)
+sfInit(parallel = TRUE, cpus = 30)
 
 vals <- seq(0, 1, by=0.01)
 
@@ -50,16 +50,23 @@ brt.run <- function(j){
   x <- subset.pol(set)
   
   zeros <- colSums(x$calib.pol, na.rm = TRUE) == 0
+  bad.clim <- is.na(x$calib.clim)
 
   # Construct the brt formula based on x$calib.clim and x$calib.pol column 
   # names, omitting the zero-sum taxa which won't be used.
   
-  brt.formula <- as.formula(paste (colnames(x$calib.clim)[1], "~", paste(colnames(x$calib.pol)[!zeros], collapse=" + ")))     
+  brt.formula <- as.formula(paste (colnames(x$calib.clim)[1], "~", 
+                                   paste(colnames(x$calib.pol)[!zeros], collapse=" + ")))
 
   # Construct brt model. x$calib.clim and x$calib.pol combined into a single
   # data frame and passed as the "data" parameter.
   
-  brt.model <- gbm(brt.formula, distribution="gaussian", n.trees=500, shrinkage=0.005, interaction.depth=6, cv.folds=4, verbose=FALSE, data=cbind(x$calib.clim, x$calib.pol[,!zeros]))
+  brt.model <- gbm(brt.formula, distribution="gaussian", 
+                   n.trees=500, shrinkage=0.005, 
+                   interaction.depth=6, cv.folds=4, 
+                   verbose=FALSE, 
+                   data=cbind(x$calib.clim, 
+                              x$calib.pol[,!zeros])[!bad.clim,])
 
   # This probably always returns 500
   brt.ntrees <- gbm.perf(brt.model, method="cv", plot.it=FALSE)   
@@ -118,13 +125,14 @@ for(k in samp){
   
   keep.pol <- aaply(diag.dist, 1,
                     function(x) {x > vals[i]})
+  
   diag(keep.pol) <- FALSE
   
   sfExport(list = list('keep.pol'))
     
   if(is.na(brt.res$mean_prediction[j,i])){
 
-    prediction <- unlist(sfLapply(rep(j, 50), fun = brt.run))
+    prediction <- unlist(sfLapply(rep(j, 30), fun = brt.run))
   
[…]

### 2015-05-06 R/bootstrap_mat.R

@@ -1,7 +1,7 @@
 library(snowfall)
 
 sfStop()
-sfInit(parallel = TRUE, cpus = 3)
+sfInit(parallel = TRUE, cpus = 30)
 
 # Set up the analogue distance exclusion values.
 #  This was originally a set of quantiles, but it played havoc on plotting and
@@ -80,7 +80,7 @@ for(i in i:length(vals)){
 
   for(j in 1:nrow(new.pol)){
 
-    prediction <- unlist(sfLapply(1:100, mat.fun, j = j, min = which.min(rmse)))
+    prediction <- unlist(sfLapply(1:30, mat.fun, j = j, min = which.min(rmse)))
     
     if(is.na(mat.res$mean_prediction[j,i])){
       mat.res$mean_prediction[j,i] <- mean(prediction, na.rm=TRUE)
@@ -94,5 +94,5 @@ for(i in i:length(vals)){
   }
     
   save(mat.res, file = 'data/mat.res.RData')
-  cat(i)
+  cat(i, '\n')
 }

### 2015-05-06 R/bootstrap_rfor.R

@@ -3,7 +3,7 @@ library(snowfall)
 library(randomForest)
 
 sfStop()
-sfInit(parallel = TRUE, cpus = 6)
+sfInit(parallel = TRUE, cpus = 30)
 
 vals <- seq(0, 1, by=0.01)
 
@@ -34,14 +34,15 @@ subset.pol <- function(set){
 
 rfor.run <- function(j){
   
-  set <- keep.pol[j,]
+  set <- keep.pol
   
   x <- subset.pol(set)
   
   zeros <- colSums(x$calib.pol, na.rm = TRUE) == 0
+  bad.clim <- is.na(x$calib.clim)
   
-  rfor <- try(randomForest(x=x$calib.pol[,!zeros], 
-                     y = x$calib.clim))
+  rfor <- try(randomForest(x=x$calib.pol[!bad.clim, !zeros], 
+                     y = x$calib.clim[!bad.clim]))
   
   if (length(rfor) > 1){
     output <- predict(rfor, newdata = new.pol[j,!zeros])
@@ -54,11 +55,9 @@ rfor.run <- function(j){
 }
 
 #  Parallelize:
-sfExport(list = list('rfor.run'))
-sfExport(list = list('new.pol'))
-sfExport(list = list('subset.pol'))
-sfExport(list = list('climate'))
+sfExport(list = list('rfor.run', 'new.pol', 'subset.pol', 'climate'))
 sfLibrary(randomForest)
+sfClusterSetupRNG()
 
 #  Create a set of all possible xy pairs in the rfor tables, and then look to see if
 #  they've been sampled yet.
@@ -90,36 +89,42 @@ longlist <- longlist[longlist[,3],]
 
 for(k in samp){
   #  This runs through each analogue distance
+  
   run.time <- proc.time()
   
   i <- longlist[k,2]
   j <- longlist[k,1]
   
-  keep.pol <- aaply(diag.dist, 1, 
-                    function(x) {x > vals[i]})
-  diag(keep.pol) <- FALSE
-  
-  sfExport(list = list('keep.pol'))
-    
+  if(!is.na(rfor.res$mean_prediction[j,i])){
+    cat('.')
+  }
   if(is.na(rfor.res$mean_prediction[j,i])){
+    
+    keep.pol <- aaply(diag.dist, 1, 
+                      function(x) {x[j] > vals[i]})
+    
+    keep.pol[j] <- FALSE
+    
[…]

### 2015-05-06 R/bootstrap_wa.R

@@ -1,7 +1,8 @@
 library(snowfall)
+library(analogue)
 
 sfStop()
-sfInit(parallel = TRUE, cpus = 6)
+sfInit(parallel = TRUE, cpus = 30)
 
 # Set up the analogue distance exclusion values.
 #  This was originally a set of quantiles, but it played havoc on plotting and
@@ -39,9 +40,10 @@ wa.run <- function(j){
   x <- subset.pol(set)
   
   zeros <- colSums(x$calib.pol, na.rm = TRUE) == 0
+  bad.clim <- is.na(x$calib.clim)
   
-  pred.wa <- try(predict(wa(x=x$calib.pol, 
-                            env = x$calib.clim, 
+  pred.wa <- try(predict(wa(x=x$calib.pol[!bad.clim, ], 
+                            env = x$calib.clim[!bad.clim], 
                             deshrinking='monotonic'), 
                          newdata = new.pol[j,!x$zeros]))
   
@@ -78,7 +80,7 @@ for(i in i:length(vals)){
     
     if(is.na(wa.res$mean_prediction[j,i])){
       #  Run WA with monotonic deshrinking.
-      prediction <- unlist(sfLapply(rep(j, 100), fun = wa.run))
+      prediction <- unlist(sfLapply(rep(j, 30), fun = wa.run))
     
       wa.res$mean_prediction[j,i] <- mean(prediction, na.rm=TRUE)
       wa.res$sample_size[j,i] <- sum(keep.pol[j,], na.rm=TRUE)
@@ -88,11 +90,8 @@ for(i in i:length(vals)){
     }
      
   }
-  if(ncol(wa.res$bias) > i){
-    if(!sum(is.na(wa.res$bias[,i+1])) == 0){
-      save(wa.res, file = 'data/wa.res.RData')
-    }
-  }
+  
+  save(wa.res, file = 'data/wa.res.RData')
   
   cat(i)  
   

### 2015-05-06 R/bootstrap_wapls.R

@@ -1,7 +1,7 @@
 library(snowfall)
 
 sfStop()
-sfInit(parallel = TRUE, cpus = 6)
+sfInit(parallel = TRUE, cpus = 30)
 
 # Set up the analogue distance exclusion values.
 #  This was originally a set of quantiles, but it played havoc on plotting and
@@ -40,9 +40,10 @@ wapls.run <- function(j){
   x <- subset.pol(set)
   
   zeros <- colSums(x$calib.pol, na.rm = TRUE) == 0
+  bad.clim <- is.na(x$calib.clim)
   
-  wapls <- try(WAPLS(y=x$calib.pol[,!zeros], 
-                     x = x$calib.clim))
+  wapls <- try(WAPLS(y=x$calib.pol[!bad.clim, !zeros], 
+                     x = x$calib.clim[!bad.clim]))
   
   if (length(wapls) > 1){
     pred.wapls <- predict(wapls, newdata = new.pol[j,!zeros])
@@ -74,8 +75,8 @@ for(i in 1:length(vals)){
   for(j in 1:nrow(new.pol)){
     
     if(is.na(wapls.res$mean_prediction[j,i])){
-      #  Run WA with monotonic deshrinking.
-      prediction <- unlist(sfLapply(rep(j, 100), fun = wapls.run))
+      #  Run WAPLS
+      prediction <- unlist(sfLapply(rep(j, 30), fun = wapls.run))
     
       wapls.res$mean_prediction[j,i] <- mean(prediction, na.rm=TRUE)
       wapls.res$sample_size[j,i] <- sum(keep.pol[j,], na.rm=TRUE)

### 2015-05-06 R/data_setup.R
`data_setup` gets more explicit cleaning of the pollen data file, so that the columns are explicitly excized and we clearly remove everything west of 100^o^ Long.
I also removed all of the old RData files.  I'm going to switch to RDS files, and also the old stuff is junk.

@@ -13,11 +13,20 @@ library(maps)
 library(maptools)
 library(ggmap)
 library(reshape2)
+library(analogue)
 
 climate <- read.csv('data/Clim_Pollen_Zirbel.csv', row.names=1)
 mod_pol <- read.csv('data/ModP35c_East.csv', row.names=1, sep='\t')
 
-new.pol <- mod_pol[,7:ncol(mod_pol)] / rowSums(mod_pol[,7:ncol(mod_pol)])
+mod_pol <- mod_pol[mod_pol$LONDD > -100, ]
+climate <- climate[climate$LONDD > -100, ]
+
+bad.pol <- c("LATDD", "LONDD", "XXX", "XXX.1", "ELEVATION", "PolSum", 
+             'SPHAGNUM', 'CYPERACE', 'PTERIDIUM')
+
+new.pol <- mod_pol[, !colnames(mod_pol)%in% bad.pol]
+  
+new.pol <- new.pol / rowSums(new.pol, na.rm=TRUE)
 
 climate <- climate[rownames(mod_pol),]
 
@@ -42,9 +51,3 @@ max.diss <- apply(diag.dist, 2, max, na.rm=TRUE)
 
 min.dist.dens <- density(min.diss, from=0, to=1.5, n=100)
 max.dist.dens <- density(max.diss, from=0, to=1.5, n=100)
-
-load('data/wa.res.RData')
-load('data/mat.res.RData')
-load('data/wapls.res.RData')
-load('data/rfor.res.RData')
-load('data/brt.res.RData')

### 2015-05-06 R/shared_run.R
I've modified Shared run but won't be using it since I'm going to try to power through this all in one go.
I can't remember what unimodal_test.R was supposed to do :)

Removed old data files.

@@ -9,10 +9,11 @@ vals <- seq(0, 1, by=0.01)
 
 #  Sakari, I doubt it will, but you need to make sure that 'goring' isn't anywhere in your
 #  working directory path!
-analyst <- ifelse(regexpr('goring', tolower(getwd()))>0, 'Simon', 'Sakari')
+#analyst <- ifelse(regexpr('goring', tolower(getwd()))>0, 'Simon', 'Sakari')
+analyst <- 'Work'
 
 if(paste('longlist.', analyst, '.RData', sep='') %in% list.files('data')){
-  load(paste('data/longlist.', analyst, '.RData', sep=''))
+  load(paste0('data/longlist.', analyst, '.RData'))
 }
 
 if(!exists('longlist')){
@@ -29,11 +30,11 @@ if(exists('longlist')){
 save(longlist, file=paste('data/longlist.', analyst, '.RData', sep=''))
 
 library(snowfall)
-
 library(gbm)
 
 sfStop()
-sfInit(parallel = TRUE, cpus = 2)
+sfInit(parallel = TRUE, cpus = 3)
+
 
 subset.pol <- function(set){
   
@@ -52,11 +53,12 @@ subset.pol <- function(set){
 
 brt.run <- function(px, pa){
   
-  set <- keep.pol[px,]
+  set <- keep.pol
   
   x <- subset.pol(set)
   
   zeros <- colSums(x$calib.pol, na.rm = TRUE) == 0
+  bad.clim <- is.na(x$calib.clim)
   
   # Construct the brt formula based on x$calib.clim and x$calib.pol column 
   # names, omitting the zero-sum taxa which won't be used.
@@ -66,12 +68,14 @@ brt.run <- function(px, pa){
   # Construct brt model. x$calib.clim and x$calib.pol combined into a single
   # data frame and passed as the "data" parameter.
   
-  brt.model <- gbm(brt.formula, distribution="gaussian", n.trees=500, shrinkage=0.005, interaction.depth=6, cv.folds=4, verbose=FALSE, data=cbind(x$calib.clim, x$calib.pol[,!zeros]))
+  brt.model <- try(gbm(brt.formula, distribution="gaussian", 
+                   n.trees=500, shrinkage=0.005, interaction.depth=6, 
+                   cv.folds=4, verbose=FALSE, 
+                   data=cbind(x$calib.clim, x$calib.pol[,!zeros])[!bad.clim,]))
   
[…]

### 2015-05-06 R/unimodal_test.R

@@ -1,0 +1,31 @@
+install.packages('diptest')
+
+library(diptest)
+library(mgcv)
+
+dips <- matrix(ncol = ncol(new.pol)+1,
+               nrow = ncol(climate))
+
+for(i in 1:ncol(new.pol)){
+  
+  #  First row is to see whether the pollen data itself has a unimodal distribution
+  dips[1,i]    <- dip.test(new.pol[,i], simulate.p.value = TRUE)$p.value  
+  
+  for(j in 4:ncol(climate)){
+    
+    samples <- data.frame(y = new.pol[,i], x = climate[,j])
+    
+    mod.test <- gam(round(cbind(y*1000, 1000),0) ~ s(x, k = 10), data = samples, family = binomial)
+    output <- predict(mod.test, 
+                      newdata = data.frame(x = seq(min(samples$x, na.rm=TRUE), 
+                                                    max(samples$x, na.rm=TRUE), 
+                                                   length.out = 100)),
+                      type = 'response')
+    
+    dips[j,1]    <- dip.test(climate[,j], simulate.p.value = TRUE)$p.value  
+    dips[j, i+1] <- dip.test(output, simulate.p.value = TRUE)$p.value
+    
+    cat(ifelse(dips[j, i]<0.05, 't', 'f'))
+  }
+  cat(i, '\n')
+}

### 2015-05-06 Paper_Markdown.Rmd
I'm not sure what's wrog, but it looks fine to me.

--- 2014-07-15 Paper_Markdown.Rmd
@@
 MAT (Overpeck, 1985) is commonly applied to reconstruct climate, for pollen, diatoms, chironomids and other taxa.
-WA has a long history in the literature, and the recent development of monotone deshrinking has apparently improved performance further (Simpson, 2014?), but WA continues to suffer fro a basic assumption of unimodal taxon distributions along climate gradients.
+WA has a long history in the literature, and the recent development of monotone deshrinking has apparently improved performance further (Simpson, 2014?), but WA continues to suffer from a basic assumption of unimodal taxon distributions along climate gradients.
 

### 2015-05-06 R/data_setup.R
I've marked out the ggmap package which seems to have trouble installing in Linux.

@@ -9,9 +9,9 @@ library(plyr)
 library(rioja)
 library(randomForest)
 library(vegan)
-library(maps)
+#library(maps)
 library(maptools)
-library(ggmap)
+#library(ggmap)
 library(reshape2)
 library(analogue)
 

### 2015-05-06 R/bootstrap_mat.R
Didn't have the auto-load for MAT and WA.  Added it in now.

@@ -8,12 +8,17 @@ sfInit(parallel = TRUE, cpus = 30)
 #  didn't seem to add much to the understanding of what was going on.
 vals <- seq(0, 1, by=0.01)
 
-mat.res <- list(mean_prediction  = matrix(ncol=length(vals), nrow=nrow(new.pol)),
-                sample_size      = matrix(ncol=length(vals), nrow=nrow(new.pol)),
-                bias             = matrix(ncol=length(vals), nrow=nrow(new.pol)),
-                variance         = matrix(ncol=length(vals), nrow=nrow(new.pol)),
-                expectation      = matrix(ncol=length(vals), nrow=nrow(new.pol)))
+if('mat.res.RData' %in% list.files('data')){
+  load('data/mat.res.RData')
+} else {
+  
+  mat.res <- list(mean_prediction  = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                  sample_size      = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                  bias             = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                  variance         = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                  expectation      = matrix(ncol=length(vals), nrow=nrow(new.pol)))
 
+}  
 #  Turn the distance table into a matrix.
 diag.dist <- as.matrix(dists); diag(diag.dist) <- NA
 
@@ -80,9 +85,8 @@ for(i in i:length(vals)){
 
   for(j in 1:nrow(new.pol)){
 
-    prediction <- unlist(sfLapply(1:30, mat.fun, j = j, min = which.min(rmse)))
-    
     if(is.na(mat.res$mean_prediction[j,i])){
+      prediction <- unlist(sfLapply(1:30, mat.fun, j = j, min = which.min(rmse)))
       mat.res$mean_prediction[j,i] <- mean(prediction, na.rm=TRUE)
       mat.res$sample_size[j,i] <- sum(keep.pol[j,], na.rm=TRUE)
       mat.res$bias[j, i] <- (climate[j,10] - mean(prediction, na.rm=TRUE))^2

### 2015-05-06 R/bootstrap_wa.R

@@ -9,14 +9,14 @@ sfInit(parallel = TRUE, cpus = 30)
 #  didn't seem to add much to the understanding of what was going on.
 vals <- seq(0, 1, by=0.01)
 
-wa.res <- list(mean_prediction  = matrix(ncol=length(vals), nrow=nrow(new.pol)),
-               sample_size      = matrix(ncol=length(vals), nrow=nrow(new.pol)),
-               bias             = matrix(ncol=length(vals), nrow=nrow(new.pol)),
-               variance         = matrix(ncol=length(vals), nrow=nrow(new.pol)),
-               expectation      = matrix(ncol=length(vals), nrow=nrow(new.pol)))
-
-if('wa.res.RData' %in% list.files('data/')) {
+if('wa.res.RData' %in% list.files('data')){
   load('data/wa.res.RData')
+} else {
+  wa.res <- list(mean_prediction  = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                 sample_size      = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                 bias             = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                 variance         = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                 expectation      = matrix(ncol=length(vals), nrow=nrow(new.pol)))
 }
 
 i <- 1; j <- 1

### 2015-05-06 R/bootstrap_mat.R
Speeding up the MAT runs when data is already there.

@@ -71,32 +71,33 @@ sfExport(list = list('climate'))
 for(i in i:length(vals)){
   #  At each quantile, figure out which samples should be acceptable for a calibration set targeting
   #  each sample return 'keep.pol'
-  keep.pol <- aaply(diag.dist, 1, 
-                    function(x) {x > vals[i]})
 
-  diag(keep.pol) <- FALSE
-  
   if(any(is.na(mat.res$mean_prediction[,i]))){
+    keep.pol <- aaply(diag.dist, 1, 
+                      function(x) {x > vals[i]})
+    
+    diag(keep.pol) <- FALSE
+    
     fast.mat <- laply(1:nrow(diag.dist), fmat)
     rmse <- aaply(fast.mat, 2, function(x) sqrt(mean((x - climate[,10])^2, na.rm = T)))
-  }
   
-  sfExport(list = list('keep.pol'))
-
-  for(j in 1:nrow(new.pol)){
-
-    if(is.na(mat.res$mean_prediction[j,i])){
-      prediction <- unlist(sfLapply(1:30, mat.fun, j = j, min = which.min(rmse)))
-      mat.res$mean_prediction[j,i] <- mean(prediction, na.rm=TRUE)
-      mat.res$sample_size[j,i] <- sum(keep.pol[j,], na.rm=TRUE)
-      mat.res$bias[j, i] <- (climate[j,10] - mean(prediction, na.rm=TRUE))^2
-      mat.res$expectation[j, i]  <- mean((climate[j,10] - prediction)^2, na.rm=TRUE)
-      mat.res$variance[j, i]  <- mean((mean(prediction, na.rm=TRUE) - prediction)^2, na.rm=TRUE)
+    sfExport(list = list('keep.pol'))
+  
+    for(j in 1:nrow(new.pol)){
+  
+      if(is.na(mat.res$mean_prediction[j,i])){
+        prediction <- unlist(sfLapply(1:30, mat.fun, j = j, min = which.min(rmse)))
+        mat.res$mean_prediction[j,i] <- mean(prediction, na.rm=TRUE)
+        mat.res$sample_size[j,i] <- sum(keep.pol[j,], na.rm=TRUE)
+        mat.res$bias[j, i] <- (climate[j,10] - mean(prediction, na.rm=TRUE))^2
+        mat.res$expectation[j, i]  <- mean((climate[j,10] - prediction)^2, na.rm=TRUE)
[…]

### 2015-05-07 R/bootstrap_brt.R
Added try-catch for bootstrap_brt to see if that helps the errors on the ec2.

@@ -61,15 +61,19 @@ brt.run <- function(j){
   # Construct brt model. x$calib.clim and x$calib.pol combined into a single
   # data frame and passed as the "data" parameter.
   
-  brt.model <- gbm(brt.formula, distribution="gaussian", 
+  brt.model <- try(gbm(brt.formula, distribution="gaussian", 
                    n.trees=500, shrinkage=0.005, 
                    interaction.depth=6, cv.folds=4, 
                    verbose=FALSE, 
                    data=cbind(x$calib.clim, 
-                              x$calib.pol[,!zeros])[!bad.clim,])
+                              x$calib.pol[,!zeros])[!bad.clim,]))
 
-  # This probably always returns 500
-  brt.ntrees <- gbm.perf(brt.model, method="cv", plot.it=FALSE)   
+  if(!class(brt.model) == 'try-error'){
+    # This probably always returns 500
+    brt.ntrees <- gbm.perf(brt.model, method="cv", plot.it=FALSE)   
+  } else {
+    brt.model <- NA
+  }  
   
   if (length(brt.model) > 1){
     output <- predict(brt.model, new.pol[j,!zeros], brt.ntrees, type="response")

### 2015-05-07 R/bootstrap_mat.R
There was a problem with the assignment of 'i' here.

@@ -68,7 +68,7 @@ sfExport(list = list('mat.fun'))
 sfExport(list = list('new.pol'))
 sfExport(list = list('climate'))
 
-for(i in i:length(vals)){
+for(i in 1:length(vals)){
   #  At each quantile, figure out which samples should be acceptable for a calibration set targeting
   #  each sample return 'keep.pol'
 

### 2015-05-08 R/bootstrap_brt.R
Wrapping all the BRT stuff into `try` calls.

@@ -70,14 +70,20 @@ brt.run <- function(j){
 
   if(!class(brt.model) == 'try-error'){
     # This probably always returns 500
-    brt.ntrees <- gbm.perf(brt.model, method="cv", plot.it=FALSE)   
+    brt.ntrees <- try(gbm.perf(brt.model, method="cv", plot.it=FALSE))
+    if(class(brt.ntrees) == 'try-error'){
+      brt.model <- NA
+    }
   } else {
     brt.model <- NA
   }  
   
   if (length(brt.model) > 1){
-    output <- predict(brt.model, new.pol[j,!zeros], brt.ntrees, type="response")
+    output <- try(predict(brt.model, new.pol[j,!zeros], brt.ntrees, type="response"))
+    if(class(output) == 'try-error'){
+      output <- NA
     }
+  }
   else{
     output <- NA
   }

### 2015-05-08 R/bootstrap_brt.R
Dropped the number of cores to see if I can stop running out of memory.

@@ -11,7 +11,7 @@ library(snowfall)
 library(gbm)
 
 sfStop()
-sfInit(parallel = TRUE, cpus = 30)
+sfInit(parallel = TRUE, cpus = 15)
 
 vals <- seq(0, 1, by=0.01)
 

### 2015-05-20 Paper_Markdown.Rmd
Explicitly load the RData files into memory.  These ought to be rds files at some point, but there's too much intertia right now.
Other minor reformatting.

@@
 ```{r Fig4SampleLoss, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3, fig.width=6}
+load('data/mat.res.RData')
+load('data/wa.res.RData')
+load('data/wapls.res.RData')
+load('data/brt.res.RData')
+load('data/rfor.res.RData')
 
@@
                          size = as.vector(mat.res$sample_size),
-                         distance = rep(seq(0, 1, by=0.01), each = nrow(mat.res$sample_size)))
+                         distance = rep(seq(0, 1, by=0.01), 
+                                        each = nrow(mat.res$sample_size)))
 
@@
 loss <- ggplot(data = na.samples, aes(x=distance, y=size)) + 
-  geom_path(aes(group=site, color = bin)) +
+  geom_path(aes(group=site, color = bin), alpha = 0.4) +
   scale_color_brewer(palette = 'PiYG', type = 'div') +
@@
   geom_point(data = loc.loss, 
-             aes(x = long, y = lat, color = bin), alpha=0.7, size = 3) +
+             aes(x = long, y = lat, color = bin), alpha=0.4, size = 3) +
   theme_bw() +
@@
                         dissim = rep(rep(vals, each=nrow(wa.res$bias)),5),
-                        model  = rep(c('MAT', 'WA', 'WAPLS', 'rFor', 'BRT'), each = length(wa.res$bias)),
+                        model  = rep(c('MAT', 'WA', 'WAPLS', 'rFor', 'BRT'), 
+                                     each = length(wa.res$bias)),
                         init = rep(rep(climate[,10], 101), 5))

### 2015-06-15 R/bootstrap_brt.R
Changed the formula script, changed the number of CPUs for a local run (will change back).

@@ -11,7 +11,7 @@ library(snowfall)
 library(gbm)
 
 sfStop()
-sfInit(parallel = TRUE, cpus = 15)
+sfInit(parallel = TRUE, cpus = 2)
 
 vals <- seq(0, 1, by=0.01)
 
@@ -55,8 +55,14 @@ brt.run <- function(j){
   # Construct the brt formula based on x$calib.clim and x$calib.pol column 
   # names, omitting the zero-sum taxa which won't be used.
   
-  brt.formula <- as.formula(paste (colnames(x$calib.clim)[1], "~", 
-                                   paste(colnames(x$calib.pol)[!zeros], collapse=" + ")))
+  if(sum(!zeros) == 0){
+    stop('Whoops.')
+  }
+  
+  new.formula <- sprintf("%s ~ %s", colnames(x$calib.clim)[1], 
+                         paste(colnames(x$calib.pol)[!zeros], collapse=" + ")) 
+  
+  brt.formula <- as.formula(new.formula)
 
   # Construct brt model. x$calib.clim and x$calib.pol combined into a single
   # data frame and passed as the "data" parameter.
@@ -142,7 +148,7 @@ for(k in samp){
     
   if(is.na(brt.res$mean_prediction[j,i])){
 
-    prediction <- unlist(sfLapply(rep(j, 30), fun = brt.run))
+    prediction <- unlist(sfLapply(rep(j, 30), fun = function(z)try(brt.run(z))))
   
     brt.res$mean_prediction[j,i] <- mean(prediction, na.rm=TRUE)
     brt.res$sample_size[j,i] <- sum(keep.pol[j,], na.rm=TRUE)

### 2015-06-15 R/bootstrap_brt.R
Upped the number of CPUs again.

@@ -11,7 +11,7 @@ library(snowfall)
 library(gbm)
 
 sfStop()
-sfInit(parallel = TRUE, cpus = 2)
+sfInit(parallel = TRUE, cpus = 15)
 
 vals <- seq(0, 1, by=0.01)
 

### 2015-06-15 R/bootstrap_brt.R
There are cases where the BRT returns only a set of NAs because it fails (for whatever reason).  We put in 9999 as a placeholder (that we can strip out later) so that we don't wind up re-running the points over and over.

@@ -149,12 +149,19 @@ for(k in samp){
   if(is.na(brt.res$mean_prediction[j,i])){
 
     prediction <- unlist(sfLapply(rep(j, 30), fun = function(z)try(brt.run(z))))
-  
-    brt.res$mean_prediction[j,i] <- mean(prediction, na.rm=TRUE)
-    brt.res$sample_size[j,i] <- sum(keep.pol[j,], na.rm=TRUE)
-    brt.res$bias[j, i] <- (climate[j,10] - mean(prediction, na.rm=TRUE))^2
-    brt.res$expectation[j, i] <- mean((climate[j,10] - prediction)^2)
-    brt.res$variance[j, i] <- mean((mean(prediction) - prediction)^2)
+    if(!is.na(mean(prediction)){
+      brt.res$mean_prediction[j,i] <- mean(prediction, na.rm=TRUE)
+      brt.res$sample_size[j,i] <- sum(keep.pol[j,], na.rm=TRUE)
+      brt.res$bias[j, i] <- (climate[j,10] - mean(prediction, na.rm=TRUE))^2
+      brt.res$expectation[j, i] <- mean((climate[j,10] - prediction)^2)
+      brt.res$variance[j, i] <- mean((mean(prediction) - prediction)^2)
+    } else {
+      brt.res$mean_prediction[j,i] <- 9999
+      brt.res$sample_size[j,i] <- 9999
+      brt.res$bias[j, i] <- 9999
+      brt.res$expectation[j, i] <- 9999
+      brt.res$variance[j, i] <- 9999
+    }
   }
 
   save(brt.res, file = 'data/brt.res.RData')

### 2015-06-15 R/bootstrap_brt.R
Missed a bracket.

@@ -149,7 +149,8 @@ for(k in samp){
   if(is.na(brt.res$mean_prediction[j,i])){
 
     prediction <- unlist(sfLapply(rep(j, 30), fun = function(z)try(brt.run(z))))
-    if(!is.na(mean(prediction)){
+    
+    if(!is.na(mean(prediction))){
       brt.res$mean_prediction[j,i] <- mean(prediction, na.rm=TRUE)
       brt.res$sample_size[j,i] <- sum(keep.pol[j,], na.rm=TRUE)
       brt.res$bias[j, i] <- (climate[j,10] - mean(prediction, na.rm=TRUE))^2

### 2015-06-15 Paper_Markdown.Rmd
Working on cleaning things up for writing.  Moved figure 3 over to a function in its own file.

@@
 ---
-output: word_document
+output: html_document
 ---
@@
 The site is notable for very high proportions of *Fraxinus pennsylvanica/americana* pollen, more than 5x higher than the next closest site.
-The lowest minimum distance occurs across four sites, two from Thinn Lake, identified as potentially anomalous in Williams et al (2009),
+The lowest minimum distance occurs across four sites, two from Thinn Lake, identified as potentially anomalous in Williams et al (2009) and XXXX.
+The geographic distribution of non-analogue sites is particularly interesting since many of the sites are in close proximity to other modern sites.
+This indicates that While spatial autocorreltaion is likely to play a role in the proximity of analogues (and thus their predictions) [@telfordSpatialAuto] it is not always the case that spatially proximate samples are compositionally proximate.
 
@@
 
-mds <- data.frame(pol.mds$points,
-                  non.ana = factor(min.diss > 0.48))
-
-non.ana <- ggplot(data = mds, aes(x = X1, y = X2)) +
-  geom_point(data = mds[mds$non.ana == FALSE,], alpha=0.5, size = 2, color='black') +   
-  geom_point(data = mds[mds$non.ana == TRUE,], alpha=0.7, size = 4, color='red') +
-  theme_bw() +
-  xlab('MDS Axis One') + ylab('MDS Axis Two') +
-  theme(text = element_text(size=24, family='serif'))
-
-non.ana.map <- ggplot(data = data.frame(map), aes(long, lat)) + 
-  geom_polygon(aes(group=group), color='blue', alpha=0.2) +
-  geom_point(data = data.frame(climate)[mds$non.ana == FALSE,], 
-             aes(x = LONDD, y = LATDD), alpha=0.5) +
-  geom_point(data = data.frame(climate)[mds$non.ana == TRUE,], 
-             aes(x = LONDD, y = LATDD), color = 'red', size = 3) +
-   theme_bw() +
-  xlab('Longitude') + ylab('Latitude') +
-  theme(axis.text=element_text(size=12, family='serif'),
-        panel.background = element_rect(fill = 'lightblue')) +
-  coord_map(projection='albers', lat0=35, lat1=55, xlim=c(-100, -60), ylim=c(35, 80))
-
-grid.arrange(non.ana, non.ana.map, ncol=2)
+source('R/figures/figure3.R')
+run_figure3(pol.mds, min.diss, map, climate)
 
@@
   scale_y_continuous(expand=c(0,0), limits=c(0, 40)) +
-  scale_color_continuous(low = 'blue', high='pink', trans='sqrt') +
+  #scale_color_continuous(low = 'blue', high='pink', trans='sqrt') +
   theme(legend.position = 'none',

### 2015-06-15 R/figures/figure3.R

@@ -1,0 +1,26 @@
+run_figure3 <- function(pol_mds, min.diss, map, climate){
+  
+  mds <- data.frame(pol_mds$points,
+                    non.ana = factor(min.diss > 0.48))
+  
+  non.ana <- ggplot(data = mds, aes(x = X1, y = X2)) +
+    geom_point(data = mds[mds$non.ana == FALSE,], alpha=0.5, size = 2, color='black') +   
+    geom_point(data = mds[mds$non.ana == TRUE,], alpha=0.7, size = 4, color='red') +
+    theme_bw() +
+    xlab('MDS Axis One') + ylab('MDS Axis Two') +
+    theme(text = element_text(size=24, family='serif'))
+  
+  non.ana.map <- ggplot(data = data.frame(map), aes(long, lat)) + 
+    geom_polygon(aes(group=group), color='blue', alpha=0.2) +
+    geom_point(data = data.frame(climate)[mds$non.ana == FALSE,], 
+               aes(x = LONDD, y = LATDD), alpha=0.5) +
+    geom_point(data = data.frame(climate)[mds$non.ana == TRUE,], 
+               aes(x = LONDD, y = LATDD), color = 'red', size = 3) +
+    theme_bw() +
+    xlab('Longitude') + ylab('Latitude') +
+    theme(axis.text=element_text(size=12, family='serif'),
+          panel.background = element_rect(fill = 'lightblue')) +
+    coord_map(projection='albers', lat0=35, lat1=55, xlim=c(-100, -60), ylim=c(35, 80))
+  
+  grid.arrange(non.ana, non.ana.map, ncol=2)
+}

### 2015-07-21 Paper_Markdown.Rmd
Adding spaces after the plots to make plotting look better.

@@
 ---
-output: html_document
+output: pdf_document
 ---
@@
 ```
+
 **Figure 1**. *Location of sample plots used for analysis.  Samples are derived from a modern pollen dataset.*
@@
 ```
+
 **Figure 2**.
@@
 ```
+
 **Figure 3**.
@@
 
-Key variables based on the $latex \frac {\lambda_1} {\lambda_2} $ using pCCA appear to be climate variables relating to mean monthly percent summer sunshine (variables preceded by an 's').
+Key variables based on the $\frac{\lambda_1}{\lambda_2}$ using pCCA appear to be climate variables relating to mean monthly percent summer sunshine (variables preceded by an 's').
 These variables are likely to be variables that are combinations of temperature and preciptiation variables, however this value is not linear, nor is it continuous, meaning that .
@@
 ```
+
 **Figure 4**.
@@
 To some degree this is surprising, however, the dominance of Pinus and Picea in these forests likely means that they have smaller analogue distances from one another than samples in the southern forests (with more diversity and therfore, likely fewer close analogues), but also, surpisingly, arctic samples also have higer inter-sample distances, and are thus less sensitive to losses of close analogues.
+```{r}
+
+get_analoguegradient <- function(dist){
+
+  x <- findInterval(dist, vals)
+
+  climate_stuff <- data.frame(tjul = c(climate$tjul, 
+                                     brt.res$mean_prediction[,x],
+                                     mat.res$mean_prediction[,x],
+                                     rfor.res$mean_prediction[,x],
+                                     wa.res$mean_prediction[,x],
+                                     wapls.res$mean_prediction[,x]),
+                            class = factor(c(rep('Actual', nrow(climate)),
+                                             rep(c('BRT', 'MAT', 'rFor', 'WA', 'WAPLS'),
+                                                 each = nrow(mat.res$mean_prediction)))),
+                            dist = dist)
+
+}
+
+dist_dists <- na.omit(do.call(rbind.data.frame,lapply(c(0, .5, 1), get_analoguegradient)))
+
+density_plot <- ggplot(dist_dists, aes(x = tjul, color = class, size = class, fill = class)) + 
+  geom_density() +
+  scale_size_manual(values = c(2, 1, 1, 1, 1, 1)) +
+  scale_fill_manual(values = c('#d3d3d3', NA, NA, NA, NA, NA)) +
+  theme_bw() +
+  coord_cartesian(xlim=c(-4.2, 28.8)) +
+  facet_wrap(~dist, ncol=1, scales='free_y')
+
+density_plot
+
+```
+**Figure 5**.
+*Climate predictions for July temperature using the full data set at an analogue distance of 0, 0.5 and 1.
+These results indicate that most predictive methods capture the full environmental gradient, however WA and WAPLS appear to over-emphasise the middle of the environmental gradient, BRT appears to truncate the gradient at both high and low ends.
+As expected, MAT appears to capture the true gradient best when the full data set is used, but rapidly regresses to the mean as distance to the nearest neighbor increases.*
 
@@
 ```
+
 **Figure 5**.
@@
 ```
+
 **Figure 6**. *Change in model bias with increasing analogue distance.*
@@
 ```
+
 **Figure 7**. *Change in bootstrap variance with increasing analogue distance.*
@@
 What does this tell us about predicting climate in non-analogue space?
+
+Key points:  
+Most predictive methods are stable until the no-analogue threshold is met.
+Beyond this threshold the collapse of most methods is (surprisingly) rapid, in part due to the fact that the number of samples available for the calibration function drops substantially, indicating high sensitivity to the sample neighborhood for each of these methods.
+
+* WAPLS can't be trusted given its bizarre behaviour beyond analogue space.
+
+* MAT is incredibly dependent on finding close neighbors.
+As analogue distance increases to near neighbours the likelihood of having poor fit increases substantially.
+This is manifested in the apparent 'flip' in predictions.
+We don't get regression to the mean, we get regression to the nearest available sample.
+* BRT seems most robust, but for the purposes of this simulation (in lage part) it is mind-bogglingly slow, even when compared to rForest, howver (1) for a single run this is unlikely to be an issue (single runs can take somewhere between 1 and 10 minutes) and (2) interstingly BRT predictions are the most constrained, i.e., the predictive output has a much narrower range of predictions than the underlying data set.
 

### 2015-07-22 Paper_Markdown.Rmd
Added new Ladd paper to the bibliography.  Formatted a few references so thay'll work properly.

--- 2015-05-20 Paper_Markdown.Rmd
@@
 =========================
-Pollen assemblages are widely used for paleoclimatic reconstruction during the Holocene, but this method is also applied to pollen assemblages from the last Glacial period (bartlein2011pollen, kaufman2012multi), and earlier (tarasov2011progress, brewer2008climate, cheddadi1998climate).
+Pollen assemblages are widely used for paleoclimatic reconstruction during the Holocene, but this method is also applied to pollen assemblages from the last Glacial period [@bartlein2011pollen; kaufman2012multi], and earlier [@tarasov2011progress;@brewer2008climate;@cheddadi1998climate].
 A particular challenge for climate reconstruction using these much earlier samples is the presence of non-analogue vegetation.
-Non-analogue pollen assemblages pose a problem for many reconstruction techniques, these vegetation assemblages may be the result of sets of climate variables that no longer co-exist, ecological conditions resulting from differing rates of species migration () or release from herbivores (gill2009pleistocene), or as a result of changes in human land use (Goring et al.
+Non-analogue pollen assemblages pose a problem for many reconstruction techniques, these vegetation assemblages may be the result of sets of climate variables that no longer co-exist, ecological conditions resulting from differing rates of species migration () or release from herbivores [@gill2009pleistocene], or as a result of changes in human land use (Goring et al.
 in review?).
@@
 MAT (Overpeck, 1985) is commonly applied to reconstruct climate, for pollen, diatoms, chironomids and other taxa.
-WA has a long history in the literature, and the recent development of monotone deshrinking has apparently improved performance further (Simpson, 2014?), but WA continues to suffer from a basic assumption of unimodal taxon distributions along climate gradients.
+WA has a long history in the literature, and the recent development of monotone deshrinking has apparently improved performance further (Simpson, 2014?), but WA continues to suffer fro a basic assumption of unimodal taxon distributions along climate gradients.
 
@@
 ```{r Fig4SampleLoss, echo=FALSE, message=FALSE, warning=FALSE, fig.height=3, fig.width=6}
-load('data/mat.res.RData')
-load('data/wa.res.RData')
-load('data/wapls.res.RData')
-load('data/brt.res.RData')
-load('data/rfor.res.RData')
 
@@
                          size = as.vector(mat.res$sample_size),
-                         distance = rep(seq(0, 1, by=0.01), 
-                                        each = nrow(mat.res$sample_size)))
+                         distance = rep(seq(0, 1, by=0.01), each = nrow(mat.res$sample_size)))
 
@@
 loss <- ggplot(data = na.samples, aes(x=distance, y=size)) + 
-  geom_path(aes(group=site, color = bin), alpha = 0.4) +
+  geom_path(aes(group=site, color = bin)) +
   scale_color_brewer(palette = 'PiYG', type = 'div') +
@@
   geom_point(data = loc.loss, 
-             aes(x = long, y = lat, color = bin), alpha=0.4, size = 3) +
+             aes(x = long, y = lat, color = bin), alpha=0.7, size = 3) +
   theme_bw() +
@@
                         dissim = rep(rep(vals, each=nrow(wa.res$bias)),5),
-                        model  = rep(c('MAT', 'WA', 'WAPLS', 'rFor', 'BRT'), 
-                                     each = length(wa.res$bias)),
+                        model  = rep(c('MAT', 'WA', 'WAPLS', 'rFor', 'BRT'), each = length(wa.res$bias)),
                         init = rep(rep(climate[,10], 101), 5))

### 2015-07-23 Paper_Markdown.Rmd
Modified the paper header so that it can compile the bibliography.  Fixed one chunk so that it doesn't display the raw code.
Added explicit references in the text and into the bibliography.

Added a csl formatting file.

--- 2015-07-21 Paper_Markdown.Rmd
@@
 ---
-output: pdf_document
+title: "Non-analogues in paleoecological reconstruction - the effects of analogue distance"
+author: "Simon Goring *et al*."
+date: "`r format(Sys.time(), '%d %B, %Y')`"
+output:
+  md_document:
+    variant: markdown
+  pdf_document:
+    pandoc_args: "-V geometry:vmargin=1in -V geometry:hmargin=1in"
+bibliography: non-analogue_clim.bib
+csl: elsevier-harvard.csl
 ---
@@
 ========================================================
+
+<!--
+
+```{r setup-knitr, echo = FALSE, message = FALSE}
+# Simon: Note that these are global options that apply to all chunks below.
+
+knitr::opts_chunk$set(
+  comment = " ",
+  error = FALSE,
+  # Note: I have set EVAL to false just to make the bib work. 
+  # Once the bib is fixed, you'll set this to TRUE or just delete this line.
+  eval = TRUE,
+  cache = TRUE,
+  tidy = FALSE,
+  fig.path = "figure/",
+  cache.path = "cache/"
+)
+
+library("pander")
+library("plyr")
+panderOptions('table.style', 'rmarkdown')
+panderOptions('table.split.table', 300)
+```
+-->
+
+Simon Goring^1^
+
+J Sakari Salonen^2^
+
+Miska Luoto^2,3^
+
+Jack Williams^1,4^
+
+^1^ Department of Geography, University of Wisconsin-Madison, Madison, WI, USA
+
+^2^ ?????????
+
+^3^
+
+^4^ Center for Climatic Research, University of Wisconsin-Madison, Madison, WI, USA
 
@@
 =========================
-Pollen assemblages are widely used for paleoclimatic reconstruction during the Holocene, but this method is also applied to pollen assemblages from the last Glacial period (bartlein2011pollen, kaufman2012multi), and earlier (tarasov2011progress, brewer2008climate, cheddadi1998climate).
+Pollen assemblages are widely used for paleoclimatic reconstruction during the Holocene, but this method is also applied to pollen assemblages from the last Glacial period [@bartlein2011pollen, @kaufman2012multi], and earlier [@tarasov2011progress; @brewer2008climate, @cheddadi1998climate].
 A particular challenge for climate reconstruction using these much earlier samples is the presence of non-analogue vegetation.
-Non-analogue pollen assemblages pose a problem for many reconstruction techniques, these vegetation assemblages may be the result of sets of climate variables that no longer co-exist, ecological conditions resulting from differing rates of species migration () or release from herbivores (gill2009pleistocene), or as a result of changes in human land use (Goring et al.
-in review?).
+Non-analogue pollen assemblages pose a problem for many reconstruction techniques, these vegetation assemblages may be the result of sets of climate variables that no longer co-exist, ecological conditions resulting from differing rates of species migration (REF) or release from herbivores [@gill2009pleistocene], or as a result of changes in human land use [@goring2015preset].
 
@@
 
-Although we cannot know the values for temperature or precipitation during this time, statistical non-analogues do exist in a modern context, and we can construct false non-analogues using a modified h-block sampling procedure (e.g., Telford & Birks) that examines analogue distance rather than geographical distance.
+Although we cannot know the values for temperature or precipitation during this time, statistical non-analogues do exist in a modern context, and we can construct false non-analogues using a modified h-block sampling procedure [*sensu*, @telford2009evaluation] that examines analogue distance rather than geographical distance.
 
@@
 Sites from across North America were used to build a modern analogue dataset.
-Samples were obtained from the North American Modern Pollen Database east of 100oW.
+Samples were obtained from the North American Modern Pollen Database east of 100^o^W.
 Pollen taxa were standardized as in Gill et al.
-(in prep), using a set of 54 taxa, similar to those reported in Whitmore et al.
-(2005).
+[@gill2015noanalogues], using a set of 54 taxa, similar to those reported in Whitmore et al.
+[-@whitmore2005modern].
 We chose to exclude Western North America since pollen taxa such as *Tsuga canadensis* and *Tsuga heterophylla* may be morphologically indistinct, but their parent plant taxa have clearly differentiated distributions, thus it may be possible to find false close-analogues using the entire North American dataset.
@@
 Each of these models has been used previously in the pollen literature, and has strengths and weaknesses.
-MAT (Overpeck, 1985) is commonly applied to reconstruct climate, for pollen, diatoms, chironomids and other taxa.
+MAT [@overpeck1985quantitative] is commonly applied to reconstruct climate, for pollen, diatoms, chironomids and other taxa.
 WA has a long history in the literature, and the recent development of monotone deshrinking has apparently improved performance further (Simpson, 2014?), but WA continues to suffer from a basic assumption of unimodal taxon distributions along climate gradients.
+
+WAPLS does this and that. . .
+
+SAKARI or MISKA can you add text here about BRT?
 
@@
 The site is notable for very high proportions of *Fraxinus pennsylvanica/americana* pollen, more than 5x higher than the next closest site.
-The lowest minimum distance occurs across four sites, two from Thinn Lake, identified as potentially anomalous in Williams et al (2009) and XXXX.
+The lowest minimum distance occurs across four sites, two from Thinn Lake, identified as potentially anomalous in Williams et al [-@williams2009rapid] and XXXX.
 The geographic distribution of non-analogue sites is particularly interesting since many of the sites are in close proximity to other modern sites.
-This indicates that While spatial autocorreltaion is likely to play a role in the proximity of analogues (and thus their predictions) [@telfordSpatialAuto] it is not always the case that spatially proximate samples are compositionally proximate.
+This indicates that While spatial autocorreltaion is likely to play a role in the proximity of analogues (and thus their predictions) [@telford2009evaluation] it is not always the case that spatially proximate samples are compositionally proximate, and indeed, pollen assemblages in densely sampled regions may exhibit no-analogue characteristics.
 
@@
 To some degree this is surprising, however, the dominance of Pinus and Picea in these forests likely means that they have smaller analogue distances from one another than samples in the southern forests (with more diversity and therfore, likely fewer close analogues), but also, surpisingly, arctic samples also have higer inter-sample distances, and are thus less sensitive to losses of close analogues.
-```{r}
+```{r, , echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.height=5}
 

### 2015-07-23 Paper_Markdown.Rmd
Removed the `md_document` flag in the output header.
Knit to PDF for most recent update.

@@
 output:
-  md_document:
-    variant: markdown
   pdf_document:

### 2015-07-24 Paper_Markdown.Rmd
Added pandoc/knit header.

--- 2015-07-22 Paper_Markdown.Rmd
@@
 ---
-output: pdf_document
+title: "Non-analogues in paleoecological reconstruction"
+author: "Simon Goring *et al*."
+date: '`r format(Sys.time(), ''%d %B, %Y'')`'
+output:
+  pdf_document:
+    pandoc_args: -V geometry:vmargin=1in -V geometry:hmargin=1in
+  md_document:
+    variant: markdown_github
+  word_document:
+    fig_height: 6
+    fig_width: 6
+csl: ecology.csl
+pandoc_args: -H margins.sty
+bibliography: non-analogue_clim.bib
 ---

### 2015-07-24 Paper_Markdown.Rmd

--- 2015-07-23 Paper_Markdown.Rmd
@@
 ---
-title: "Non-analogues in paleoecological reconstruction - the effects of analogue distance"
+title: "Non-analogues in paleoecological reconstruction"
 author: "Simon Goring *et al*."
@@
   pdf_document:
-    pandoc_args: "-V geometry:vmargin=1in -V geometry:hmargin=1in"
+    pandoc_args: -V geometry:vmargin=1in -V geometry:hmargin=1in
+  md_document:
+    variant: markdown_github
+  word_document:
+    fig_height: 6
+    fig_width: 6
+csl: elsevier-harvard.csl
+pandoc_args: -H margins.sty
 bibliography: non-analogue_clim.bib
-csl: elsevier-harvard.csl
 ---
@@
 
-Miska Luoto^2,3^
+Miska Luoto^2^
 
-Jack Williams^1,4^
+Jack Williams^1,3^
 
@@
 
-^2^ ?????????
+^2^ Department of Geosciences and Geography, University of Helsinki, Helsinki, Finland
 
-^3^
-
-^4^ Center for Climatic Research, University of Wisconsin-Madison, Madison, WI, USA
+^3^ Center for Climatic Research, University of Wisconsin-Madison, Madison, WI, USA
 
@@
 =========================
+
 Pollen assemblages are widely used for paleoclimatic reconstruction during the Holocene, but this method is also applied to pollen assemblages from the last Glacial period [@bartlein2011pollen, @kaufman2012multi], and earlier [@tarasov2011progress; @brewer2008climate, @cheddadi1998climate].
@@
 To some degree this is surprising, however, the dominance of Pinus and Picea in these forests likely means that they have smaller analogue distances from one another than samples in the southern forests (with more diversity and therfore, likely fewer close analogues), but also, surpisingly, arctic samples also have higer inter-sample distances, and are thus less sensitive to losses of close analogues.
+
 ```{r, , echo=FALSE, message=FALSE, warning=FALSE, fig.width=6, fig.height=5}

### 2015-07-30 Paper_Markdown.Rmd
Minor changes from Sakari's comments on the PDF.  Added some methods papers.  Will add more.

@@
 To analyse the effects of non-analogue pollen on model skill we chose five pollen-based reconstruction methods from the literature and apply them to the modern pollen data from eastern North America.
-The models are MAT, WA, WAPLS, BRT and randomForest.
+The models are MAT [@overpeck1985quantitative], WA [@ter1993weighted], WAPLS [@ter1989inferring], BRT and randomForest.
 Each of these models has been used previously in the pollen literature, and has strengths and weaknesses.
-MAT [@overpeck1985quantitative] is commonly applied to reconstruct climate, for pollen, diatoms, chironomids and other taxa.
+MAT is commonly applied to reconstruct climate, for pollen, diatoms, chironomids and other taxa.
 WA has a long history in the literature, and the recent development of monotone deshrinking has apparently improved performance further (Simpson, 2014?), but WA continues to suffer from a basic assumption of unimodal taxon distributions along climate gradients.
@@
 
-SAKARI or MISKA can you add text here about BRT?
+BRT [@de2007boosted] and randomForest [@breiman2001random] are ensemble machine-learning approaches which have seen increasing use in fossil pollen literature during the past few years [][SOME REFS HERE IF NECESSARY ...
+GORING ET AL.
+2010, ZHAO ET AL.
+2012, SALONEN ET AL.
+2014...].
+Both methods are based on creating an ensemble of regression tree models with all pollen taxa used as predictors and an environmental variable of choice as the response.
+The randomForest creates an ensemble of independent tree models (here 500 trees), while using bootstrap samples of the calibration data when building each tree, and randomly selected subsets of the predictor set when determining each tree split, thus creating diversity in the tree ensemble.
+In BRT, by comparison, the tree ensemble is built sequentially, with each added tree aiming to explain the residuals of the previously fitted ensemble.
+The final BRT model can thus be likened to an additive regression model in which the individual terms are regression trees.
 
@@
 
-Mean minimum dissimilarity for pollen samples in the modern pollen dataset ranges from `r min(min.diss)` to `r min(min.diss)`, with a 95% confidence interval from `r quantile(min.diss, 0.025)` to `r quantile(min.diss, 0.975)`.
+Mean minimum dissimilarity for pollen samples in the modern pollen dataset ranges from `r min(min.diss)` to `r max(min.diss)`, with a 95% confidence interval from `r quantile(min.diss, 0.025)` to `r quantile(min.diss, 0.975)`.
 The highest minimum analogue distance is for an unpublished site along the New York, Pennsylvannia border.
@@
 The geographic distribution of non-analogue sites is particularly interesting since many of the sites are in close proximity to other modern sites.
-This indicates that While spatial autocorreltaion is likely to play a role in the proximity of analogues (and thus their predictions) [@telford2009evaluation] it is not always the case that spatially proximate samples are compositionally proximate, and indeed, pollen assemblages in densely sampled regions may exhibit no-analogue characteristics.
+This indicates that While spatial autocorrelation is likely to play a role in the proximity of analogues (and thus their predictions) [@telford2009evaluation] it is not always the case that spatially proximate samples are compositionally proximate, and indeed, pollen assemblages in densely sampled regions may exhibit no-analogue characteristics.
 

### 2015-08-27 Paper_Markdown.Rmd

@@
 A particular challenge for climate reconstruction using these much earlier samples is the presence of non-analogue vegetation.
-Non-analogue pollen assemblages pose a problem for many reconstruction techniques, these vegetation assemblages may be the result of sets of climate variables that no longer co-exist, ecological conditions resulting from differing rates of species migration (REF) or release from herbivores [@gill2009pleistocene], or as a result of changes in human land use [@goring2015preset].
+Non-analogue pollen assemblages pose a problem for many reconstruction techniques, these vegetation assemblages may be the result of sets of climate variables that no longer co-exist, ecological conditions resulting from differing rates of species migration [@huntley1991plants] or release from herbivores [@gill2009pleistocene], or as a result of changes in human land use [@goring2015preset].
 
@@
 The final BRT model can thus be likened to an additive regression model in which the individual terms are regression trees.
+
+For each pollen model and each sample site we calibrated a model to reconstruct T_{jul} using a pool of samples selected by removing all pollen samples closer than an analogue distance that increased from 0 to 1 at a step size of 0.01.
+The calibration dataset for each model and each sample, at each exclusion step was then bootstrapped 30 times, sampling with replacement from the available calibration samples, to obtain a variance parameter, along with a mean prediction, bias estimate, root mean squared error of prediction and sample size estimate.
+This full procedure reqired 29,254,650 individual calibrated models (although a number failed due to small sample sizes with large analogue exclusion distances).
+To facilitate this large number of models, and the significant processing time, we ran the simulations on an Amazon AWS EC2 server as a parallel process using 30 cores.
 

### 2015-09-17 Paper_Markdown.Rmd
Lots of new text in the introduction.  Moved many plots into their own files (in R/figures).  Added some methods text.
New references.

Cut libraries that were unused.

Added a 'null' model.  Just randomly samples from the data without respect to underlying data.

@@
 Although we cannot know the values for temperature or precipitation during this time, statistical non-analogues do exist in a modern context, and we can construct false non-analogues using a modified h-block sampling procedure [*sensu*, @telford2009evaluation] that examines analogue distance rather than geographical distance.
+The result of no-analogue prediction is likely to vary with the method used for climate reconstruction, and possibly the climate variable (or distribution of the climate variable) of interest.
+This has been shown to occur with reconstructions from mixed depositional environments [@goring2010pollen], and more broadly, related families of methods are likely to behave similarly [KNUTTIPaper, NEWCANNONPAPER].
+
+The Modern Analogue Technique is effectively a k-nearest neighbors method, generally using a squared-chord metric of distance.
+MAT is used commonly in paleoecological reconstruction [@overpeck1985quantitative], but the k-nearest neighbor alogrithm is common to big data analysis.
+With MAT, the effects of moving toward no-analogue space will likely vary depending on the location of the sample in multi-dimentsional space.
+We expect that a sample close to the average of all samples will show increasing predictive variability, but little effect on the mean prediction as the closest analogues become progressvely further from the focal sample, but the average of distant samples remains similar.
+In general, with increasing analogue distance we expect samples should revert toward the mean, while also showing higher variability.
+Of all the methods, MAT's dependence on the local neighborhood likely suggests that we should expect a higher increase in variance with increasing analogue distance.
+
+When WA is used for reconstruction in no-analogue space it is likely that species combinations found along the climatic gradient of interest are absent, meaning higher variability of prediction and greater likelihood that species optima or tolerances will be misinterpreted.
+This is likely to be most significant for taxa with bimodal distributions,
 
@@
 
-map <- map_data('world')
+source('R/figures/figure1_map.R')
 
-map <- subset(map, map$long > -160 & map$long < -30)
-map <- subset(map, map$lat > 20 & map$lat < 83)
-
-ggplot(data = data.frame(map), aes(long, lat)) + 
-  geom_polygon(aes(group=group), color='blue', alpha=0.7) +
-  geom_point(data = data.frame(climate), aes(x = LONDD, y = LATDD), alpha = 0.5) +
-  theme_bw() +
-  xlab('') + ylab('') +
-  theme(axis.text=element_text(size=12, family='serif'),
-        panel.background = element_rect(fill = 'lightblue')) +
-  coord_map(projection='albers', lat0=35, lat1=55, xlim=c(-100, -60), ylim=c(35, 80))
+map_plot
 
@@
 
-WAPLS does this and that. . .
+WAPLS is a modification of WA, whereby the residuals for each component are again weighted and regressed and the number of components are chosen by minimizing the root mean squared error (RMSE) of the model [@ter1993weighted].
 
-BRT [@de2007boosted] and randomForest [@breiman2001random] are ensemble machine-learning approaches which have seen increasing use in fossil pollen literature during the past few years [][SOME REFS HERE IF NECESSARY ...
-GORING ET AL.
-2010, ZHAO ET AL.
-2012, SALONEN ET AL.
-2014...].
-Both methods are based on creating an ensemble of regression tree models with all pollen taxa used as predictors and an environmental variable of choice as the response.
-The randomForest creates an ensemble of independent tree models (here 500 trees), while using bootstrap samples of the calibration data when building each tree, and randomly selected subsets of the predictor set when determining each tree split, thus creating diversity in the tree ensemble.
+BRT [@de2007boosted] and randomForest [@breiman2001random] are ensemble machine-learning approaches which have seen increasing use in fossil pollen literature during the past few years [@goring2010pollen;@salonen2014reconstructing;@salonen2012north;@zhao2012application].
+Both machine learning methods are based on creating an ensemble of regression tree models with all pollen taxa used as predictors and an environmental variable of choice as the response.
+The random forest creates an ensemble of independent tree models (here 500 trees), while using bootstrap samples of the calibration data when building each tree, and randomly selected subsets of the predictor set when determining each tree split, thus creating diversity in the tree ensemble.
 In BRT, by comparison, the tree ensemble is built sequentially, with each added tree aiming to explain the residuals of the previously fitted ensemble.

### 2015-09-17 R/bootstrap_rand.R

@@ -1,0 +1,87 @@
+library(snowfall)
+
+sfStop()
+sfInit(parallel = TRUE, cpus = 2)
+
+# Set up the analogue distance exclusion values.
+#  This was originally a set of quantiles, but it played havoc on plotting and
+#  didn't seem to add much to the understanding of what was going on.
+vals <- seq(0, 1, by=0.01)
+
+if('rand.res.RData' %in% list.files('data')){
+  load('data/rand.res.RData')
+} else {
+  
+  rand.res <- list(mean_prediction  = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                   sample_size      = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                   bias             = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                   variance         = matrix(ncol=length(vals), nrow=nrow(new.pol)),
+                   expectation      = matrix(ncol=length(vals), nrow=nrow(new.pol)))
+
+}  
+
+
+i <- 1; j <- 1
+
+subset.pol <- function(set){
+  
+  calib.samples <- sample((1:length(set))[set], replace=TRUE)
+  
+  zeros <- colSums(new.pol[calib.samples,], na.rm = TRUE) == 0
+  
+  list(calib.pol  = new.pol[calib.samples,!zeros],
+       calib.clim = climate[calib.samples,10],
+       zeros = zeros)
+  
+}
+
+rand.run <- function(j){
+  
+  set <- keep.pol[j,]
+  
+  x <- subset.pol(set)
+  
+  bad.clim <- is.na(x$calib.clim)
+  
+  rand <- sample(x$calib.clim[!bad.clim],1)
+  
+  output <- rand
+  
+  return(output)
+  
+}
+
+#  Parallelize:
+sfExport(list = list('rand.run'))
+sfExport(list = list('new.pol'))
+sfExport(list = list('subset.pol'))
+sfExport(list = list('climate'))
+sfLibrary(rioja)
+
+
+for(i in 1:length(vals)){
+  #  This runs through each analogue distance
+  keep.pol <- aaply(diag.dist, 1, 
+                    function(x) {x > vals[i]})
+  diag(keep.pol) <- FALSE
+  
+  sfExport(list = list('keep.pol'))
+  
+  for(j in 1:nrow(new.pol)){
+    
[…]

### 2015-09-17 R/data_setup.R

@@ -9,11 +9,10 @@ library(plyr)
 library(rioja)
 library(randomForest)
 library(vegan)
-#library(maps)
 library(maptools)
-#library(ggmap)
 library(reshape2)
 library(analogue)
+library(RColorBrewer)
 
 climate <- read.csv('data/Clim_Pollen_Zirbel.csv', row.names=1)
 mod_pol <- read.csv('data/ModP35c_East.csv', row.names=1, sep='\t')

### 2015-09-17 R/figures/figure1_map.R

@@ -1,0 +1,16 @@
+# Figure 1 - Map
+
+
+map <- map_data('world')
+
+map <- subset(map, map$long > -160 & map$long < -30)
+map <- subset(map, map$lat > 20 & map$lat < 83)
+
+map_plot <- ggplot(data = data.frame(map), aes(x, y)) + 
+  geom_polygon(aes(group=group), color='blue', alpha=0.7) +
+  geom_point(data = data.frame(climate), aes(x = x, y = y), alpha = 0.5) +
+  theme_bw() +
+  xlab('') + ylab('') +
+  theme(axis.text=element_text(size=12, family='serif'),
+        panel.background = element_rect(fill = 'lightblue')) +
+  coord_map(projection='albers', lat0=35, lat1=55, xlim=c(-100, -60), ylim=c(35, 80))

### 2015-09-21 Paper_Markdown.Rmd
Added the null model (random selection) just to check & see.
Working on code to provide context for the bigger dataset (bound historical & modern) (`neotoma_historical.R`).

Added code for CI2015 figures in `best_fit_model.R`, will likely go into the paper, but right now I'm still figuring out the best way to get it to work.

Added this historical pollen records as `RData`, but this reflects anolder version of the Neotoma dataset, there are more newer records that could be added.

@@
 load('data/rfor.res.RData')
+load('data/rand.res.RData')
 
@@
 
-model.out <- data.frame(site   = rep(rep(climate[,1], ncol(wa.res$bias)),5),
-                        preds  = c(as.vector(mat.res$mean_prediction),
+model.out <- data.frame(site   = rep(rep(climate[,1], ncol(wa.res$bias)),6),
+                        preds  = c(as.vector(rand.res$mean_prediction),
+                                   as.vector(mat.res$mean_prediction),
                                    as.vector(wa.res$mean_prediction),
@@
                                    as.vector(brt.res$mean_prediction)),
-                        bias   = c(as.vector(mat.res$bias),
+                        bias   = c(as.vector(rand.res$bias),
+                                   as.vector(mat.res$bias),
                                    as.vector(wa.res$bias),
@@
                                    as.vector(brt.res$bias)),
-                        variance = c(as.vector(mat.res$variance),
+                        variance = c(as.vector(rand.res$variance),
+                                     as.vector(mat.res$variance),
                                    as.vector(wa.res$variance),
@@
                                    as.vector(brt.res$variance)),
-                        dissim = rep(rep(vals, each=nrow(wa.res$bias)),5),
-                        model  = rep(c('MAT', 'WA', 'WAPLS', 'rFor', 'BRT'), 
+                        dissim = rep(rep(vals, each=nrow(wa.res$bias)),6),
+                        model  = factor(rep(c('NULL','MAT', 'WA', 'WAPLS', 'rFor', 'BRT'), 
                                      each = length(wa.res$bias)),
-                        init = rep(rep(climate[,10], 101), 5))
+                                     levels=c('NULL', 'MAT', 'rFor', 'BRT', 'WA', 'WAPLS')),
+                        init = rep(rep(climate[,10], 101), 6))
 
@@
 
-samp.set <- sapply(c('MAT', 'WA', 'WAPLS', 'rFor', 'BRT'), sample.model)
+samp.set <- sapply(c('NULL','MAT', 'WA', 'WAPLS', 'rFor', 'BRT'), sample.model)
 
@@
   scale_y_continuous(expand=c(0,0), limits=c(0, 40)) +
+  coord_cartesian() +
   #scale_color_continuous(low = 'blue', high='pink', trans='sqrt') +
@@
   scale_y_sqrt(expand=c(0,0)) +
-  scale_color_continuous(low = 'blue', high='pink', trans='sqrt') +
+  #scale_color_continuous(low = 'blue', high='pink', trans='sqrt') +
   theme(legend.position = 'none',

### 2015-09-21 R/neotoma_historical.R

@@ -1,0 +1,26 @@
+library(neotoma)
+
+load('data/compiled.pollen.RData')
+
+all_melt <- melt(compiled.pollen, id.vars = c("sitename", "depth", "age", 
+                                              "date.type", "lat", "long", "dataset"))
+
+mod_pol$ID <- 1:nrow(mod_pol)
+neo_melt <- melt(mod_pol,id=c("ID","LATDD","LONDD","XXX","XXX.1","ELEVATION","PolSum"))
+
+trans_taxa <- read.csv('data/mod_to_neotoma.csv',header = FALSE)
+
+neo_melt$neo <- trans_taxa[match(neo_melt$variable,trans_taxa[,1]),2]
+
+all_melt <- all_melt[!all_melt$value ==0 & all_melt$variable %in% neo_melt$neo,]
+all_melt <- all_melt[!all_melt$date.type %in% "Radiocarbon years BP", ]
+
+neo_melt <- neo_melt[!neo_melt$value ==0,]
+neo_melt <- neo_melt[neo_melt$neo %in% levels(all_melt$variable),]
+
+all_points <- data.frame(point = c(all_melt$sitename, neo_melt$ID),
+                         age =   c(all_melt$age, rep("Modern", nrow(neo_melt))),
+                         value = c(all_melt$value, neo_melt$value),
+                         variable = c(as.character(all_melt$variable), as.character(neo_melt$variable)))
+
+all_cast <- dcast(all_melt, value ~ variable + age, fun=sum)
